### 2023-02-15 04. Analysis/scripts/01-data-preparation-joas.Rmd
Adding other examples

---
title: "Data preparation - JOAs"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: readable
    highlight: tango
    code_folding: hide
params:
  author: "Sam Kim and Ben Jaques-Leslie"
  project_number: "001"
  project_name: "Remote Work"
  data_folder_1: "03. Data Collection/Original data files from Melissa/JOAs"
  unique_identifier_1: !r c("usajobs_control_number")
  date_earliest_open: '2022-01-01'
  data_latest_close: '2022-12-01'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
tictoc::tic()
library(Hmisc)
library(tidyverse)
library(janitor)
library(skimr)
library(DataExplorer)
library(readr)
library(data.table)
library(lubridate)
library(arsenal)
```

# {.tabset .tabset-pills}

**Project number**: `r params$project_number`

**Project name**: `r params$project_name`

**Author**: `r params$author`

**Data folder 1**: `r params$data_folder`

**Unique identifier for data folder 1**: `r params$unique_identifier_1`

## Load data {.tabset .tabset-pills}

### Data files

The file names below represent all original candidate data files from Melissa.

```{r}
files <- 
  tibble(files = list.files(here::here(params$data_folder_1))
)

files
```

### Function to identify the encoding of the data files.

```{r}
encoding_data_frame <- function(in_file_name)
{
  readr::guess_encoding(here::here(params$data_folder_1,in_file_name)) %>% 
        mutate(file = in_file_name) %>% 
        select(file, everything())
}
```

### Identifying file encoding

Identify the encoding the candidates data files.

```{r}
files <- 
  files %>% 
  pull() %>% 
  map_dfr(~encoding_data_frame(.))

files
```

Filter candidate files to those with greatest confidence in the encoding.

```{r}
files <- 
  files %>% 
  group_by(file) %>% 
  filter(confidence == max(confidence))

files
```

### Functions

#### Function to load UTF-8 candidate files

```{r}
load_utf8 <- function(in_file_name, in_col_names)
{
  prep_01 <-
    fread(
      here::here(params$data_folder_1,in_file_name)
    ) %>% 
      distinct()

  return(prep_01)
}
```

#### Function to load UTF-16 candidate files

```{r}
load_utf16 <- function(in_file_name)
{
  prep_01 <- read.csv(
    file = here::here(params$data_folder_1,in_file_name),
    fileEncoding = "UTF-16",
    sep = "\t"
    )

  return(prep_01)
}
```

#### Function to prepare JOA info data.

```{r}
prepare_joa <- function(in_data)
{
prep_02 <- 
  in_data %>% 
  clean_names()

prep_03 <-
  prep_02 %>%
  select(everything())

if("announcement_closing_type" %in% names(prep_03))
{
  prep_04 <-
  prep_03 %>%
  mutate(
    across(
      .cols = c(position_open_date,position_close_date),
      .fns = ~ymd(.)
    )
  ) %>% 
  mutate(
    across(
      .cols = c(usajobs_control_number,announcement_closing_type,agency_level,hiring_department_code,series),
      .fns = ~as.character(.)
    )
  ) %>% 
  mutate(
    across(
      .cols = c(hiring_agency_code,total_openings),
      .fns = ~str_trim(.)
    )
  ) %>% 
  mutate(
    across(
      .cols = c(total_openings),
      .fns = ~as.numeric(.)
    )
  )
}

if(!("announcement_closing_type" %in% names(prep_03)) &
   !("agency_level" %in% names(prep_03)) &
   !("hiring_department_code" %in% names(prep_03)) &
   !("hiring_agency_code" %in% names(prep_03)) 
     )
{
  prep_04 <-
  prep_03 %>%
  mutate(
    across(
      .cols = c(position_open_date,position_close_date),
      .fns = ~ymd(.)
    )
  ) %>% 
  mutate(
    across(
      .cols = c(usajobs_control_number,
                # announcement_closing_type,
                # agency_level,
                # hiring_department_code,
                series),
      .fns = ~as.character(.)
    )
  ) %>% 
  mutate(
    across(
      .cols = c(
        # hiring_agency_code,
        total_openings),
      .fns = ~str_trim(.)
    )
  ) %>% 
  mutate(
    across(
      .cols = c(total_openings),
      .fns = ~as.numeric(.)
    )
  )
}

  #   position_open_date = ymd(position_open_date),
  #   position_close_date = ymd(position_close_date),
  #   usajobs_control_number = as.character(usajobs_control_number),
  #   race = str_remove(race, " - .*"),
  #   race = case_when(race == "" ~ "Did not disclose race",
  #                    is.na(race) ~ "Did not disclose race",
  #                    TRUE ~ race),
  #   degree_level = str_remove(degree_level, " \\(.*"),
  #   degree_level = case_when(degree_level == "" ~ "Did not disclose degree",
  #                    TRUE ~ degree_level),
  #   federal_employment_status =
  #     case_when(
  #       str_detect(
  #         federal_employment_status,"former") ~ "Former",
  #       str_detect(
  #         federal_employment_status,"currently") ~ "Current",
  #       str_detect(
  #         federal_employment_status,"never") ~ "Non-Federal employee"
  #       )
  # )
# 
# prep_05 <- 
#   prep_04 %>% 
#   select(candidate_id,race) %>% 
#   distinct() %>%
#   mutate(ind = 1) %>% 
#   pivot_wider(names_from = race, values_from = ind) %>% 
#   clean_names() %>% 
#   mutate(
#     across(
#       .cols = where(is.numeric),
#       .fns = ~replace_na(.,0)
#     )
#   )
# 
# prep_06 <-
#   prep_04 %>% 
#   select(candidate_id,degree_level) %>% 
#   distinct() %>% 
#   mutate(ind = 1) %>% 
#   pivot_wider(names_from = degree_level, values_from = ind) %>% 
#   clean_names() %>% 
#   mutate(
#     across(
#       .cols = where(is.numeric),
#       .fns = ~replace_na(.,0)
#     )
#   )
# 
# prep_07 <- 
#   prep_04 %>% 
#   select(-race,-degree_level) %>% 
#   distinct()
# 
# prep_08 <- 
#   prep_05 %>% 
#   left_join(prep_06) %>% 
#   left_join(prep_07)

return(prep_04)
}
```

#### Function to prepare JOA info data.

```{r}
prepare_joa_telework <- function(in_data)
{
prep_02 <- 
  in_data %>% 
  clean_names()

prep_03 <-
  prep_02 %>%
  select(everything())

if("is_tele_work_eligible" %in% names(prep_03))
{
  prep_04 <-
  prep_03 %>%
    rename(telework_eligible = is_tele_work_eligible)
}

if(!("is_tele_work_eligible" %in% names(prep_03)))
{
  prep_04 <-
  prep_03
}

prep_05 <- 
  prep_04 %>% 
  select(usajobs_control_number,telework_eligible) %>% 
  mutate(
    across(
      .cols = c(usajobs_control_number,
               ),
      .fns = ~as.character(.)
    )
  ) %>%
  mutate(
    across(
      .cols = c(telework_eligible),
      .fns = ~as.numeric(.)
    )
  )
# 
# if(!("announcement_closing_type" %in% names(prep_03)) &
#    !("agency_level" %in% names(prep_03)) &
#    !("hiring_department_code" %in% names(prep_03)) &
#    !("hiring_agency_code" %in% names(prep_03)) 
#      )
# {
#   prep_04 <-
#   prep_03 %>%
#   mutate(
#     across(
#       .cols = c(position_open_date,position_close_date),
#       .fns = ~ymd(.)
#     )
#   ) %>% 
#   mutate(
#     across(
#       .cols = c(usajobs_control_number,
#                 # announcement_closing_type,
#                 # agency_level,
#                 # hiring_department_code,
#                 series),
#       .fns = ~as.character(.)
#     )
#   ) %>% 
#   mutate(
#     across(
#       .cols = c(
#         # hiring_agency_code,
#         total_openings),
#       .fns = ~str_trim(.)
#     )
#   ) %>% 
#   mutate(
#     across(
#       .cols = c(total_openings),
#       .fns = ~as.numeric(.)
#     )
#   )
# }

  #   position_open_date = ymd(position_open_date),
  #   position_close_date = ymd(position_close_date),
  #   usajobs_control_number = as.character(usajobs_control_number),
  #   race = str_remove(race, " - .*"),
  #   race = case_when(race == "" ~ "Did not disclose race",
  #                    is.na(race) ~ "Did not disclose race",
  #                    TRUE ~ race),
  #   degree_level = str_remove(degree_level, " \\(.*"),
  #   degree_level = case_when(degree_level == "" ~ "Did not disclose degree",
  #                    TRUE ~ degree_level),
  #   federal_employment_status =
  #     case_when(
  #       str_detect(
  #         federal_employment_status,"former") ~ "Former",
  #       str_detect(
  #         federal_employment_status,"currently") ~ "Current",
  #       str_detect(
  #         federal_employment_status,"never") ~ "Non-Federal employee"
  #       )
  # )
# 
# prep_05 <- 
#   prep_04 %>% 
#   select(candidate_id,race) %>% 
#   distinct() %>%
#   mutate(ind = 1) %>% 
#   pivot_wider(names_from = race, values_from = ind) %>% 
#   clean_names() %>% 
#   mutate(
#     across(
#       .cols = where(is.numeric),
#       .fns = ~replace_na(.,0)
#     )
#   )
# 
# prep_06 <-
#   prep_04 %>% 
#   select(candidate_id,degree_level) %>% 
#   distinct() %>% 
#   mutate(ind = 1) %>% 
#   pivot_wider(names_from = degree_level, values_from = ind) %>% 
#   clean_names() %>% 
#   mutate(
#     across(
#       .cols = where(is.numeric),
#       .fns = ~replace_na(.,0)
#     )
#   )
# 
# prep_07 <- 
#   prep_04 %>% 
#   select(-race,-degree_level) %>% 
#   distinct()
# 
# prep_08 <- 
#   prep_05 %>% 
#   left_join(prep_06) %>% 
#   left_join(prep_07)

return(prep_05)
}
```

#### Function to load and prepare JOA information

```{r}
load_prepare_joa <- function(in_file_name,in_encoding)
{
  if(in_encoding == 'ASCII')
  {
    prep_01 <- in_file_name %>% 
  load_utf8()
  }

  if(in_encoding == 'UTF-16LE')
  {
    prep_01 <- in_file_name %>% 
  load_utf16()
  }

  out <- prep_01 %>% 
  prepare_joa()

  return(out)

}
```

#### Function to load and prepare JOA telework information

```{r}
load_prepare_joa_telework <- function(in_file_name,in_encoding)
{
  if(in_encoding == 'ASCII')
  {
    prep_01 <- in_file_name %>% 
  load_utf8()
  }

  if(in_encoding == 'UTF-16LE')
  {
    prep_01 <- in_file_name %>% 
  load_utf16()
  }

  out <- prep_01 %>% 
  prepare_joa_telework()

  return(out)

}
```

#### Function to convert character to indicator

```{r}
character_to_indicator <- function(in_data,in_unique_identifier,in_variable,in_prefix)
{
  in_data %>% 
  select(all_of(in_unique_identifier),{{in_variable}}) %>% 
    distinct() %>% 
    mutate(ind = 1) %>% 
    mutate(across(
      .cols = {{in_variable}},
      .fns = ~case_when(. == "" ~ "missing",
                       TRUE ~ .)
    )) %>% 
    pivot_wider(names_from = {{in_variable}},
                values_from = ind) %>% 
    rename_with(
      .fn = ~glue::glue("{in_prefix} {.}"),
  .cols = -in_unique_identifier
  ) %>% 
    clean_names() %>% 
    mutate(
      across(
        .cols = where(is.numeric),
        .fns = ~replace_na(.,0)
      )
    )
}
```

### JOA data

#### Identify files with "full" in title or "Positions_Updated_Dec2022"

```{r}
files_full <- 
  files %>% 
  filter(str_detect(file,"_full_") | str_detect(file,"Positions_Updated_Dec2022")) %>%
  select(file,encoding)

files_full
```

#### Load 'full' files

```{r}
prep_base <- 
  map2_dfr(
    .x = files_full$file,
    .y = files_full$encoding,
    .f = ~load_prepare_joa(in_file_name = .x, in_encoding = .y)
  ) %>% 
  distinct()
```

#### Identify files with "remote_telework_positions" in title or "telework_eligible_from_09092022"

```{r}
files_telework <- 
  files %>% 
  filter(str_detect(file,"remote_telework_positions") | str_detect(file,"telework_eligible_from_09092022")) %>%
  select(file,encoding)

files_telework
```

#### Load 'telework' files

```{r}
prep_telework <- 
  map2_dfr(
    .x = files_telework$file,
    .y = files_telework$encoding,
    .f = ~load_prepare_joa_telework(in_file_name = .x, in_encoding = .y)
  ) %>% 
  distinct()
```

```{r}
prep_telework %>% 
  skim()
```

## Filter to correct dates

```{r}
prep_base <- 
  prep_base %>% 
  filter(position_open_date >= ymd(params$date_earliest_open) &
           position_close_date <= ymd(params$data_latest_close))

```

#### Review data

```{r}
prep_base %>%
  skim()
```

## Addressing duplication

Melissa recommended using `r params$unique_identifier_1` as the unique identifier.

To review specific cases you to https://www.usajobs.gov/job/ and put the USAJobs control number at the end.

### Check for duplication

```{r}
dupe_01 <- 
  prep_base %>%
  get_dupes(params$unique_identifier_1)
```

#### Data review

```{r}
dupe_01 %>% 
  skim()
```

### Dataframe of the unique identifier

```{r}
prep_uniq <- 
  prep_base %>% 
  select(params$unique_identifier_1) %>% 
  distinct()

prep_uniq %>% 
  count()
```

### Identifying and categorizing columns

Created a dataframe of variable metadata

```{r}
column_meta_data <- 
  prep_base %>% 
  contents()

column_meta_data <- 
  column_meta_data$contents %>% 
  as.data.frame()

column_meta_data <-
  column_meta_data %>% 
  rownames_to_column(var = "variable")

column_meta_data
```

Function that checks whether there are duplicates for different variables

```{r}
dupe_check <- function(in_data,in_variable_name)
{
  check_01 <- 
  in_data %>% 
  select(params$unique_identifier_1,all_of(in_variable_name)) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

out <- nrow(check_01) > 0

return(out)
}
```

Checking duplicates for each variable in the dataframe.

```{r}
duplicates_by_variable <- 
  column_meta_data %>% 
  pull(variable) %>% 
  map_lgl(~dupe_check(in_data = dupe_01, .))

duplicates_by_variable
```

Add this to the column_meta_data dataframe.

```{r}
column_meta_data <- 
  column_meta_data %>% 
  add_column(
    duplicates_by_variable = duplicates_by_variable
  )

column_meta_data
```

Create dataframe of only variables that cause duplicates.

```{r}
duplicating_variables <- 
  column_meta_data %>% 
  filter(duplicates_by_variable)

duplicating_variables
```

Grouping variables.

```{r}
duplicating_variables <- 
  duplicating_variables %>% 
  mutate(group = 
           case_when(variable == "position_open_date" ~ "dates",
                     variable == "position_close_date" ~ "dates",
                     variable == "announcement_closing_type" ~ "announcement_closing_type",
                     variable == "announcement_closing_type_description" ~ "announcement_closing_type",
                     variable == "application_limit" ~ "application_limit",
                     variable == "application_limit_set" ~ "application_limit",
                     variable == "agency_level" ~ "agency_level",
                     variable == "agency_level_sort" ~ "agency_level",
                     variable == "hiring_agency_code" ~ "agency_level",
                     variable == "hiring_agency_name" ~ "agency_level",
                     variable == "hiring_department_code" ~ "agency_level",
                     variable == "hiring_department_name" ~ "agency_level",
                     variable == "maximum_salary" ~ "maximum_salary",
                     variable == "minimum_salary" ~ "maximum_salary",
                     variable == "maximum_grade" ~ "maximum_salary",
                     variable == "minimum_grade" ~ "maximum_salary",
                     variable == "series" ~ "series",
                     variable == "series_title" ~ "series",
                     TRUE ~ variable
                     ))

duplicating_variables
```

Create list of groups.

```{r}
duplicating_groups <- 
  duplicating_variables %>% 
  select(group) %>% 
  distinct() %>% 
  pull()

duplicating_groups
```

### Duplication removal

```{r}
step_n <- 0
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)
```

Some of these duplicates may be due to small typos. Transforming the string to upper case.

```{r}
dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  mutate(
    announcement_number = str_to_upper(announcement_number)
  ) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)
```

This fixes some, but not all.

Without better information, we will take the lower of the two values for the remainder.

```{r}
dedupe <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  mutate(
    announcement_number = str_to_upper(announcement_number)
  ) %>% 
  distinct() %>% 
  arrange(params$unique_identifier_1,announcement_number) %>% 
  group_by_at(params$unique_identifier_1) %>% 
  summarise(announcement_number = first(announcement_number))

dedupe %>% 
  get_dupes(params$unique_identifier_1)
```

Creating a deduplicated dataframe of `r params$unique_identifier_1` and `r duplicating_groups[step_n]`.

First create a dataframe of distinct rows of the variables from the original data.

```{r}
working_01 <- 
  prep_base %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct()

working_01 %>% 
  glimpse()

working_01 %>% 
  skim()
```

Next remove rows that dataframe that were found to be duplicated and save that as a new dataframe.

```{r}
working_02 <- 
  working_01 %>% 
  anti_join(dedupe %>% select(params$unique_identifier_1))

working_02 %>% 
  glimpse()

working_02 %>% 
  skim()
```

Last bind this dataframe to the deduplicated set above.

```{r}
prep_xx <- 
  working_02 %>% 
  bind_rows(dedupe)

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Rename dataframe and count rows.

```{r}
prep_01 <- prep_xx

prep_01 %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_01 %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Duplication appear to be because of difference in open and close dates. After messaging with Melissa, we decide to retain the later dates as determined by the close date.

```{r}
dedupe <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>%
  arrange(params$unique_identifier_1, position_close_date) %>% 
  group_by(usajobs_control_number) %>% 
  summarise(
    across(
      .cols = all_of(duplicating_group %>% pull(variable)),
      .fns = last
    )
  )

dedupe %>% 
  get_dupes(params$unique_identifier_1)
```

Creating a deduplicated dataframe of `r params$unique_identifier_1` and `r duplicating_groups[step_n]`.

First create a dataframe of distinct rows of the variables from the original data.

```{r}
working_01 <- 
  prep_base %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct()

working_01 %>% 
  glimpse()

working_01 %>% 
  skim()
```

Next remove rows that dataframe that were found to be duplicated and save that as a new dataframe.

```{r}
working_02 <- 
  working_01 %>% 
  anti_join(dedupe %>% select(params$unique_identifier_1))

working_02 %>% 
  glimpse()

working_02 %>% 
  skim()
```

Last bind this dataframe to the deduplicated set above.

```{r}
prep_xx <- 
  working_02 %>% 
  bind_rows(dedupe)

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Rename dataframe

```{r}
prep_02 <- prep_xx
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Duplication appear to be because `r duplicating_group %>% pull(variable)` is empty in one row. Transform the empties into 'N'.

We want to have this change to affect the whole dataframe so we address the duplicates off the original dataframe.

```{r}
prep_xx <- 
  prep_base %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  mutate(
    is_remote_work_eligible = case_when(is_remote_work_eligible == "" ~ "N",
                                        TRUE ~ is_remote_work_eligible)
  ) %>% 
  distinct() %>% 
  character_to_indicator(in_unique_identifier = params$unique_identifier_1,
                         in_variable = is_remote_work_eligible,
                         in_prefix = "is_remote_work_eligible") %>% 
  # Drop the "n" indicatoris_remote_work_eligible
  select(-is_remote_work_eligible_n)

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Rename dataframe

```{r}
prep_03 <- prep_xx
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Check to see if the type codes are consistent to the description.

```{r}
prep_base %>% 
  count(announcement_closing_type,announcement_closing_type_description)
```

Codes are consistent to description, but the code adds no new information. We will drop the code and transform the description into columns of indicators.

We want to have this change to affect the whole dataframe so we address the duplicates off the original dataframe.

```{r}
prep_xx <- 
  prep_base %>% 
  character_to_indicator(params$unique_identifier_1,announcement_closing_type_description,"closing_type")

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Rename dataframe and count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Rename dataframe

```{r}
prep_04 <- prep_xx

```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Announcements have multiple open statuses.

First count the statuses across the original data.

```{r}
prep_base %>% count(position_opening_status)
```

Not too many of them. Like above, we will transform these into indicator columns.

We want to have this change to affect the whole dataframe so we address the duplicates off the original dataframe.

```{r}
prep_xx <- 
  prep_base %>% 
  character_to_indicator(params$unique_identifier_1,position_opening_status,"position_opening_status")

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```
Rename dataframe

```{r}
prep_05 <- prep_xx

```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Not too many. It appears that `application_limit_set` is "N" when `application_limit` is "Applicant Cut-Off".

First count the statuses across the original data.

```{r}
prep_base %>% count(application_limit,application_limit_set)
```

All NA `application_limit` are "N" `application_limit_set`.

We will consider "Y" the correct response for all and we will change this into an indicator.

We want to have this change to affect the whole dataframe so we address the duplicates off the original dataframe.

```{r}
prep_xx <- 
  prep_base %>% 
  character_to_indicator(params$unique_identifier_1,application_limit_set,"application_limit_set") %>% 
  # Drop the "N" column
  select(-application_limit_set_n)

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Rename dataframe

```{r}
prep_06 <- prep_xx
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Not too many. It appears that `work_schedule` might have different.

First count the statuses across the original data.

```{r}
prep_base %>% count(work_schedule)
```

There are just a few, but we'll want to transform these into indicators.

We want to have this change to affect the whole dataframe so we address the duplicates off the original dataframe.

```{r}
prep_xx <- 
  prep_base %>% 
  character_to_indicator(params$unique_identifier_1,work_schedule,"work_schedule")

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Rename dataframe

```{r}
prep_07 <- prep_xx
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Not too many. It appears that `disable_apply_online` might have different.

First count the statuses across the original data.

```{r}
prep_base %>% count(disable_apply_online)
```

There are just a few, but we'll want to transform these into indicators.

We want to have this change to affect the whole dataframe so we address the duplicates off the original dataframe.

Since there is only one instance of this type of duplication, we will still drop the no column.

```{r}
prep_xx <- 
  prep_base %>% 
  character_to_indicator(params$unique_identifier_1,disable_apply_online,"disable_apply_online") %>% 
  # Drop the "N" column
  select(-disable_apply_online_n)

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Rename dataframe

```{r}
prep_08 <- prep_xx
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Not too many.

First count the statuses across the original data.

```{r}
prep_base %>% count(appointment_type)
```

There are just a few, but we'll want to transform these into indicators.

We want to have this change to affect the whole dataframe so we address the duplicates off the original dataframe.

Since there is only one instance of this type of duplication, we will still drop the no column.

```{r}
prep_xx <- 
  prep_base %>% 
  character_to_indicator(params$unique_identifier_1,appointment_type,"appointment_type")

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Rename dataframe

```{r}
prep_09 <- prep_xx
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

A few things going on here:

1.    Department of Defense agencies are duplicated because sometimes sub-departments (e.i. Department of Navy) are considered the *Department*
2.    Many duplicates are caused by blank values in hiring agency code

First let's remove duplicates because of department. To do this, we'll use the maximum agency level to limit to just the entries with the most levels.

```{r}
dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  group_by(usajobs_control_number) %>% 
  filter(agency_level == max(agency_level)) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)
```

We can still see the duplications from blank hiring agency codes. We'll drop those.

```{r}
dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  group_by(usajobs_control_number) %>% 
  filter(agency_level == max(agency_level)) %>% 
  filter(!is.na(hiring_department_code)) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)
```

We still have one duplicate remaining. For this we'll, select the entry with the later position closing date.

```{r}
dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable)),position_close_date) %>% 
  group_by(usajobs_control_number) %>% 
  filter(agency_level == max(agency_level)) %>% 
  filter(!is.na(hiring_department_code)) %>% 
  filter(position_close_date == max(position_close_date)) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)
```

Now we will apply all of these to the duplicated rows.

```{r}
dedupe <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable)),position_close_date) %>% 
  group_by(usajobs_control_number) %>% 
  filter(agency_level == max(agency_level)) %>% 
  filter(!is.na(hiring_department_code)) %>% 
  filter(position_close_date == max(position_close_date)) %>% 
  distinct() %>% 
  ungroup() %>% 
  select(-position_close_date)

dedupe %>% 
  get_dupes(params$unique_identifier_1)
```

Creating a deduplicated dataframe of `r params$unique_identifier_1` and `r duplicating_groups[step_n]`.

First create a dataframe of distinct rows of the variables from the original data.

```{r}
working_01 <- 
  prep_base %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct()

working_01 %>% 
  glimpse()

working_01 %>% 
  skim()
```

Next remove rows that dataframe that were found to be duplicated and save that as a new dataframe.

```{r}
working_02 <- 
  working_01 %>% 
  anti_join(dedupe %>% select(params$unique_identifier_1))

working_02 %>% 
  glimpse()

working_02 %>% 
  skim()
```

Last bind this dataframe to the deduplicated set above.

```{r}
prep_xx <- 
  working_02 %>% 
  bind_rows(dedupe)

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Rename dataframe

```{r}
prep_10 <- prep_xx
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Count types.

```{r}
prep_base %>% 
  count(hiring_subelement_name)
```

Many blank entries and a level of detail we don't need. We will drop this column.

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Appears like this comes from posting the job again with revised salary information.

Like above we decide to retain the later dates as determined by the close date.

```{r}
dedupe <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable)),position_close_date) %>%
  arrange(params$unique_identifier_1, position_close_date) %>% 
  group_by(usajobs_control_number) %>% 
  summarise(
    across(
      .cols = all_of(duplicating_group %>% pull(variable)),
      .fns = last
    )
  )

dedupe %>% 
  get_dupes(params$unique_identifier_1)
```

Creating a deduplicated dataframe of `r params$unique_identifier_1` and `r duplicating_groups[step_n]`.

First create a dataframe of distinct rows of the variables from the original data.

```{r}
working_01 <- 
  prep_base %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct()

working_01 %>% 
  glimpse()

working_01 %>% 
  skim()
```

Next remove rows that dataframe that were found to be duplicated and save that as a new dataframe.

```{r}
working_02 <- 
  working_01 %>% 
  anti_join(dedupe %>% select(params$unique_identifier_1))

working_02 %>% 
  glimpse()

working_02 %>% 
  skim()
```

Last bind this dataframe to the deduplicated set above.

```{r}
prep_xx <- 
  working_02 %>% 
  bind_rows(dedupe)

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Rename dataframe

```{r}
prep_11 <- prep_xx
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```
Count types.

```{r}
prep_base %>% 
  count(series, series_title)
```

Many different titles and series.

Check if the series numbers are duplicated.

```{r}
prep_base %>% 
  count(series, series_title) %>% 
  ungroup() %>% 
  get_dupes(series)
```

None.

Now the titles.

```{r}
prep_base %>% 
  count(series, series_title) %>% 
  ungroup() %>% 
  get_dupes(series_title)
```

None.

Since the `series` adds no information and is uniquely matched to the titles, we will drop `series`.

There are too many series to want indicators, we'll use the last closing date to select the appropriate series.

Like above we decide to retain the later dates as determined by the close date.

```{r}
dedupe <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable)),position_close_date) %>%
  arrange(params$unique_identifier_1, position_close_date) %>% 
  group_by(usajobs_control_number) %>% 
  summarise(
    across(
      .cols = all_of(duplicating_group %>% pull(variable)),
      .fns = last
    )
  )

dedupe %>% 
  get_dupes(params$unique_identifier_1)
```

Creating a deduplicated dataframe of `r params$unique_identifier_1` and `r duplicating_groups[step_n]`.

First create a dataframe of distinct rows of the variables from the original data.

```{r}
working_01 <- 
  prep_base %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct()

working_01 %>% 
  glimpse()

working_01 %>% 
  skim()
```

Next remove rows that dataframe that were found to be duplicated and save that as a new dataframe.

```{r}
working_02 <- 
  working_01 %>% 
  anti_join(dedupe %>% select(params$unique_identifier_1))

working_02 %>% 
  glimpse()

working_02 %>% 
  skim()
```

Last bind this dataframe to the deduplicated set above.

```{r}
prep_xx <- 
  working_02 %>% 
  bind_rows(dedupe) %>% 
  select(-series)

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Rename dataframe

```{r}
prep_12 <- prep_xx
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Not too many duplicates. We'll follow the same procedure, taking the number from the row wiht the later closing date.

Like above we decide to retain the later dates as determined by the close date.

```{r}
dedupe <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable)),position_close_date) %>%
  arrange(params$unique_identifier_1, position_close_date) %>% 
  group_by(usajobs_control_number) %>% 
  summarise(
    across(
      .cols = all_of(duplicating_group %>% pull(variable)),
      .fns = last
    )
  )

dedupe %>% 
  get_dupes(params$unique_identifier_1)
```

Creating a deduplicated dataframe of `r params$unique_identifier_1` and `r duplicating_groups[step_n]`.

First create a dataframe of distinct rows of the variables from the original data.

```{r}
working_01 <- 
  prep_base %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct()

working_01 %>% 
  glimpse()

working_01 %>% 
  skim()
```

Next remove rows that dataframe that were found to be duplicated and save that as a new dataframe.

```{r}
working_02 <- 
  working_01 %>% 
  anti_join(dedupe %>% select(params$unique_identifier_1))

working_02 %>% 
  glimpse()

working_02 %>% 
  skim()
```

Last bind this dataframe to the deduplicated set above.

```{r}
prep_xx <- 
  working_02 %>% 
  bind_rows(dedupe)

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Rename dataframe

```{r}
prep_13 <- prep_xx
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Not too many duplicates.

Like above we decide to retain the later dates as determined by the close date.

```{r}
dedupe <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable)),position_close_date) %>%
  arrange(params$unique_identifier_1, position_close_date) %>% 
  group_by(usajobs_control_number) %>% 
  summarise(
    across(
      .cols = all_of(duplicating_group %>% pull(variable)),
      .fns = last
    )
  )

dedupe %>% 
  get_dupes(params$unique_identifier_1)
```

Creating a deduplicated dataframe of `r params$unique_identifier_1` and `r duplicating_groups[step_n]`.

First create a dataframe of distinct rows of the variables from the original data.

```{r}
working_01 <- 
  prep_base %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct()

working_01 %>% 
  glimpse()

working_01 %>% 
  skim()
```

Next remove rows that dataframe that were found to be duplicated and save that as a new dataframe.

```{r}
working_02 <- 
  working_01 %>% 
  anti_join(dedupe %>% select(params$unique_identifier_1))

working_02 %>% 
  glimpse()

working_02 %>% 
  skim()
```

Last bind this dataframe to the deduplicated set above.

```{r}
prep_xx <- 
  working_02 %>% 
  bind_rows(dedupe)

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Rename dataframe

```{r}
prep_14 <- prep_xx
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

No more deplicated columns!!

### Combinding deduplicated data

#### Created dataframe on non-duplicating columns

Collect column names of variables not creating duplicates.

```{r}
non_duplicating_variables <- 
  column_meta_data %>% 
  filter(!duplicates_by_variable) %>% 
  pull(variable)

non_duplicating_variables
```

Create dataframe of only these variables.

```{r}
prep_xx <- 
  prep_base %>% 
  select(any_of(non_duplicating_variables)) %>% 
  distinct()

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Rename dataframe.

```{r}
prep_11 <- prep_xx
```

#### Join all deduplicated dataframes

```{r}
prep_xx <- 
  ls(envir=.GlobalEnv) %>% 
  as_tibble() %>% 
  filter(str_detect(value, "prep_[0-9]")) %>% 
  pull() %>% 
  mget(envir=.GlobalEnv) %>% 
  reduce(full_join)

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Rename dataframe.

```{r}
prep_dedup <- prep_xx
```

### Checking duplicating columns have been fixed

Created a dataframe of variable metadata

```{r}
column_meta_data_dedupe <- 
  prep_dedup %>% 
  contents()

column_meta_data_dedupe <- 
  column_meta_data_dedupe$contents %>% 
  as.data.frame()

column_meta_data_dedupe <-
  column_meta_data_dedupe %>% 
  rownames_to_column(var = "variable")

column_meta_data_dedupe
```

Checking duplicates for each variable in the dataframe.

```{r}
duplicates_by_variable <- 
  column_meta_data_dedupe %>% 
  pull(variable) %>% 
  map_lgl(~dupe_check(in_data = prep_dedup, .))

duplicates_by_variable
```

Add this to the column_meta_data dataframe.

```{r}
column_meta_data_dedupe <- 
  column_meta_data_dedupe %>% 
  add_column(
    duplicates_by_variable = duplicates_by_variable
  )

column_meta_data_dedupe
```

Create dataframe of only variables that cause duplicates.

```{r}
duplicating_variables <- 
  column_meta_data_dedupe %>% 
  filter(duplicates_by_variable)

duplicating_variables
```

## Addressing blanks

Created a dataframe of variable metadata for deduplicated data

```{r}
column_meta_data_dedupe <- 
  prep_dedup %>% 
  contents()

column_meta_data_dedupe <- 
  column_meta_data_dedupe$contents %>% 
  as.data.frame()

column_meta_data_dedupe <-
  column_meta_data_dedupe %>% 
  rownames_to_column(var = "variable")

column_meta_data_dedupe
```

Function that checks whether there are duplicates for different variables

```{r}
blank_check <- function(in_data,in_variable_name)
{
  check_01 <- 
  in_data %>% 
    filter(!!sym(in_variable_name) == "") %>% 
  distinct()

out <- nrow(check_01) > 0

return(out)
}
```

Checking duplicates for each variable in the dataframe.

```{r}
blanks_by_variable <- 
  column_meta_data_dedupe %>% 
  pull(variable) %>% 
  map_lgl(~blank_check(in_data = prep_dedup,.))

blanks_by_variable
```

Add this to the `column_meta_data_dedupe` dataframe.

```{r}
column_meta_data_dedupe <- 
  column_meta_data_dedupe %>% 
  add_column(
    blanks_by_variable = blanks_by_variable
  )

column_meta_data_dedupe
```

Create dataframe of only variables that have blanks.

```{r}
blanking_variables <- 
  column_meta_data_dedupe %>% 
  filter(blanks_by_variable)

blanking_variables
```

Grouping variables.

```{r}
blanking_variables <- 
  blanking_variables %>% 
  mutate(group = 
           case_when(variable == "position_open_date" ~ "dates",
                     variable == "position_close_date" ~ "dates",
                     variable == "announcement_closing_type" ~ "announcement_closing_type",
                     variable == "announcement_closing_type_description" ~ "announcement_closing_type",
                     variable == "application_limit" ~ "application_limit",
                     variable == "application_limit_set" ~ "application_limit",
                     variable == "agency_level" ~ "agency_level",
                     variable == "agency_level_sort" ~ "agency_level",
                     variable == "hiring_agency_code" ~ "agency_level",
                     variable == "hiring_agency_name" ~ "agency_level",
                     variable == "hiring_department_code" ~ "agency_level",
                     variable == "hiring_department_name" ~ "agency_level",
                     variable == "maximum_salary" ~ "maximum_salary",
                     variable == "minimum_salary" ~ "maximum_salary",
                     variable == "maximum_grade" ~ "maximum_salary",
                     variable == "minimum_grade" ~ "maximum_salary",
                     variable == "series" ~ "series",
                     variable == "series_title" ~ "series",
                     TRUE ~ variable
                     ))

blanking_variables
```

Create list of groups.

```{r}
blanking_groups <- 
  blanking_variables %>% 
  select(group) %>% 
  distinct() %>% 
  pull()

blanking_groups
```

### Blanks removal

```{r}
step_n <- 0
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing blanks produced by:

```{r}
blanking_variable <- 
  blanking_variables %>% 
  filter(group == blanking_groups[step_n])

blanking_variable
```

Reviewing blanks.

```{r}
blanks_01 <- 
  prep_dedup %>% 
  filter(if_any(all_of(blanking_variable %>% pull(variable)),  ~.x == "")) %>% 
  select(all_of(blanking_variable %>% pull(variable)),everything())

blanks_01
```

Blanks appear to be caused by a few columns:

1.    `hiring_agency_code` appears to be blank for agencies that don't fall into departments
2.    `hiring_agency_name` appears to be blank for the Air Force

Let's count the extent of the issues.

```{r}
blanking_variable %>% 
  pull(variable) %>% 
  map(
    ~blanks_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(!!sym(.x) == "")
  )
```

A lot of blanks in `hiring_agency_code`, some in `hiring_agency_name`, and very few in `hiring_department_name`.

Let's test filling blanks in `hiring_agency_name` with `hiring_department_name`.

```{r}
blanks_02 <- 
  blanks_01 %>% 
    select(starts_with("hiring_agency_"),starts_with("hiring_department_name"),everything()) %>%
  mutate(
    hiring_agency_name = case_when(
      hiring_agency_name == "" & !is.na(hiring_department_name) ~ hiring_department_name,
      TRUE ~ hiring_agency_name
      ),
    ) 
```

Reviewing blanks.

```{r}
  blanks_02 %>% 
  filter(if_any(all_of(blanking_variable %>% pull(variable)),  ~.x == "")) %>% 
  select(all_of(blanking_variable %>% pull(variable)),everything())
```

Reduced a lot.

Let's count the extent of the issues.

```{r}
blanking_variable %>% 
  pull(variable) %>% 
  map(
    ~blanks_02 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(!!sym(.x) == "")
  )
```

All the `hiring_agency_name` are gone.

Let's look into the issues with `hiring_department_name`.

```{r}
blanks_02 %>%
  filter(hiring_department_name == "")
```

Let's check what `hiring_department_name` this `hiring_agency_name` has in the larger data.

```{r}
prep_dedup %>% 
  filter(hiring_agency_name == "Public Defender Service for the District of Columbia") %>% 
  count(hiring_agency_name,hiring_department_name)
```

This suggests we should replace `hiring_department_name` this `hiring_agency_name`.

Let's do both steps.

```{r}
blanks_03 <- 
  blanks_01 %>% 
    select(starts_with("hiring_agency_"),starts_with("hiring_department_name"),everything()) %>%
  mutate(
    hiring_agency_name = case_when(
      hiring_agency_name == "" & !is.na(hiring_department_name) ~ hiring_department_name,
      TRUE ~ hiring_agency_name
      ),
    hiring_department_name = case_when(
      hiring_department_name == "" & !is.na(hiring_agency_name) ~ hiring_agency_name,
      TRUE ~ hiring_department_name
      ),
    ) 
```

Reviewing blanks.

```{r}
  blanks_03 %>% 
  filter(if_any(all_of(blanking_variable %>% pull(variable)),  ~.x == "")) %>% 
  select(all_of(blanking_variable %>% pull(variable)),everything())
```

Reduced a lot.

Let's count the extent of the issues.

```{r}
blanking_variable %>% 
  pull(variable) %>% 
  map(
    ~blanks_03 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(!!sym(.x) == "")
  )
```

Still have many issues with the `hiring_agency_code`. `hiring_agency_name` appears to have all of the important information and is more complete. So we will drop `hiring_agency_code`.

Now apply the steps above to the larger dataframe.

```{r}
prep_xx <- 
  prep_dedup %>% 
  mutate(
    hiring_agency_name = case_when(
      hiring_agency_name == "" & !is.na(hiring_department_name) ~ hiring_department_name,
      TRUE ~ hiring_agency_name
      ),
    hiring_department_name = case_when(
      hiring_department_name == "" & !is.na(hiring_agency_name) ~ hiring_agency_name,
      TRUE ~ hiring_department_name
      ),
    ) %>% 
  select(-hiring_agency_code)
```

Check for blanks.

```{r}
prep_xx %>% 
  filter(if_any(all_of(blanking_variable %>% filter(variable != 'hiring_agency_code') %>%  pull(variable)),  ~.x == ""))
```

Rename dataframe

```{r}
prep_blank <- prep_xx
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing blanks produced by:

```{r}
blanking_variable <- 
  blanking_variables %>% 
  filter(group == blanking_groups[step_n])

blanking_variable
```

Reviewing blanks.

```{r}
blanks_01 <- 
  prep_dedup %>% 
  filter(if_any(all_of(blanking_variable %>% pull(variable)),  ~.x == "")) %>% 
  select(all_of(blanking_variable %>% pull(variable)),everything())

blanks_01
```

Looks like `who_may_apply` is just blank.

Let's count the extent of the issues.

```{r}
blanking_variable %>% 
  pull(variable) %>% 
  map(
    ~blanks_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(!!sym(.x) == "")
  )
```

They are all blank. We will drop the `who_may_apply` column.

```{r}
prep_xx <- 
  prep_blank %>% 
  select(-who_may_apply)
```

Rename dataframe

```{r}
prep_blank <- prep_xx
```

```{r}
step_n <- step_n + 1
```

### Check that blanks are removed

Created a dataframe of variable metadata for deduplicated data

```{r}
column_meta_data_blank <- 
  prep_blank %>% 
  contents()

column_meta_data_blank <- 
  column_meta_data_blank$contents %>% 
  as.data.frame()

column_meta_data_blank <-
  column_meta_data_blank %>% 
  rownames_to_column(var = "variable")

column_meta_data_blank
```

Checking duplicates for each variable in the dataframe.

```{r}
blanks_by_variable <- 
  column_meta_data_blank %>% 
  pull(variable) %>% 
  map_lgl(~blank_check(in_data = prep_blank,.))

blanks_by_variable
```

Add this to the `column_meta_data_dedupe` dataframe.

```{r}
column_meta_data_blank <- 
  column_meta_data_blank %>% 
  add_column(
    blanks_by_variable = blanks_by_variable
  )

column_meta_data_blank
```

Create dataframe of only variables that have blanks.

```{r}
blanking_variables <- 
  column_meta_data_blank %>% 
  filter(blanks_by_variable)

blanking_variables
```

## Addressing nulls

Created a dataframe of variable metadata for blanks data

```{r}
column_meta_data_null <- 
  prep_blank %>% 
  contents()

column_meta_data_null <- 
  column_meta_data_null$contents %>% 
  as.data.frame()

column_meta_data_null <-
  column_meta_data_null %>% 
  rownames_to_column(var = "variable")

column_meta_data_null
```

Function that checks whether there are duplicates for different variables

```{r}
null_check <- function(in_data,in_variable_name)
{
  check_01 <- 
  in_data %>% 
    filter(is.na(!!sym(in_variable_name))) %>% 
  distinct()

out <- nrow(check_01) > 0

return(out)
}
```

Checking duplicates for each variable in the dataframe.

```{r}
nulls_by_variable <- 
  column_meta_data_null %>% 
  pull(variable) %>% 
  map_lgl(~null_check(in_data = prep_dedup,.))

nulls_by_variable
```

Add this to the `column_meta_data_null` dataframe.

```{r}
column_meta_data_null <- 
  column_meta_data_null %>% 
  add_column(
    nulls_by_variable = nulls_by_variable
  )

column_meta_data_null
```

Create dataframe of only variables that have blanks.

```{r}
null_variables <- 
  column_meta_data_null %>% 
  filter(nulls_by_variable)

null_variables
```

Grouping variables.

```{r}
null_variables <- 
  null_variables %>% 
  mutate(group = 
           case_when(variable == "position_open_date" ~ "dates",
                     variable == "position_close_date" ~ "dates",
                     variable == "announcement_closing_type" ~ "announcement_closing_type",
                     variable == "announcement_closing_type_description" ~ "announcement_closing_type",
                     variable == "application_limit" ~ "application_limit",
                     variable == "application_limit_set" ~ "application_limit",
                     variable == "agency_level" ~ "agency_level",
                     variable == "agency_level_sort" ~ "agency_level",
                     variable == "hiring_agency_code" ~ "agency_level",
                     variable == "hiring_agency_name" ~ "agency_level",
                     variable == "hiring_department_code" ~ "agency_level",
                     variable == "hiring_department_name" ~ "agency_level",
                     variable == "maximum_salary" ~ "maximum_salary",
                     variable == "minimum_salary" ~ "maximum_salary",
                     variable == "maximum_grade" ~ "maximum_salary",
                     variable == "minimum_grade" ~ "maximum_salary",
                     variable == "series" ~ "series",
                     variable == "series_title" ~ "series",
                     TRUE ~ variable
                     ))

null_variables
```

Create list of groups.

```{r}
null_groups <- 
  null_variables %>% 
  select(group) %>% 
  distinct() %>% 
  pull()

null_groups
```

### Nulls removal

```{r}
step_n <- 0
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing nulls produced by:

```{r}
null_variable <- 
  null_variables %>% 
  filter(group == null_groups[step_n])

null_variable
```

Reviewing blanks.

```{r}
null_01 <- 
  prep_blank %>% 
  filter(if_any(all_of(null_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(null_variable %>% pull(variable)),everything())

null_01
```

Looks like the `hiring_department_code` is null for some rows.

Let's count the extent of the issues.

```{r}
null_variable %>% 
  pull(variable) %>% 
  map(
    ~null_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's see how the hiring codes look in comparison to the department names.

```{r}
prep_blank %>% 
  count(hiring_department_code,hiring_department_name)
```

There are a lot of nulls, and 0s.

Check for any duplications in the codes.

```{r}
prep_blank %>% 
  count(hiring_department_code,hiring_department_name) %>% 
  get_dupes(hiring_department_code)
```

Lots of 0, nulls, and 5s.

The department name appears more complete. We will drop `hiring_department_code`.

```{r}
prep_xx <- 
  prep_blank %>% 
  select(-hiring_department_code)
```

Rename dataframe

```{r}
prep_null <- prep_xx
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing nulls produced by:

```{r}
null_variable <- 
  null_variables %>% 
  filter(group == null_groups[step_n])

null_variable
```

Reviewing blanks.

```{r}
null_01 <- 
  prep_blank %>% 
  filter(if_any(all_of(null_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(null_variable %>% pull(variable)),everything())

null_01
```

Looks like the `telework_eligibility_description` is null.

Let's count the extent of the issues.

```{r}
null_variable %>% 
  pull(variable) %>% 
  map(
    ~null_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

The whole column is null. We should drop.

```{r}
prep_xx <- 
  prep_null %>% 
  select(-telework_eligibility_description)
```

Rename dataframe

```{r}
prep_null <- prep_xx
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing nulls produced by:

```{r}
null_variable <- 
  null_variables %>% 
  filter(group == null_groups[step_n])

null_variable
```

Reviewing blanks.

```{r}
null_01 <- 
  prep_blank %>% 
  filter(if_any(all_of(null_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(null_variable %>% pull(variable)),everything())

null_01
```

Looks like the `second_announcement` is null.

Let's count the extent of the issues.

```{r}
null_variable %>% 
  pull(variable) %>% 
  map(
    ~null_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

The whole column is null. We should drop.

```{r}
prep_xx <- 
  prep_null %>% 
  select(-second_announcement)
```

Rename dataframe

```{r}
prep_null <- prep_xx
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing nulls produced by:

```{r}
null_variable <- 
  null_variables %>% 
  filter(group == null_groups[step_n])

null_variable
```

Reviewing blanks.

```{r}
null_01 <- 
  prep_blank %>% 
  filter(if_any(all_of(null_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(null_variable %>% pull(variable)),everything())

null_01
```

Looks like the `pay_scale` is null.

Let's count the extent of the issues.

```{r}
null_variable %>% 
  pull(variable) %>% 
  map(
    ~null_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's see how the `pay_scale` look in comparison to the `pay_scale_description`.

```{r}
prep_blank %>% 
  count(pay_scale,pay_scale_description)
```

Looks like the NAs in `pay_scale` are "Nonappropriated Funds--Nonsupervisory And Nonleader--Federal Wage System." We'll retain `pay_scale_description`and drop `pay_scale`.

```{r}
prep_xx <- 
  prep_null %>% 
  select(-pay_scale)
```

Rename dataframe

```{r}
prep_null <- prep_xx
```

```{r}
step_n <- step_n + 1
```

### Check that nulls are removed

Created a dataframe of variable metadata for blanks data

```{r}
column_meta_data_null <- 
  prep_null %>% 
  contents()

column_meta_data_null <- 
  column_meta_data_null$contents %>% 
  as.data.frame()

column_meta_data_null <-
  column_meta_data_null %>% 
  rownames_to_column(var = "variable")

column_meta_data_null
```

Checking duplicates for each variable in the dataframe.

```{r}
nulls_by_variable <- 
  column_meta_data_null %>% 
  pull(variable) %>% 
  map_lgl(~null_check(in_data = prep_null,.))

nulls_by_variable
```

Add this to the `column_meta_data_null` dataframe.

```{r}
column_meta_data_null <- 
  column_meta_data_null %>% 
  add_column(
    nulls_by_variable = nulls_by_variable
  )

column_meta_data_null
```

Create dataframe of only variables that have blanks.

```{r}
null_variables <- 
  column_meta_data_null %>% 
  filter(nulls_by_variable)

null_variables
```

## Other data cleaning

Review the data.

```{r}
prep_null %>% 
  skim()
```

Only two variables stand out: `telework_eligible` and `supervisory_status`.

First count them.

```{r}
prep_null %>% 
  count(telework_eligible,supervisory_status)
```

No positions appear to be telework eligible, which is wrong.

<!-- Talk to Melissa about this. -->

For supervisory status, it's preferable to transform this into an indicator.

```{r}
prep_xx <- 
  prep_null %>% 
  character_to_indicator(params$unique_identifier_1,supervisory_status,"supervisory_status") %>% 
  select(-supervisory_status_n)

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Rename dataframe.

```{r}
prep_12 <- prep_xx
```

## Creating final data set

Remove `supervisory_status` and attach `supervisory_status_y` the new data

```{r}
prep_xx <- 
  prep_null %>% 
  select(-supervisory_status) %>% 
  left_join(prep_12)

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Rename dataframe.

```{r}
joas <- prep_xx
```

## Save data {.tabset .tabset-pills}

```{r}
save(joas, 
     file = here::here("03. Data Collection",
                       "prepared-data",
                       glue::glue("joas-{today()}.rdata")))
```

## Extraction time {.tabset .tabset-pills}

```{r}
tictoc::toc()
```