### 2023-03-15 04. Analysis/scripts/01-data-preparation-joas.Rmd
Update 01-data-preparation-joas.Rmd
example from opm work updated

---
title: "Data preparation - JOAs"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: readable
    highlight: tango
    code_folding: hide
params:
  author: "Sam Kim and Ben Jaques-Leslie"
  project_number: "001"
  project_name: "Remote Work"
  data_folder_1: "03. Data Collection/Original data files from Melissa/JOAs"
  unique_identifier_1: !r c("usajobs_control_number")
  date_earliest_open: '2022-01-01'
  data_latest_close: '2022-12-31'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
tictoc::tic()
library(Hmisc)
library(tidyverse)
library(janitor)
library(skimr)
library(DataExplorer)
library(readr)
library(data.table)
library(lubridate)
library(arsenal)
```

# {.tabset .tabset-pills}

**Project number**: `r params$project_number`

**Project name**: `r params$project_name`

**Author**: `r params$author`

**Data folder 1**: `r params$data_folder`

**Unique identifier for data folder 1**: `r params$unique_identifier_1`

## Load data {.tabset .tabset-pills}

### Data files

The file names below represent all original candidate data files from Melissa.

```{r}
files <- 
  tibble(files = list.files(here::here(params$data_folder_1))
)

files
```

### Function to identify the encoding of the data files.

```{r}
encoding_data_frame <- function(in_file_name)
{
  readr::guess_encoding(here::here(params$data_folder_1,in_file_name)) %>% 
        mutate(file = in_file_name) %>% 
        select(file, everything())
}
```

### Identifying file encoding

Identify the encoding the candidates data files.

```{r}
files <- 
  files %>% 
  pull() %>% 
  map_dfr(~encoding_data_frame(.))

files
```

Filter candidate files to those with greatest confidence in the encoding.

```{r}
files <- 
  files %>% 
  group_by(file) %>% 
  filter(confidence == max(confidence))

files
```

### Functions

#### Function to load UTF-8 candidate files

```{r}
load_utf8 <- function(in_file_name, in_col_names)
{
  prep_01 <-
    fread(
      here::here(params$data_folder_1,in_file_name)
    ) %>% 
      distinct()

  return(prep_01)
}
```

#### Function to load UTF-16 candidate files

```{r}
load_utf16 <- function(in_file_name)
{
  prep_01 <- read.csv(
    file = here::here(params$data_folder_1,in_file_name),
    fileEncoding = "UTF-16",
    sep = "\t"
    )

  return(prep_01)
}
```

#### Function to prepare JOA info data.

```{r}
prepare_joa <- function(in_data)
{
prep_02 <- 
  in_data %>% 
  clean_names()

prep_03 <-
  prep_02 %>%
  select(everything())

if("announcement_closing_type" %in% names(prep_03))
{
  prep_04 <-
  prep_03 %>%
  mutate(
    across(
      .cols = c(position_open_date,position_close_date),
      .fns = ~ymd(.)
    )
  ) %>% 
  mutate(
    across(
      .cols = c(usajobs_control_number,announcement_closing_type,agency_level,hiring_department_code,series),
      .fns = ~as.character(.)
    )
  ) %>% 
  mutate(
    across(
      .cols = c(hiring_agency_code,total_openings),
      .fns = ~str_trim(.)
    )
  ) %>% 
  mutate(
    across(
      .cols = c(total_openings),
      .fns = ~as.numeric(.)
    )
  )
}

if(!("announcement_closing_type" %in% names(prep_03)) &
   !("agency_level" %in% names(prep_03)) &
   !("hiring_department_code" %in% names(prep_03)) &
   !("hiring_agency_code" %in% names(prep_03)) 
     )
{
  prep_04 <-
  prep_03 %>%
  mutate(
    across(
      .cols = c(position_open_date,position_close_date),
      .fns = ~ymd(.)
    )
  ) %>% 
  mutate(
    across(
      .cols = starts_with("usajobs_control_number")|
                # announcement_closing_type,
                # agency_level,
                # hiring_department_code,
                starts_with("series"),
      .fns = ~as.character(.)
    )
  ) %>% 
  mutate(
    across(
      .cols = c(
        # hiring_agency_code,
        total_openings),
      .fns = ~str_trim(.)
    )
  ) %>% 
  mutate(
    across(
      .cols = c(total_openings),
      .fns = ~as.numeric(.)
    )
  )
}

  #   position_open_date = ymd(position_open_date),
  #   position_close_date = ymd(position_close_date),
  #   usajobs_control_number = as.character(usajobs_control_number),
  #   race = str_remove(race, " - .*"),
  #   race = case_when(race == "" ~ "Did not disclose race",
  #                    is.na(race) ~ "Did not disclose race",
  #                    TRUE ~ race),
  #   degree_level = str_remove(degree_level, " \\(.*"),
  #   degree_level = case_when(degree_level == "" ~ "Did not disclose degree",
  #                    TRUE ~ degree_level),
  #   federal_employment_status =
  #     case_when(
  #       str_detect(
  #         federal_employment_status,"former") ~ "Former",
  #       str_detect(
  #         federal_employment_status,"currently") ~ "Current",
  #       str_detect(
  #         federal_employment_status,"never") ~ "Non-Federal employee"
  #       )
  # )
# 
# prep_05 <- 
#   prep_04 %>% 
#   select(candidate_id,race) %>% 
#   distinct() %>%
#   mutate(ind = 1) %>% 
#   pivot_wider(names_from = race, values_from = ind) %>% 
#   clean_names() %>% 
#   mutate(
#     across(
#       .cols = where(is.numeric),
#       .fns = ~replace_na(.,0)
#     )
#   )
# 
# prep_06 <-
#   prep_04 %>% 
#   select(candidate_id,degree_level) %>% 
#   distinct() %>% 
#   mutate(ind = 1) %>% 
#   pivot_wider(names_from = degree_level, values_from = ind) %>% 
#   clean_names() %>% 
#   mutate(
#     across(
#       .cols = where(is.numeric),
#       .fns = ~replace_na(.,0)
#     )
#   )
# 
# prep_07 <- 
#   prep_04 %>% 
#   select(-race,-degree_level) %>% 
#   distinct()
# 
# prep_08 <- 
#   prep_05 %>% 
#   left_join(prep_06) %>% 
#   left_join(prep_07)

return(prep_04)
}
```

#### Function to prepare JOA info data.

```{r}
prepare_joa_telework <- function(in_data)
{
prep_02 <- 
  in_data %>% 
  clean_names()

prep_03 <-
  prep_02 %>%
  select(everything())

if("is_tele_work_eligible" %in% names(prep_03))
{
  prep_04 <-
  prep_03 %>%
    rename(telework_eligible = is_tele_work_eligible)
}

if(!("is_tele_work_eligible" %in% names(prep_03)))
{
  prep_04 <-
  prep_03
}

prep_05 <- 
  prep_04 %>% 
  select(usajobs_control_number,telework_eligible) %>% 
  mutate(
    across(
      .cols = c(usajobs_control_number,
               ),
      .fns = ~as.character(.)
    )
  ) %>%
  mutate(
    across(
      .cols = c(telework_eligible),
      .fns = ~as.numeric(.)
    )
  )
# 
# if(!("announcement_closing_type" %in% names(prep_03)) &
#    !("agency_level" %in% names(prep_03)) &
#    !("hiring_department_code" %in% names(prep_03)) &
#    !("hiring_agency_code" %in% names(prep_03)) 
#      )
# {
#   prep_04 <-
#   prep_03 %>%
#   mutate(
#     across(
#       .cols = c(position_open_date,position_close_date),
#       .fns = ~ymd(.)
#     )
#   ) %>% 
#   mutate(
#     across(
#       .cols = c(usajobs_control_number,
#                 # announcement_closing_type,
#                 # agency_level,
#                 # hiring_department_code,
#                 series),
#       .fns = ~as.character(.)
#     )
#   ) %>% 
#   mutate(
#     across(
#       .cols = c(
#         # hiring_agency_code,
#         total_openings),
#       .fns = ~str_trim(.)
#     )
#   ) %>% 
#   mutate(
#     across(
#       .cols = c(total_openings),
#       .fns = ~as.numeric(.)
#     )
#   )
# }

  #   position_open_date = ymd(position_open_date),
  #   position_close_date = ymd(position_close_date),
  #   usajobs_control_number = as.character(usajobs_control_number),
  #   race = str_remove(race, " - .*"),
  #   race = case_when(race == "" ~ "Did not disclose race",
  #                    is.na(race) ~ "Did not disclose race",
  #                    TRUE ~ race),
  #   degree_level = str_remove(degree_level, " \\(.*"),
  #   degree_level = case_when(degree_level == "" ~ "Did not disclose degree",
  #                    TRUE ~ degree_level),
  #   federal_employment_status =
  #     case_when(
  #       str_detect(
  #         federal_employment_status,"former") ~ "Former",
  #       str_detect(
  #         federal_employment_status,"currently") ~ "Current",
  #       str_detect(
  #         federal_employment_status,"never") ~ "Non-Federal employee"
  #       )
  # )
# 
# prep_05 <- 
#   prep_04 %>% 
#   select(candidate_id,race) %>% 
#   distinct() %>%
#   mutate(ind = 1) %>% 
#   pivot_wider(names_from = race, values_from = ind) %>% 
#   clean_names() %>% 
#   mutate(
#     across(
#       .cols = where(is.numeric),
#       .fns = ~replace_na(.,0)
#     )
#   )
# 
# prep_06 <-
#   prep_04 %>% 
#   select(candidate_id,degree_level) %>% 
#   distinct() %>% 
#   mutate(ind = 1) %>% 
#   pivot_wider(names_from = degree_level, values_from = ind) %>% 
#   clean_names() %>% 
#   mutate(
#     across(
#       .cols = where(is.numeric),
#       .fns = ~replace_na(.,0)
#     )
#   )
# 
# prep_07 <- 
#   prep_04 %>% 
#   select(-race,-degree_level) %>% 
#   distinct()
# 
# prep_08 <- 
#   prep_05 %>% 
#   left_join(prep_06) %>% 
#   left_join(prep_07)

return(prep_05)
}
```

#### Function to load and prepare JOA information

```{r}
load_prepare_joa <- function(in_file_name,in_encoding)
{
  if(in_encoding == 'ASCII')
  {
    prep_01 <- in_file_name %>% 
  load_utf8()
  }

  if(in_encoding == 'UTF-16LE')
  {
    prep_01 <- in_file_name %>% 
  load_utf16()
  }

  out <- prep_01 %>% 
  prepare_joa()

  return(out)

}
```

#### Function to load and prepare JOA telework information

```{r}
load_prepare_joa_telework <- function(in_file_name,in_encoding)
{
  if(in_encoding == 'ASCII')
  {
    prep_01 <- in_file_name %>% 
  load_utf8()
  }

  if(in_encoding == 'UTF-16LE')
  {
    prep_01 <- in_file_name %>% 
  load_utf16()
  }

  out <- prep_01 %>% 
  prepare_joa_telework()

  return(out)

}
```

#### Function to convert character to indicator

```{r}
character_to_indicator <- function(in_data,in_unique_identifier,in_variable,in_prefix)
{
  in_data %>% 
  select(all_of(in_unique_identifier),{{in_variable}}) %>% 
    distinct() %>% 
    mutate(ind = 1) %>% 
    mutate(across(
      .cols = {{in_variable}},
      .fns = ~case_when(. == "" ~ "missing",
                       TRUE ~ .)
    )) %>% 
    pivot_wider(names_from = {{in_variable}},
                values_from = ind) %>% 
    rename_with(
      .fn = ~glue::glue("{in_prefix} {.}"),
  .cols = -in_unique_identifier
  ) %>% 
    clean_names() %>% 
    mutate(
      across(
        .cols = where(is.numeric),
        .fns = ~replace_na(.,0)
      )
    )
}
```

### JOA data

#### Identify files with "full" in title or "Positions_Updated_Dec2022"

```{r}
files_full <- 
  files %>% 
  filter(str_detect(file,"_full_") | 
           str_detect(file,"Positions_Updated_Dec2022") 
         ) %>%
  select(file,encoding)

files_full
```

#### Load 'allpositionsopenedduring2022'

Load position headers for the data.

```{r}
positions_headers <- 
  here::here(params$data_folder_1,"positions_headers.csv") %>% 
  read_csv() %>% 
      distinct()
```

Load the data and apply the position headers.

```{r}
prep_all_positions <- 
  files %>% 
  filter(
    str_detect(file,"allpositionsopenedduring2022")
         ) %>%
  pull(file) %>%  
  load_utf8(
  ) %>% 
  rename_all(~names(positions_headers)) %>% 
  clean_names() %>% 
  prepare_joa()
```

#### Load 'full' files

```{r}
prep_base <- 
  map2_dfr(
    .x = files_full$file,
    .y = files_full$encoding,
    .f = ~load_prepare_joa(in_file_name = .x, in_encoding = .y)
  ) %>% 
  bind_rows(prep_all_positions) %>% 
  distinct()
```

#### Identify files with "remote_telework_positions" in title or "telework_eligible_from_09092022"

```{r}
files_telework <- 
  files %>% 
  filter(str_detect(file,"remote_telework_positions") | str_detect(file,"telework_eligible_from_09092022")) %>%
  select(file,encoding)

files_telework
```

#### Load 'telework' files

```{r}
prep_telework <- 
  map2_dfr(
    .x = files_telework$file,
    .y = files_telework$encoding,
    .f = ~load_prepare_joa_telework(in_file_name = .x, in_encoding = .y)
  ) %>% 
  distinct()
```

```{r}
prep_telework %>% 
  skim()
```

#### Load 'hiring path' files

Load first hiring path file.

```{r}
prep_hiring_path_01 <- 
   read_csv(here::here(params$data_folder_1,"hiringpath_remotework.csv"),
   col_types = cols(`Hiring Path` = col_character(), 
        `Position Open Date` = col_skip(), 
        `Position Close Date` = col_skip(), 
        `USAJOBS Control Number` = col_character())
   ) %>% 
  clean_names() %>% 
  distinct()
```

Collect the names from the dataframe.

```{r}
prep_hiring_path_01_names <- sort(names(prep_hiring_path_01), decreasing = TRUE)
```

Load second hiring path file, and apply names from the first dataset.

```{r}
prep_hiring_path_02 <- read_csv(here::here(params$data_folder_1,"positions - hiring path - 2022.csv"),
    col_names = FALSE, col_types = cols(X1 = col_character(), 
        X2 = col_skip(), X3 = col_skip(), 
        X4 = col_character())) %>% 
  distinct() %>% 
  rename_at(
    vars(everything()),
    ~prep_hiring_path_01_names
  )
```

Bind dataframes and convert hiring path to indicators.

```{r}
prep_hiring_path <- 
  prep_hiring_path_01 %>% 
  bind_rows(prep_hiring_path_02) %>% 
  distinct() %>% 
  character_to_indicator(in_unique_identifier = params$unique_identifier_1,
                         in_variable = hiring_path,
                         in_prefix = "hiring_path")

```

```{r}
prep_hiring_path %>% 
  skim()
```

#### Load 'work schedule and appointment type' files

Load first work schedule and appointment type file.

```{r}
prep_work_schedule_appointment_01 <- 
   read_csv(here::here(params$data_folder_1,"positions - work schedule and appointment type.csv"), 
    col_types = cols(Announcement_Number = col_skip(), 
        USAJOBS_Control_Number = col_character())
   ) %>% 
  clean_names() %>% 
  distinct()
```

Convert work schedule to indicators.

```{r}
prep_work_schedule <- 
  prep_work_schedule_appointment_01 %>% 
  distinct() %>% 
  character_to_indicator(in_unique_identifier = params$unique_identifier_1,
                         in_variable = position_work_schedule_type,
                         in_prefix = "work_schedule")

```

Convert appointment type to indicators.

```{r}
prep_appointment <- 
  prep_work_schedule_appointment_01 %>% 
  distinct() %>% 
  character_to_indicator(in_unique_identifier = params$unique_identifier_1,
                         in_variable = position_appointment_type,
                         in_prefix = "appointment_type")

```

Join indicator dataframes.

```{r}
prep_work_schedule_appointment <- 
  prep_work_schedule %>% 
  full_join(prep_appointment)
```

```{r}
prep_work_schedule_appointment %>% 
  skim()
```

#### Load 'series' files

Load first series file.

```{r}
prep_series_01 <- 
   read_csv(here::here(params$data_folder_1,"positions - series - 2022.csv"), 
    col_types = cols(USAJOBS_Control_Number = col_character(), 
        Announcement_Number = col_skip(), 
        Series = col_character(), Series_Title = col_character())) %>% 
  clean_names() %>% 
  distinct()
```

Rename.

```{r}
prep_series <- 
  prep_series_01
```

```{r}
prep_series %>% 
  skim()
```

#### Load 'locations' files

Load first locations file.

```{r}
prep_locations_01 <- 
   read_csv(here::here(params$data_folder_1,"positions - locations - 2022.csv"), 
   col_types = cols(`Announcement Number` = col_skip(), 
        `USAJOBS Control Number` = col_character(), 
        `Position Open Date` = col_date(format = "%m/%d/%Y"), 
        `Position Close Date` = col_date(format = "%m/%d/%Y"), 
        `Position Location City` = col_character(), 
        `Position Location Country` = col_character(), 
        `Position Location Geo-location Code` = col_character(), 
        `Position Location State` = col_character(), 
        `Position Location Zip` = col_character())) %>% 
  clean_names() %>% 
  distinct()
```

Rename.

```{r}
prep_location <- 
  prep_locations_01
```

```{r}
prep_location %>% 
  skim()
```

Check for duplicates.

```{r}
dupes_location <- 
  prep_location %>% 
  get_dupes(params$unique_identifier_1)

dupes_location
```

Count by number of duplications.

```{r}
dupes_location %>% 
  count(dupe_count)
```
Count number of `position_location_geo_location_code`.

```{r}
location_count <- 
  prep_location %>% 
  group_by(position_location_city,position_location_country,
           position_location_geo_location_code,
           position_location_state,position_location_zip) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  rownames_to_column(var = 'rank')

location_count
```

Check to see if any of these 'NULL' are duplicated.

```{r}
dupes_location %>% 
  filter(position_location_geo_location_code == "NULL")
```

Aggregate location data to deduplicate, counting multiple locations.

```{r}
prep_location_n_distinct <- 
  prep_location %>%
  group_by(usajobs_control_number
           ,position_close_date, 
           position_open_date
           ) %>%
  summarise(
    across(
      .cols = everything(),
      .fns = ~n_distinct(.),
      .names = "{.col}_n_distinct"
      )
  ) %>% 
  ungroup() %>% 
  mutate(
    multiple_location_ind = 
      case_when(if_any(ends_with("_n_distinct"), ~ .x > 1) ~ 1,
                TRUE ~ 0
                ),
  ) %>% 
  arrange(position_location_geo_location_code_n_distinct)

prep_location_n_distinct %>% 
  skim()
```

Aggregate location data to deduplicate, assigning locations by least and most common.

```{r}
prep_location_common <-
  prep_location %>%
  left_join(location_count) %>% 
  group_by(usajobs_control_number
           ,position_close_date, 
           position_open_date
           ) %>%
  arrange(usajobs_control_number
           ,position_close_date, 
           position_open_date,
          rank) %>% 
  summarise(
    across(
      .cols = everything(),
      .fns = list(most_common = first, least_common = last),
      .names = "{.col}_{.fn}"
      )
  ) %>% 
  ungroup()

prep_location_common %>% 
  skim()
```

# CONTINUE FROM HERE

## Filter to correct dates

```{r}
prep_base <- 
  prep_base %>% 
  filter(position_open_date >= ymd(params$date_earliest_open) &
           position_close_date <= ymd(params$data_latest_close))

```

#### Review data

```{r}
prep_base %>%
  skim()
```

## Addressing duplication

Melissa recommended using `r params$unique_identifier_1` as the unique identifier.

To review specific cases you to https://www.usajobs.gov/job/ and put the USAJobs control number at the end.

### Check for duplication

```{r}
dupe_01 <- 
  prep_base %>%
  get_dupes(params$unique_identifier_1)
```

#### Data review

```{r}
dupe_01 %>% 
  skim()
```

### Dataframe of the unique identifier

```{r}
prep_uniq <- 
  prep_base %>% 
  select(params$unique_identifier_1) %>% 
  distinct()

prep_uniq %>% 
  count()
```

Also create a building dataframe to add other deduplicated dataframes.

```{r}
prep_building <- prep_uniq
```

### Identifying and categorizing columns

Created a dataframe of variable metadata

```{r}
column_meta_data <- 
  prep_base %>% 
  contents()

column_meta_data <- 
  column_meta_data$contents %>% 
  as.data.frame()

column_meta_data <-
  column_meta_data %>% 
  rownames_to_column(var = "variable")

column_meta_data
```

Function that checks whether there are duplicates for different variables

```{r}
dupe_check <- function(in_data,in_variable_name)
{
  check_01 <- 
  in_data %>% 
  select(params$unique_identifier_1,all_of(in_variable_name)) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

out <- nrow(check_01) > 0

return(out)
}
```

Checking duplicates for each variable in the dataframe.

```{r}
duplicates_by_variable <- 
  column_meta_data %>% 
  pull(variable) %>% 
  map_lgl(~dupe_check(in_data = dupe_01, .))

duplicates_by_variable
```

Add this to the column_meta_data dataframe.

```{r}
column_meta_data <- 
  column_meta_data %>% 
  add_column(
    duplicates_by_variable = duplicates_by_variable
  )

column_meta_data
```

Create dataframe of only variables that cause duplicates.

```{r}
duplicating_variables <- 
  column_meta_data %>% 
  filter(duplicates_by_variable)

duplicating_variables
```

Grouping variables.

```{r}
duplicating_variables <- 
  duplicating_variables %>% 
  mutate(group = 
           case_when(variable == "position_open_date" ~ "dates",
                     variable == "position_close_date" ~ "dates",
                     variable == "announcement_closing_type" ~ "announcement_closing_type",
                     variable == "announcement_closing_type_description" ~ "announcement_closing_type",
                     variable == "announcement_closing_type_option" ~ "announcement_closing_type",
                     variable == "application_limit" ~ "application_limit",
                     variable == "application_limit_set" ~ "application_limit",
                     variable == "application_limit_number" ~ "application_limit",
                     variable == "agency_level" ~ "agency_level",
                     variable == "agency_level_sort" ~ "agency_level",
                     variable == "hiring_agency_code" ~ "agency_level",
                     variable == "hiring_agency_name" ~ "agency_level",
                     variable == "hiring_department_code" ~ "agency_level",
                     variable == "hiring_department_name" ~ "agency_level",
                     variable == "maximum_salary" ~ "maximum_salary",
                     variable == "minimum_salary" ~ "maximum_salary",
                     variable == "maximum_grade" ~ "maximum_salary",
                     variable == "minimum_grade" ~ "maximum_salary",
                     variable == "series" ~ "series",
                     variable == "series_title" ~ "series",
                     variable == "mission_critical_code" ~ "mission_critical_code",
                     variable == "mission_critical_code_description" ~ "mission_critical_code",
                     variable == "mission_critical_source" ~ "mission_critical_code",
                     TRUE ~ variable
                     ))

duplicating_variables
```

Create list of groups.

```{r}
duplicating_groups <- 
  duplicating_variables %>% 
  select(group) %>% 
  distinct() %>% 
  arrange(group) %>% 
  pull()

duplicating_groups
```

### Duplication removal

```{r}
step_n <- 0
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `agency_level`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

A few things going on here:

1.    Department of Defense agencies are duplicated because sometimes sub-departments (e.i. Department of Navy) are considered the *Department*
2.    Many duplicates are caused by blank values in hiring agency code and agency level.

First let's remove duplicates because of department. To do this, we will group by control number, then arrange by and control number and descending agency level. Then, we take the data in the first entry.

```{r}
dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  group_by_at(params$unique_identifier_1) %>% 
  arrange(usajobs_control_number,desc(agency_level)) %>% 
  summarise(
    across(
      .cols = everything(),
      .fns = first
    )
  ) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)
```

Now we will apply all of these to the whole dataframe.

```{r}
dedupe <- 
  prep_base %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  group_by_at(params$unique_identifier_1) %>% 
  arrange(usajobs_control_number,desc(agency_level)) %>% 
  summarise(
    across(
      .cols = everything(),
      .fns = first
    )
  ) %>% 
  distinct() %>% 
  ungroup()

dedupe %>% 
  get_dupes(params$unique_identifier_1)
```

Let's review this data set.

```{r}
prep_xx <- 
  dedupe

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Join to building dataframe.

```{r}
prep_building <- prep_building %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `announcement_closing_type`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Check to see if the type codes are consistent to the description.

```{r}
prep_base %>% 
  count(announcement_closing_type,announcement_closing_type_description)
```

Codes are consistent to description, but the code adds no new information. We will drop the code and transform the description into columns of indicators.

We want to have this change to affect the whole dataframe so we address the duplicates off the original dataframe. We will update the NA column to include announcements where there are no other indicators equal to one except NA.

```{r}
prep_xx <-
  prep_base %>% 
  character_to_indicator(params$unique_identifier_1,announcement_closing_type_description,"closing_type") %>% 
  mutate(
    closing_type_na = case_when(
      closing_type_closing_date == 1 ~ 0,
      closing_type_applicant_cut_off == 1 ~ 0,
      closing_type_open_continuous == 1 ~ 0,
      closing_type_missing == 1 ~ 0,
      closing_type_closing_time == 1 ~ 0,
      closing_type_na == 1 ~ 1,
    )
  )

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Rename dataframe and count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Join to building dataframe.

```{r}
prep_building <- prep_building %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `announcement_number`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)
```

Some of these duplicates may be due to small typos. Transforming the string to upper case.

```{r}
dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  mutate(
    announcement_number = str_to_upper(announcement_number)
  ) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)
```

This fixes some, but not all.

Without better information, we will take the lower of the two values for the remainder.

Apply this to the entire dataframe.

```{r}
dedupe <- 
  prep_base %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  mutate(
    announcement_number = str_to_upper(announcement_number)
  ) %>% 
  distinct() %>% 
  arrange(params$unique_identifier_1,announcement_number) %>% 
  group_by_at(params$unique_identifier_1) %>% 
  summarise(announcement_number = first(announcement_number)) %>% 
  ungroup()

dedupe %>% 
  get_dupes(params$unique_identifier_1)
```

Review the dataframe.

```{r}
prep_xx <- 
  dedupe

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Rename dataframe and count rows.

```{r}
prep_01 <- prep_xx

prep_01 %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_01 %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Join to building dataframe.

```{r}
prep_building <- prep_building %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `application_limit`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Not too many. It appears that `application_limit_set` is "N" when `application_limit` is "Applicant Cut-Off". Also `application_limit_number` appears to only be valid when `application_limit_set` is NA.

First count the statuses across the original data.

```{r}
prep_base %>% count(application_limit,application_limit_set)
```

Now let's look at `application_limit_number`.

```{r}
prep_base %>% 
  filter(!is.na(application_limit_number)) %>% 
  select(all_of(params$unique_identifier_1),application_limit_number) %>% 
  distinct()
```

Only a very few rows have values for `application_limit_number`. We will drop this column.

All NA `application_limit` are "N" `application_limit_set`. Also NA in `application_limit_set` have values in `application_limit`. We will apply the `application_limit` values to NAs in `application_limit_set`.

We will consider "Y" the correct response for all and we will change this into an indicator.

We want to have this change to affect the whole dataframe so we address the duplicates off the original dataframe.

```{r}
prep_xx <- 
  prep_base %>% 
  mutate(
    application_limit_set = case_when(is.na(application_limit_set) ~ application_limit,
                                      TRUE ~ application_limit_set)
  ) %>% 
  character_to_indicator(params$unique_identifier_1,application_limit_set,"application_limit_set") %>% 
  # Drop the "N" column
  select(-application_limit_set_n)

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Join to building dataframe.

```{r}
prep_building <- prep_building %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `appointment_type``

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Lots of them, but appears to be mainly because of NAs.

First count the statuses across the original data.

```{r}
prep_base %>% count(appointment_type)
```

We'll want to transform these into indicators and transform the NAs so that they are only for *true* NAs.

We want to have this change to affect the whole dataframe so we address the duplicates off the original dataframe.

```{r}
prep_xx <- 
  prep_base %>% 
  character_to_indicator(params$unique_identifier_1,appointment_type,"appointment_type") %>% 
  mutate(
    appointment_type_na = case_when(
      appointment_type_permanent == 1 ~ 0,
      appointment_type_term == 1 ~ 0,
      appointment_type_intermittent == 1 ~ 0,
      appointment_type_temporary == 1 ~ 0,
      appointment_type_internships == 1 ~ 0,
      appointment_type_multiple_appointment_types == 1 ~ 0,
      appointment_type_temporary_promotion == 1 ~ 0,
      appointment_type_recent_graduates == 1 ~ 0,
      appointment_type_detail == 1 ~ 0,
      appointment_type_seasonal == 1 ~ 0,
      appointment_type_agency_employees_only == 1 ~ 0,
      appointment_type_summer == 1 ~ 0,
      appointment_type_ictap_only == 1 ~ 0,
      appointment_type_presidential_management_fellows == 1 ~ 0,
      appointment_type_telework == 1 ~ 0,
      appointment_type_na == 1 ~ 1
    )
  )

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Join to building dataframe.

```{r}
prep_building <- prep_building %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `dates`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Duplication appear to be because of difference in open and close dates. After messaging with Melissa, we decide to retain the dates as determined by the close date.

We want to have this change to affect the whole dataframe so we address the duplicates off the original dataframe.

```{r}
prep_xx <- 
  prep_base %>% 
   select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>%
  arrange(params$unique_identifier_1, position_close_date) %>% 
  group_by(usajobs_control_number) %>% 
  summarise(
    across(
      .cols = all_of(duplicating_group %>% pull(variable)),
      .fns = last
    )
  )

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Join to building dataframe.

```{r}
prep_building <- prep_building %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `disable_apply_online`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Quite a few. It appears that `disable_apply_online` in NA in many duplicates.

First count the statuses across the original data.

```{r}
prep_base %>% count(disable_apply_online)
```

There are just a few, but we'll want to transform these into indicators.

We want to have this change to affect the whole dataframe so we address the duplicates off the original dataframe.

Since there is only one instance of this type of duplication, we will still drop the no column.

```{r}
prep_xx <- 
  prep_base %>% 
  character_to_indicator(params$unique_identifier_1,disable_apply_online,"disable_apply_online") %>% 
  mutate(
    disable_apply_online_na = case_when(
      disable_apply_online_n == 1 ~ 0,
      disable_apply_online_y == 1 ~ 0,
      disable_apply_online_na == 1 ~ 1
    )
  )

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Join to building dataframe.

```{r}
prep_building <- prep_building %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `hiring_subelement_name`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Count types.

```{r}
prep_base %>% 
  count(hiring_subelement_name)
```

Many blank entries and a level of detail we don't need. We will drop this column.

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `is_remote_work_eligible`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Duplication appear to be because `r duplicating_group %>% pull(variable)` is empty in one row. Transform the empties into 'N'.

We want to have this change to affect the whole dataframe so we address the duplicates off the original dataframe.

```{r}
prep_xx <- 
  prep_base %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  mutate(
    is_remote_work_eligible = case_when(is_remote_work_eligible == "" ~ "N",
                                        is_remote_work_eligible == "NULL" ~ "N",
                                        TRUE ~ is_remote_work_eligible)
  ) %>% 
  distinct() %>% 
  character_to_indicator(in_unique_identifier = params$unique_identifier_1,
                         in_variable = is_remote_work_eligible,
                         in_prefix = "is_remote_work_eligible") %>% 
  # Drop the "n" indicatoris_remote_work_eligible
  select(-is_remote_work_eligible_n)

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Join to building dataframe.

```{r}
prep_building <- prep_building %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `maximum_salary`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Appears like this comes from posting the job again with revised salary information.

Like above we decide to retain the later dates as determined by the close date.

```{r}
prep_xx <- 
  prep_base %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable)),position_close_date) %>%
  arrange(params$unique_identifier_1, position_close_date) %>% 
  group_by(usajobs_control_number) %>% 
  summarise(
    across(
      .cols = all_of(duplicating_group %>% pull(variable)),
      .fns = last
    )
  )

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Join to building dataframe.

```{r}
prep_building <- prep_building %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `mission_critical_code`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Let's count the entries here.

```{r}
prep_base %>% 
  group_by(across(all_of(duplicating_group %>% pull(variable)))) %>% 
  count()
```

Lots of entries here. So we don't want to make indicators. Looks like they are created by rows with a null. We'll count how many nulls and entries with values.

```{r}
dupe_02 %>% 
  mutate(cause = case_when(is.na(mission_critical_code) ~ 'null',
                           TRUE ~ 'value'),
         ind = 1) %>% 
  select(usajobs_control_number,cause,ind) %>% 
  distinct() %>% 
  pivot_wider(names_from = cause, values_from = ind) %>% 
  summarise(
    unique = n_distinct(usajobs_control_number),
    nulls = sum(null),
    values = sum(value, na.rm = TRUE)
  )
```

Select the rows that have an entry if they have that.

```{r}
prep_xx <- 
  prep_base %>% 
    select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>%
  arrange(usajobs_control_number,mission_critical_code) %>% 
  group_by(usajobs_control_number) %>% 
  summarise(
    across(
      .fns = first
    )
  ) %>% 
  ungroup()

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Join to building dataframe.

```{r}
prep_building <- prep_building %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `number_of_views`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Let's count the entries here.

```{r}
prep_base %>% 
  group_by(across(all_of(duplicating_group %>% pull(variable)))) %>% 
  count()
```

This looks pretty unreliable. We dill drop `number_of_views`.

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `position_opening_status`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Let's count the entries.

```{r}
prep_base %>% 
  group_by(across(all_of(duplicating_group %>% pull(variable)))) %>% 
  count()
```

We want to have this change to affect the whole dataframe so we address the duplicates off the original dataframe. We will update the NA column to include announcements where there are no other indicators equal to one except NA.

```{r}
prep_xx <-
  prep_base %>% 
  character_to_indicator(params$unique_identifier_1,
                         position_opening_status,
                         "position_opening_status") %>%
  mutate(
    position_opening_status_na = case_when(
      position_opening_status_job_canceled == 1 ~ 0,
      position_opening_status_applications_under_review == 1 ~ 0,
      position_opening_status_candidate_selected == 1 ~ 0,
      position_opening_status_accepting_applications == 1 ~ 0,
      position_opening_status_job_closed == 1 ~ 0,
      position_opening_status_missing == 1 ~ 0,
      position_opening_status_na	== 1 ~ 1,
    )
  )

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Join to building dataframe.

```{r}
prep_building <- prep_building %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `position_title`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Let's count the entries here.

```{r}
prep_base %>% 
  group_by(across(all_of(duplicating_group %>% pull(variable)))) %>% 
  count()
```

Lots of options. We'll try to use the ones associated with later closing dates.

```{r}
prep_xx <- 
  prep_base %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable)),position_close_date) %>%
  arrange(params$unique_identifier_1, position_close_date) %>% 
  group_by(usajobs_control_number) %>% 
  summarise(
    across(
      .cols = all_of(duplicating_group %>% pull(variable)),
      .fns = last
    )
  )

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Join to building dataframe.

```{r}
prep_building <- prep_building %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `series`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Let's count the entries here.

```{r}
prep_base %>% 
  group_by(across(all_of(duplicating_group %>% pull(variable)))) %>% 
  count()
```

There are more series types than we want to make indicators for.

Let's reviewing duplicates, but include closing date.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable)),position_close_date) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1,position_close_date)

dupe_02
```

Duplication also exists for `position_close_date`. If an announcement has multiple series on a the same closing date, we'll define the announcement as "(Multiple)". For others, we'll use close date to select the correct value.

```{r}
prep_xx_01 <- 
  prep_base %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable)),position_close_date) %>%
  group_by(across(all_of(params$unique_identifier_1)),position_close_date) %>% 
  summarise(
    across(
      .cols = all_of(duplicating_group %>% pull(variable)),
      .fns = ~n_distinct(.),
      .names = "{.col}_n_distinct"
    )
  ) %>% 
  ungroup()

prep_xx_02 <- 
  prep_base %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable)),position_close_date) %>%
  left_join(prep_xx_01) %>% 
  mutate(
    series = case_when(series_n_distinct > 1 ~ "(Multiple)",
                       TRUE ~ series),
    series_title = case_when(series_title_n_distinct > 1 ~ "(Multiple)",
                       TRUE ~ series_title)
  )

prep_xx <-
  prep_xx_02 %>% 
  arrange(params$unique_identifier_1, position_close_date) %>% 
  group_by(across(all_of(params$unique_identifier_1))) %>% 
  summarise(
    across(
      .cols = all_of(c(duplicating_group %>% pull(variable),"series_n_distinct","series_title_n_distinct")),
      .fns = last
    )
  )

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Join to building dataframe.

```{r}
prep_building <- prep_building %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `supervisory_status`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Let's count the entries here.

```{r}
prep_base %>% 
  group_by(across(all_of(duplicating_group %>% pull(variable)))) %>% 
  count()
```

Only two options, but they don't make sense as indicators. We'll use close date to select the correct value.

```{r}
prep_xx <- 
  prep_base %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable)),position_close_date) %>%
  arrange(params$unique_identifier_1, position_close_date) %>% 
  group_by(usajobs_control_number) %>% 
  summarise(
    across(
      .cols = all_of(duplicating_group %>% pull(variable)),
      .fns = last
    )
  )

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Join to building dataframe.

```{r}
prep_building <- prep_building %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `telework_eligibility_description`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Let's count the entries here.

```{r}
prep_base %>% 
  group_by(across(all_of(duplicating_group %>% pull(variable)))) %>% 
  count()
```

No useful information here. We will drop this column.

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `total_openings`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Let's count the entries here.

```{r}
prep_base %>% 
  group_by(across(all_of(duplicating_group %>% pull(variable)))) %>% 
  count()
```

We'll take the entries associated with the later closing date.

```{r}
prep_xx <- 
  prep_base %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable)),position_close_date) %>%
  arrange(params$unique_identifier_1, position_close_date) %>% 
  group_by(usajobs_control_number) %>% 
  summarise(
    across(
      .cols = all_of(duplicating_group %>% pull(variable)),
      .fns = last
    )
  )

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Join to building dataframe.

```{r}
prep_building <- prep_building %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `who_may_apply`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Let's count the entries here.

```{r}
prep_base %>% 
  group_by(across(all_of(duplicating_group %>% pull(variable)))) %>% 
  count()
```

Only a few categories. We will transform these into indicators.

```{r}
prep_xx <- 
  prep_base %>% 
  character_to_indicator(params$unique_identifier_1,who_may_apply,"who_may_apply") %>% 
  mutate(
    who_may_apply_missing = case_when(
      who_may_apply_united_states_citizens == 1 ~ 0,
      who_may_apply_agency_employees_only == 1 ~ 0,
      who_may_apply_status_candidates_merit_promotion_and_veoa_eligibles == 1 ~ 0,
      who_may_apply_missing == 1 ~ 1,
      who_may_apply_null == 1 ~ 1,
    )
  ) %>% 
  select(-who_may_apply_null)

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Join to building dataframe.

```{r}
prep_building <- prep_building %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `work_schedule`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_02 <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

dupe_02
```

Let's count the entries here.

```{r}
prep_base %>% 
  group_by(across(all_of(duplicating_group %>% pull(variable)))) %>% 
  count()
```

Only a few categories. We will transform these into indicators. Many NAs.

```{r}
prep_xx <- 
  prep_base %>% 
  character_to_indicator(params$unique_identifier_1,work_schedule,"work_schedule") %>%
  mutate(
    work_schedule_na = case_when(
      work_schedule_full_time == 1 ~ 0,
      work_schedule_part_time == 1 ~ 0,
      work_schedule_intermittent == 1 ~ 0,
      work_schedule_multiple_schedules == 1 ~ 0,
      work_schedule_shift_work == 1 ~ 0,
      work_schedule_job_sharing == 1 ~ 0,
      work_schedule_na == 1 ~ 1
    )
  )

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Join to building dataframe.

```{r}
prep_building <- prep_building %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

All done with duplicates!

### Combinding deduplicated data

#### Create dataframe off non-duplicating columns

Collect column names of variables not creating duplicates.

```{r}
non_duplicating_variables <- 
  column_meta_data %>% 
  filter(!duplicates_by_variable) %>% 
  pull(variable)

non_duplicating_variables
```

Create dataframe of only these variables.

```{r}
prep_xx <- 
  prep_base %>% 
  select(any_of(non_duplicating_variables)) %>% 
  distinct()

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Join to building dataframe.

```{r}
prep_building <- prep_building %>% 
  left_join(prep_xx)
```

### Checking duplicating columns have been fixed

Created a dataframe of variable metadata.

```{r}
column_meta_data_dedupe <- 
  prep_building %>% 
  contents()

column_meta_data_dedupe <- 
  column_meta_data_dedupe$contents %>% 
  as.data.frame()

column_meta_data_dedupe <-
  column_meta_data_dedupe %>% 
  rownames_to_column(var = "variable")

column_meta_data_dedupe
```

Checking duplicates for each variable in the dataframe.

```{r}
duplicates_by_variable <- 
  column_meta_data_dedupe %>% 
  pull(variable) %>% 
  map_lgl(~dupe_check(in_data = prep_building, .))

duplicates_by_variable
```

Add this to the column_meta_data dataframe.

```{r}
column_meta_data_dedupe <- 
  column_meta_data_dedupe %>% 
  add_column(
    duplicates_by_variable = duplicates_by_variable
  )

column_meta_data_dedupe
```

Create dataframe of only variables that cause duplicates.

```{r}
duplicating_variables <- 
  column_meta_data_dedupe %>% 
  filter(duplicates_by_variable)

duplicating_variables
```

## Addressing blanks

Function that checks whether there are blanks for different variables

```{r}
blank_check <- function(in_data,in_variable_name)
{
  check_01 <- 
  in_data %>% 
    filter(!!sym(in_variable_name) == "") %>% 
  distinct()

out <- nrow(check_01) > 0

return(out)
}
```

Checking duplicates for each variable in the dataframe.

```{r}
blanks_by_variable <- 
  column_meta_data_dedupe %>% 
  pull(variable) %>% 
  map_lgl(~blank_check(in_data = prep_building,.))

blanks_by_variable
```

Add this to the `column_meta_data_dedupe` dataframe.

```{r}
column_meta_data_dedupe <- 
  column_meta_data_dedupe %>% 
  add_column(
    blanks_by_variable = blanks_by_variable
  )

column_meta_data_dedupe
```

Create dataframe of only variables that have blanks.

```{r}
blanking_variables <- 
  column_meta_data_dedupe %>% 
  filter(blanks_by_variable)

blanking_variables
```

Grouping variables.

```{r}
blanking_variables <- 
  blanking_variables %>% 
  mutate(group = 
           case_when(variable == "position_open_date" ~ "dates",
                     variable == "position_close_date" ~ "dates",
                     variable == "announcement_closing_type" ~ "announcement_closing_type",
                     variable == "announcement_closing_type_description" ~ "announcement_closing_type",
                     variable == "application_limit" ~ "application_limit",
                     variable == "application_limit_set" ~ "application_limit",
                     variable == "agency_level" ~ "agency_level",
                     variable == "agency_level_sort" ~ "agency_level",
                     variable == "hiring_agency_code" ~ "agency_level",
                     variable == "hiring_agency_name" ~ "agency_level",
                     variable == "hiring_department_code" ~ "agency_level",
                     variable == "hiring_department_name" ~ "agency_level",
                     variable == "maximum_salary" ~ "maximum_salary",
                     variable == "minimum_salary" ~ "maximum_salary",
                     variable == "maximum_grade" ~ "maximum_salary",
                     variable == "minimum_grade" ~ "maximum_salary",
                     variable == "series" ~ "series",
                     variable == "series_title" ~ "series",
                     variable == "mission_critical_code" ~ "mission_critical_code",
                     variable == "mission_critical_code_description" ~ "mission_critical_code",
                     variable == "mission_critical_source" ~ "mission_critical_code",
                     TRUE ~ variable
                     ))

blanking_variables
```

Create list of groups.

```{r}
blanking_groups <- 
  blanking_variables %>% 
  select(group) %>% 
  distinct() %>% 
  arrange(group) %>% 
  pull()

blanking_groups
```

Also create a building dataframe to add other dataframes where missing have been removed.

```{r}
prep_building_missing <- prep_uniq
```

### Blanks removal

```{r}
step_n <- 0
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `agency_level`

Removing blanks produced by

```{r}
blanking_variable <- 
  blanking_variables %>% 
  filter(group == blanking_groups[step_n])

blanking_variable
```

Reviewing blanks.

```{r}
blanks_01 <- 
  prep_building %>% 
  filter(if_any(all_of(blanking_variable %>% pull(variable)),  ~.x == "")) %>% 
  select(all_of(blanking_variable %>% pull(variable)),everything())

blanks_01
```

Blanks appear to be caused by a few columns:

1.    `hiring_agency_code` appears to be blank for agencies that don't fall into departments
2.    `hiring_agency_name` appears to be blank for the Air Force

Let's count the extent of the issues.

```{r}
blanking_variable %>% 
  pull(variable) %>% 
  map(
    ~blanks_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(!!sym(.x) == "")
  )
```

A lot of blanks in `hiring_agency_code`, some in `hiring_agency_name`, and very few in `hiring_department_name`.

Let's see what happens when we assign rows with `hiring_department_name` "Department of the Air Force" with "Department of Defense" and place "Department of the Air Force" in `hiring_agency_name`.

```{r}
blanks_01 <- 
  prep_building %>% 
  mutate(
    hiring_agency_name = case_when(
      hiring_agency_name == "" & hiring_department_name == "Department of the Air Force" ~ "Department of the Air Force",
      TRUE ~ hiring_agency_name
    ),
    hiring_department_name = case_when(
      hiring_agency_name == "" & hiring_department_name == "Department of the Air Force" ~ "Department of Defense",
      TRUE ~ hiring_department_name
    )
  ) %>% 
  filter(if_any(all_of(blanking_variable %>% pull(variable)),  ~.x == "")) %>% 
  select(all_of(blanking_variable %>% pull(variable)),everything())

blanks_01
```

That gets rid of many. Now let's investigate `hiring_agency_code`.

```{r}
prep_building %>% 
  count(hiring_agency_code,hiring_agency_name)
```

This doesn't appear to add addition information to `hiring_agency_name`. We will drop and test for blanks.

```{r}
blanks_01 <- 
  prep_building %>% 
  mutate(
    hiring_agency_name = case_when(
      hiring_agency_name == "" & hiring_department_name == "Department of the Air Force" ~ "Department of the Air Force",
      TRUE ~ hiring_agency_name
    ),
    hiring_department_name = case_when(
      hiring_agency_name == "" & hiring_department_name == "Department of the Air Force" ~ "Department of Defense",
      TRUE ~ hiring_department_name
    )
  ) %>% 
  select(-hiring_agency_code) %>% 
  filter(if_any(c(hiring_agency_name,hiring_department_name),  ~.x == "")) %>% 
  select(hiring_agency_name,hiring_department_name,everything())

blanks_01
```

Still a few thousand with missing `hiring_agency_name`. Replace these with "(Missing)" and test.

```{r}
blanks_01 <- 
  prep_building %>% 
  mutate(
    hiring_agency_name = case_when(
      hiring_agency_name == "" & hiring_department_name == "Department of the Air Force" ~ "Department of the Air Force",
      TRUE ~ hiring_agency_name
    ),
    hiring_department_name = case_when(
      hiring_agency_name == "" & hiring_department_name == "Department of the Air Force" ~ "Department of Defense",
      TRUE ~ hiring_department_name
    )
  ) %>% 
  select(-hiring_agency_code) %>% 
  mutate(
    hiring_agency_name = case_when(
      hiring_agency_name == "" ~ "(Missing)",
      TRUE ~ hiring_agency_name
    )
  ) %>% 
  filter(if_any(c(hiring_agency_name,hiring_department_name),  ~.x == "")) %>% 
  select(hiring_agency_name,hiring_department_name,everything())

blanks_01
```

Now a few with missing `hiring_department_name`. These appear to have "Public Defender Service for the District of Columbia" in `hiring_agency_name`. Replace the blank in `hiring_department_name` for "Other Agencies and Independent Organizations".

```{r}
blanks_01 <- 
  prep_building %>% 
  mutate(
    hiring_agency_name = case_when(
      hiring_agency_name == "" & hiring_department_name == "Department of the Air Force" ~ "Department of the Air Force",
      TRUE ~ hiring_agency_name
    ),
    hiring_department_name = case_when(
      hiring_agency_name == "" & hiring_department_name == "Department of the Air Force" ~ "Department of Defense",
      TRUE ~ hiring_department_name
    )
  ) %>% 
  select(-hiring_agency_code) %>% 
  mutate(
    hiring_agency_name = case_when(
      hiring_agency_name == "" ~ "(Missing)",
      TRUE ~ hiring_agency_name
    )
  ) %>% 
  mutate(
    hiring_department_name = case_when(
      hiring_department_name == "" & hiring_agency_name == "Public Defender Service for the District of Columbia" ~ 
        "Other Agencies and Independent Organizations",
      TRUE ~ hiring_department_name
    )
  ) %>% 
  filter(if_any(c(hiring_agency_name,hiring_department_name),  ~.x == "")) %>% 
  select(hiring_agency_name,hiring_department_name,everything())

blanks_01
```

Now apply this to the entire dataframe.

```{r}
prep_xx <- 
  prep_building %>% 
  mutate(
    hiring_agency_name = case_when(
      hiring_agency_name == "" & hiring_department_name == "Department of the Air Force" ~ "Department of the Air Force",
      TRUE ~ hiring_agency_name
    ),
    hiring_department_name = case_when(
      hiring_agency_name == "" & hiring_department_name == "Department of the Air Force" ~ "Department of Defense",
      TRUE ~ hiring_department_name
    )
  ) %>% 
  select(-hiring_agency_code) %>% 
  mutate(
    hiring_agency_name = case_when(
      hiring_agency_name == "" ~ "(Missing)",
      TRUE ~ hiring_agency_name
    )
  ) %>% 
  mutate(
    hiring_department_name = case_when(
      hiring_department_name == "" & hiring_agency_name == "Public Defender Service for the District of Columbia" ~ 
        "Other Agencies and Independent Organizations",
      TRUE ~ hiring_department_name
    )
  ) %>% 
  select(usajobs_control_number,hiring_agency_name,hiring_department_name)
```

Join to building dataframe.

```{r}
prep_building_missing <- prep_building_missing %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `mission_critical_code`

Removing blanks produced by

```{r}
blanking_variable <- 
  blanking_variables %>% 
  filter(group == blanking_groups[step_n])

blanking_variable
```

Reviewing blanks.

```{r}
blanks_01 <- 
  prep_building %>% 
  filter(if_any(all_of(blanking_variable %>% pull(variable)),  ~.x == "")) %>% 
  select(all_of(blanking_variable %>% pull(variable)),everything())

blanks_01
```

Let's count the extent of the issues.

```{r}
blanking_variable %>% 
  pull(variable) %>% 
  map(
    ~blanks_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(!!sym(.x) == "")
  )
```

The same number or missing from each group.

Let's count the entries here.

```{r}
prep_building %>% 
  group_by(across(all_of(blanking_variable %>% pull(variable)))) %>% 
  count()
```

The case majority are missing. For all of these, we will recode to "(Missing)". We'll do that and test.

```{r}
blanks_01 <- 
  prep_building %>% 
  mutate(
    across(
      .cols = all_of(blanking_variable %>% pull(variable)),
      .fns = ~case_when(
        . == "" ~ "(Missing)",
        . == "NULL" ~ "(Missing)",
        is.na(.) ~ "(Missing)",
        TRUE ~ .
      )
    )
  ) %>% 
  filter(if_any(all_of(blanking_variable %>% pull(variable)),  ~.x == "")) %>% 
  select(all_of(blanking_variable %>% pull(variable)),everything())

blanks_01
```

Now apply this to the entire dataframe.

```{r}
prep_xx <- 
  prep_building %>% 
   mutate(
    across(
      .cols = all_of(blanking_variable %>% pull(variable)),
      .fns = ~case_when(
        . == "" ~ "(Missing)",
        . == "NULL" ~ "(Missing)",
        is.na(.) ~ "(Missing)",
        TRUE ~ .
      )
    )
  ) %>% 
    select(usajobs_control_number,all_of(blanking_variable %>% pull(variable)))
```

Join to building dataframe.

```{r}
prep_building_missing <- prep_building_missing %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `series_title`

Removing blanks produced by

```{r}
blanking_variable <- 
  blanking_variables %>% 
  filter(group == blanking_groups[step_n])

blanking_variable
```

Reviewing blanks.

```{r}
blanks_01 <- 
  prep_building %>% 
  filter(if_any(all_of(blanking_variable %>% pull(variable)),  ~.x == "")) %>% 
  select(all_of(blanking_variable %>% pull(variable)),everything())

blanks_01
```

Let's count the extent of the issues.

```{r}
blanking_variable %>% 
  pull(variable) %>% 
  map(
    ~blanks_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(!!sym(.x) == "")
  )
```

The same number or missing from each group.

Let's count the entries here. We'll add `series` as well.

```{r}
prep_building %>% 
  group_by(across(all_of(blanking_variable %>% pull(variable))),series) %>% 
  count()
```

Just one row is blank. We'll replace `series_title` with "(Missing)" and test.

```{r}
blanks_01 <- 
  prep_building %>% 
  mutate(
    series_title = case_when(
      series_title == "" ~ "(Missing)",
      TRUE ~ series_title
    )
  ) %>% 
  filter(if_any(all_of(blanking_variable %>% pull(variable)),  ~.x == "")) %>% 
  select(all_of(blanking_variable %>% pull(variable)),everything())

blanks_01
```

Now apply this to the entire dataframe.

```{r}
prep_xx <- 
  prep_building %>% 
    mutate(
    series_title = case_when(
      series_title == "" ~ "(Missing)",
      TRUE ~ series_title
    )
  ) %>% 
    select(usajobs_control_number,all_of(blanking_variable %>% pull(variable)))
```

Join to building dataframe.

```{r}
prep_building_missing <- prep_building_missing %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing blanks produced by

```{r}
blanking_variable <- 
  blanking_variables %>% 
  filter(group == blanking_groups[step_n])

blanking_variable
```

All done!

### Combinding blank data

#### Create dataframe off non-blank columns

Collect column names of variables not creating duplicates.

```{r}
non_blanking_variables <- 
  column_meta_data_dedupe %>% 
  filter(!blanks_by_variable) %>% 
  pull(variable)

non_blanking_variables
```

Create dataframe of only these variables.

```{r}
prep_xx <- 
  prep_building %>% 
  select(any_of(non_blanking_variables)) %>% 
  distinct()

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Join to building dataframe.

```{r}
prep_building_missing <- prep_building_missing %>% 
  left_join(prep_xx)
```

### Checking blanking columns have been fixed

Created a dataframe of variable metadata.

```{r}
column_meta_data_dedupe_missing <- 
  prep_building_missing %>% 
  contents()

column_meta_data_dedupe_missing <- 
  column_meta_data_dedupe_missing$contents %>% 
  as.data.frame()

column_meta_data_dedupe_missing <-
  column_meta_data_dedupe_missing %>% 
  rownames_to_column(var = "variable")

column_meta_data_dedupe_missing
```

Checking duplicates for each variable in the dataframe.

```{r}
blanks_by_variable <- 
  column_meta_data_dedupe_missing %>% 
  pull(variable) %>% 
  map_lgl(~blank_check(in_data = prep_building_missing,.))

blanks_by_variable
```

Add this to the `column_meta_data_dedupe` dataframe.

```{r}
column_meta_data_dedupe_missing <- 
  column_meta_data_dedupe_missing %>% 
  add_column(
    blanks_by_variable = blanks_by_variable
  )

column_meta_data_dedupe_missing
```

Create dataframe of only variables that have blanks.

```{r}
blanking_variables <- 
  column_meta_data_dedupe_missing %>% 
  filter(blanks_by_variable)

blanking_variables
```

## Addressing nulls

Create variable to identify columns with NAs.

```{r}
column_meta_data_dedupe_missing <- 
  column_meta_data_dedupe_missing %>% 
  mutate(
    nulls_by_variable = case_when(
      NAs == 0 ~ FALSE,
      NAs > 0 ~ TRUE
    )
  )

column_meta_data_dedupe_missing
```

Create dataframe of only variables that have blanks.

```{r}
nulling_variables <- 
  column_meta_data_dedupe_missing %>% 
  filter(nulls_by_variable)

nulling_variables
```

Grouping variables.

```{r}
nulling_variables <- 
  nulling_variables %>% 
  mutate(group = 
           case_when(variable == "position_open_date" ~ "dates",
                     variable == "position_close_date" ~ "dates",
                     variable == "announcement_closing_type" ~ "announcement_closing_type",
                     variable == "announcement_closing_type_description" ~ "announcement_closing_type",
                     variable == "application_limit" ~ "application_limit",
                     variable == "application_limit_set" ~ "application_limit",
                     variable == "agency_level" ~ "agency_level",
                     variable == "agency_level_sort" ~ "agency_level",
                     variable == "hiring_agency_code" ~ "agency_level",
                     variable == "hiring_agency_name" ~ "agency_level",
                     variable == "hiring_department_code" ~ "agency_level",
                     variable == "hiring_department_name" ~ "agency_level",
                     variable == "maximum_salary" ~ "maximum_salary",
                     variable == "minimum_salary" ~ "maximum_salary",
                     variable == "maximum_grade" ~ "maximum_salary",
                     variable == "minimum_grade" ~ "maximum_salary",
                     variable == "series" ~ "series",
                     variable == "series_title" ~ "series",
                     variable == "mission_critical_code" ~ "mission_critical_code",
                     variable == "mission_critical_code_description" ~ "mission_critical_code",
                     variable == "mission_critical_source" ~ "mission_critical_code",
                     TRUE ~ variable
                     ))

nulling_variables
```

Create list of groups.

```{r}
nulling_groups <- 
  nulling_variables %>% 
  select(group) %>% 
  distinct() %>% 
  arrange(group) %>% 
  pull()

nulling_groups
```

Also create a building dataframe to add other dataframes where missing have been removed.

```{r}
prep_building_missing_nulls <- prep_uniq
```

### Nulls removal

```{r}
step_n <- 0
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `agency_level`

Removing nulls produced by

```{r}
nulling_variable <- 
  nulling_variables %>% 
  filter(group == nulling_groups[step_n])

nulling_variable
```

Reviewing blanks.

```{r}
nulls_01 <- 
  prep_building_missing %>% 
  filter(if_any(all_of(nulling_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(nulling_variable %>% pull(variable)),everything())

nulls_01
```

Let's count the extent of the issues.

```{r}
nulling_variable %>% 
  pull(variable) %>% 
  map(
    ~nulls_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's count the entries here.

```{r}
prep_building_missing %>% 
  group_by(across(all_of(nulling_variable %>% pull(variable))),hiring_department_name, hiring_agency_name) %>% 
  count() %>% 
  arrange(agency_level_sort)
```

None of these columns appear to added information that isn't part of `hiring_department_name` or `hiring_agency_name`. We will drop them

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `maximum_salary`

Removing nulls produced by

```{r}
nulling_variable <- 
  nulling_variables %>% 
  filter(group == nulling_groups[step_n])

nulling_variable
```

Reviewing blanks.

```{r}
nulls_01 <- 
  prep_building_missing %>% 
  filter(if_any(all_of(nulling_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(nulling_variable %>% pull(variable)),everything())

nulls_01
```

Let's count the extent of the issues.

```{r}
nulling_variable %>% 
  pull(variable) %>% 
  map(
    ~nulls_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's count the entries here.

```{r}
prep_building_missing %>% 
  group_by(across(all_of(nulling_variable %>% pull(variable)))) %>% 
  count() %>% 
  arrange()
```

Very few examples. We will transform these to "(Missing)" and then test.

```{r}
nulls_01 <- 
  prep_building_missing %>% 
  mutate(
    across(
      .cols = all_of(nulling_variable %>% pull(variable)),
      .fns = ~case_when(is.na(.) ~ "(Missing)",
                        TRUE ~ .)
    )
  ) %>% 
  filter(if_any(all_of(nulling_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(nulling_variable %>% pull(variable)),everything())

nulls_01
```

Now apply this to the entire dataframe.

```{r}
prep_xx <- 
  prep_building_missing %>% 
    mutate(
    across(
      .cols = all_of(nulling_variable %>% pull(variable)),
      .fns = ~case_when(is.na(.) ~ "(Missing)",
                        TRUE ~ .)
    )
  ) %>% 
    select(usajobs_control_number,all_of(nulling_variable %>% pull(variable)))
```

Join to building dataframe.

```{r}
prep_building_missing_nulls <- prep_building_missing_nulls %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `pay_scale`

Removing nulls produced by

```{r}
nulling_variable <- 
  nulling_variables %>% 
  filter(group == nulling_groups[step_n])

nulling_variable
```

Reviewing blanks.

```{r}
nulls_01 <- 
  prep_building_missing %>% 
  filter(if_any(all_of(nulling_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(nulling_variable %>% pull(variable)),everything())

nulls_01
```

Let's count the extent of the issues.

```{r}
nulling_variable %>% 
  pull(variable) %>% 
  map(
    ~nulls_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's count the entries here.

```{r}
prep_building_missing %>% 
  group_by(across(all_of(nulling_variable %>% pull(variable))),pay_scale_description) %>% 
  count() %>% 
  arrange()
```

No additional information in `pay_scale`. We will drop this column.

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `second_announcement`

Removing nulls produced by

```{r}
nulling_variable <- 
  nulling_variables %>% 
  filter(group == nulling_groups[step_n])

nulling_variable
```

Reviewing blanks.

```{r}
nulls_01 <- 
  prep_building_missing %>% 
  filter(if_any(all_of(nulling_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(nulling_variable %>% pull(variable)),everything())

nulls_01
```

Let's count the extent of the issues.

```{r}
nulling_variable %>% 
  pull(variable) %>% 
  map(
    ~nulls_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's count the entries here.

```{r}
prep_building_missing %>% 
  group_by(across(all_of(nulling_variable %>% pull(variable)))) %>% 
  count() %>% 
  arrange()
```

Completely blank. We will drop.

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `series`

Removing nulls produced by

```{r}
nulling_variable <- 
  nulling_variables %>% 
  filter(group == nulling_groups[step_n])

nulling_variable
```

Reviewing blanks.

```{r}
nulls_01 <- 
  prep_building_missing %>% 
  filter(if_any(all_of(nulling_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(nulling_variable %>% pull(variable)),everything())

nulls_01
```

Let's count the extent of the issues.

```{r}
nulling_variable %>% 
  pull(variable) %>% 
  map(
    ~nulls_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's count the entries here.

```{r}
prep_building_missing %>% 
  group_by(across(all_of(nulling_variable %>% pull(variable)))) %>% 
  count()
```

All examples have missing series titles. We will transform these to "(Missing)" and then test.

```{r}
nulls_01 <- 
  prep_building_missing %>% 
  mutate(
    across(
      .cols = all_of(nulling_variable %>% pull(variable)),
      .fns = ~case_when(is.na(.) ~ "(Missing)",
                        TRUE ~ .)
    )
  ) %>% 
  filter(if_any(all_of(nulling_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(nulling_variable %>% pull(variable)),everything())

nulls_01
```

Now apply this to the entire dataframe.

```{r}
prep_xx <- 
  prep_building_missing %>% 
    mutate(
    across(
      .cols = all_of(nulling_variable %>% pull(variable)),
      .fns = ~case_when(is.na(.) ~ "(Missing)",
                        TRUE ~ .)
    )
  ) %>% 
    select(usajobs_control_number,all_of(nulling_variable %>% pull(variable)))
```

Join to building dataframe.

```{r}
prep_building_missing_nulls <- prep_building_missing_nulls %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `total_openings`

Removing nulls produced by

```{r}
nulling_variable <- 
  nulling_variables %>% 
  filter(group == nulling_groups[step_n])

nulling_variable
```

Reviewing blanks.

```{r}
nulls_01 <- 
  prep_building_missing %>% 
  filter(if_any(all_of(nulling_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(nulling_variable %>% pull(variable)),everything())

nulls_01
```

Let's count the extent of the issues.

```{r}
nulling_variable %>% 
  pull(variable) %>% 
  map(
    ~nulls_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's count the entries here.

```{r}
prep_building_missing %>% 
  group_by(across(all_of(nulling_variable %>% pull(variable)))) %>% 
  count() %>% 
  arrange()
```

We will replace the nulls with the median value of the existing values for the column and create an indicator for rows that had data missing.

First, calculate the median value.

```{r}
total_openings_median <- 
  prep_building_missing %>% 
  summarise(
    median(total_openings,na.rm = TRUE)
  ) %>% 
  pull()

total_openings_median
```

Now apply to nulls and create indicator.

```{r}
nulls_01 <- 
  prep_building_missing %>% 
  mutate(
    total_openings_na = case_when(is.na(total_openings) ~ 1,
                                  TRUE ~ 0),
    total_openings = case_when(is.na(total_openings) ~ total_openings_median,
                                  TRUE ~ total_openings),
  ) %>% 
  filter(if_any(all_of(c(nulling_variable %>% pull(variable),"total_openings_na")),  ~is.na(.x))) %>% 
  select(all_of(nulling_variable %>% pull(variable)),total_openings_na,everything())

nulls_01
```

Now apply this to the entire dataframe.

```{r}
prep_xx <- 
  prep_building_missing %>% 
    mutate(
    total_openings_na = case_when(is.na(total_openings) ~ 1,
                                  TRUE ~ 0),
    total_openings = case_when(is.na(total_openings) ~ total_openings_median,
                                  TRUE ~ total_openings),
  ) %>% 
    select(usajobs_control_number,all_of(nulling_variable %>% pull(variable)),total_openings_na)
```

Join to building dataframe.

```{r}
prep_building_missing_nulls <- prep_building_missing_nulls %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `telework_eligibility_description`

Removing nulls produced by

```{r}
nulling_variable <- 
  nulling_variables %>% 
  filter(group == nulling_groups[step_n])

nulling_variable
```

Reviewing blanks.

```{r}
nulls_01 <- 
  prep_building_missing %>% 
  filter(if_any(all_of(nulling_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(nulling_variable %>% pull(variable)),everything())

nulls_01
```

Let's count the extent of the issues.

```{r}
nulling_variable %>% 
  pull(variable) %>% 
  map(
    ~nulls_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's count the entries here.

```{r}
prep_building_missing %>% 
  group_by(across(all_of(nulling_variable %>% pull(variable)))) %>% 
  count() %>% 
  arrange()
```

`r nulling_variable %>% pull(variable)` is either null or has the value "NULL".

We will drop this column.

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing nulls produced by

```{r}
nulling_variable <- 
  nulling_variables %>% 
  filter(group == nulling_groups[step_n])

nulling_variable
```

Done with nulls.

### Combinding nulls data

#### Create dataframe off non-null columns

Collect column names of variables not creating duplicates.

```{r}
non_nulling_variables <- 
  column_meta_data_dedupe_missing %>% 
  filter(!nulls_by_variable) %>% 
  pull(variable)

non_nulling_variables
```

Create dataframe of only these variables.

```{r}
prep_xx <- 
  prep_building_missing %>% 
  select(any_of(non_nulling_variables)) %>% 
  distinct()

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Join to building dataframe.

```{r}
prep_building_missing_nulls <- prep_building_missing_nulls %>% 
  left_join(prep_xx)
```

### Checking nulling columns have been fixed

Created a dataframe of variable metadata.

```{r}
column_meta_data_dedupe_missing_nulls <- 
  prep_building_missing_nulls %>% 
  contents()

column_meta_data_dedupe_missing_nulls <- 
  column_meta_data_dedupe_missing_nulls$contents %>% 
  as.data.frame()

column_meta_data_dedupe_missing_nulls <-
  column_meta_data_dedupe_missing_nulls %>% 
  rownames_to_column(var = "variable")

column_meta_data_dedupe_missing_nulls
```

Function that checks whether there are duplicates for different variables

```{r}
null_check <- function(in_data,in_variable_name)
{
  check_01 <- 
  in_data %>% 
    filter(is.na(!!sym(in_variable_name))) %>% 
  distinct()

out <- nrow(check_01) > 0

return(out)
}
```

Checking duplicates for each variable in the dataframe.

```{r}
nulls_by_variable <- 
  column_meta_data_dedupe_missing_nulls %>% 
  pull(variable) %>% 
  map_lgl(~null_check(in_data = prep_building_missing_nulls,.))

nulls_by_variable
```

Add this to the `column_meta_data_dedupe_missing_nulls` dataframe.

```{r}
column_meta_data_dedupe_missing_nulls <- 
  column_meta_data_dedupe_missing_nulls %>% 
  add_column(
    nulls_by_variable = nulls_by_variable
  )

column_meta_data_dedupe_missing_nulls
```

Create dataframe of only variables that have blanks.

```{r}
nulling_variables <- 
  column_meta_data_dedupe_missing_nulls %>% 
  filter(nulls_by_variable)

nulling_variables
```

## Attaching telework data

Left join telework data to announcement data and then review.

```{r}
prep_building_missing_nulls %>% 
  left_join(prep_telework, by = "usajobs_control_number", suffix = c("_building", "_telework")) %>% 
  skim(usajobs_control_number,starts_with("telework_eligible"))
```

Left join and count results.

```{r}
prep_building_missing_nulls %>% 
  left_join(prep_telework, by = "usajobs_control_number", suffix = c("_building", "_telework")) %>% 
  count(telework_eligible_building, telework_eligible_telework)
```

`telework_eligible_building` has no variable and `telework_eligible_telework` has lots of NULLs.

We will attach the data and drop `telework_eligible_building` from the `prep_building_missing_nulls` dataframe. We will transform this into indicators for a position being telework eligible and for the value being null.

```{r}
prep_xx <- 
  prep_building_missing_nulls %>% 
  select(-telework_eligible) %>% 
  left_join(prep_telework, by = "usajobs_control_number") %>% 
  mutate(
    telework_eligible = case_when(telework_eligible == 1 ~ "y",
                                  telework_eligible == 0 ~ "n")
  ) %>% 
  character_to_indicator(in_unique_identifier = params$unique_identifier_1, in_variable = telework_eligible, in_prefix = "telework_eligible") %>% 
  select(-telework_eligible_n)

prep_xx %>% 
  skim(starts_with("telework_eligible"))
```

Join onto existing data.

```{r}
prep_building_missing_nulls <- 
  prep_building_missing_nulls %>% 
  left_join(prep_xx)

prep_building_missing_nulls %>% 
  skim(starts_with("telework_eligible"))
```

## Attaching hiring path data

Left join hiring path data to announcement data and then review.

```{r}
prep_building_missing_nulls %>% 
  left_join(prep_hiring_path, by = "usajobs_control_number", suffix = c("_building", "_telework")) %>% 
  skim(usajobs_control_number,starts_with("hiring_path"))
```

Looks good, but we have some nulls. Let's review those.

```{r}
names_hiring_path <- prep_hiring_path %>% 
  select(starts_with("hiring_path")) %>% 
  names()
prep_building_missing_nulls %>% 
  left_join(prep_hiring_path, by = "usajobs_control_number", suffix = c("_building", "_telework")) %>% 
  filter(if_any(names_hiring_path,  ~is.na(.x))) %>% 
  select(all_of(names_hiring_path),everything())
```

All the same entries. We'll set these `hiring_path_na` to 1 for these rows and the other `hiring_path_*` variables to 0.

```{r}
prep_building_missing_nulls %>% 
  left_join(prep_hiring_path, by = "usajobs_control_number", suffix = c("_building", "_telework")) %>% 
  mutate(
    hiring_path_na = case_when(is.na(hiring_path_na) ~ 1, 
                               TRUE ~ hiring_path_na)
  ) %>% 
  mutate(
    across(
      .cols = starts_with("hiring_path"),
      .fns = ~case_when(is.na(.) ~ 0, 
                               TRUE ~ .)
    )
  ) %>% 
  skim(usajobs_control_number,starts_with("hiring_path"))
```

```{r}
prep_xx <- 
  prep_uniq %>% 
 left_join(prep_hiring_path, by = "usajobs_control_number", suffix = c("_building", "_telework")) %>% 
  mutate(
    hiring_path_na = case_when(is.na(hiring_path_na) ~ 1, 
                               TRUE ~ hiring_path_na)
  ) %>% 
  mutate(
    across(
      .cols = starts_with("hiring_path"),
      .fns = ~case_when(is.na(.) ~ 0, 
                               TRUE ~ .)
    )
  )

prep_xx %>% 
  skim()
```

Join onto existing data.

```{r}
prep_building_missing_nulls <- 
  prep_building_missing_nulls %>% 
  left_join(prep_xx)

prep_building_missing_nulls %>% 
  skim(starts_with("hiring_path"))
```

## Attaching appointment and work schedule data

Left join appointment and work schedule data to announcement data and then review.

```{r}
prep_xx <- 
  prep_building_missing_nulls %>% 
  left_join(prep_work_schedule_appointment, by = "usajobs_control_number", suffix = c("_building", "_join")) %>% 
  select(usajobs_control_number,starts_with("appointment_"),starts_with("work_schedule_")) 
prep_xx %>% 
  select(order(colnames(prep_xx))) %>% 
  skim()
```

Collect names.

```{r}
names_xx <- prep_work_schedule_appointment %>% names()
```

Function to count differences between columns by variable prefix.

```{r}
counting_differences <- function(in_data,in_variable_prefix)
{
   in_data %>%
    select(starts_with(in_variable_prefix)) %>%
    group_by(across(all_of(starts_with(in_variable_prefix)))) %>% 
    count()
}
```

Counting differences.

```{r}
 names_xx[-1] %>% 
  map(
    ~counting_differences(in_data = prep_xx, in_variable_prefix = .)
  )
```

We want to take the values from the attached dataframe, unless they are null.

Exact matches.

```{r}
 prep_xx_01 <- 
  prep_building_missing_nulls %>% 
  select(all_of(names_xx)) %>% 
  inner_join(prep_work_schedule_appointment)
```

`r params$unique_identifier_1` not in `prep_work_schedule_appointment`.

```{r}
 prep_xx_02 <- 
  prep_building_missing_nulls %>% 
  select(usajobs_control_number) %>% 
  anti_join(prep_work_schedule_appointment) %>% 
  inner_join(prep_building_missing_nulls) %>% 
  select(all_of(names_xx)) 
```

In `prep_work_schedule_appointment`, but not above two.

```{r}
prep_xx_03 <- 
  prep_work_schedule_appointment %>% 
  anti_join(
    prep_xx_01 %>% 
      bind_rows(prep_xx_02) %>% 
      select(usajobs_control_number) %>% 
      distinct()

  )
```
Bind the above.

```{r}
prep_xx_04 <- 
  prep_xx_01 %>% 
  bind_rows(prep_xx_02) %>% 
  bind_rows(prep_xx_03) %>% 
  distinct()
```

Join to the unique `r params$unique_identifier_1`.

```{r}
prep_xx <- 
  prep_uniq %>% 
  left_join(prep_xx_04)

prep_xx %>% 
  skim()
```

Join onto existing data, while droping the original colums.

```{r}
prep_building_missing_nulls <-
  prep_building_missing_nulls  %>% 
  select(-all_of(names_xx[-1])) %>% 
  left_join(prep_xx)

prep_building_missing_nulls %>% 
  skim(all_of(names_xx))
```

## Attaching series data

Left join series data to announcement data and then review.

```{r}
prep_xx <- 
  prep_building_missing_nulls %>% 
  left_join(prep_series, by = "usajobs_control_number", suffix = c("_building", "_join")) %>% 
  select(usajobs_control_number,starts_with("series_")) 
prep_xx %>% 
  select(order(colnames(prep_xx))) %>% 
  skim()
```

Collect names.

```{r}
names_xx <- prep_series %>% names()
```

Counting differences.

```{r}
 names_xx[-1] %>% 
  map(
    ~counting_differences(in_data = prep_xx, in_variable_prefix = .)
  )
```

We want to take the values from the attached dataframe, unless they are null.

Exact matches.

```{r}
 prep_xx_01 <-
  prep_building_missing_nulls %>% 
  select(all_of(names_xx)) %>% 
  inner_join(prep_series)
```

`r params$unique_identifier_1` not in `prep_work_schedule_appointment`.

```{r}
 prep_xx_02 <-
  prep_building_missing_nulls %>% 
  select(usajobs_control_number) %>% 
  anti_join(prep_series) %>% 
  inner_join(prep_building_missing_nulls) %>% 
  select(all_of(names_xx)) 
```

In `prep_work_schedule_appointment`, but not above two.

```{r}
prep_xx_03 <- 
  prep_series %>% 
  anti_join(
    prep_xx_01 %>% 
      bind_rows(prep_xx_02) %>% 
      select(usajobs_control_number) %>% 
      distinct()

  )
```

Bind the above.

```{r}
prep_xx_04 <- 
  prep_xx_01 %>% 
  bind_rows(prep_xx_02) %>% 
  bind_rows(prep_xx_03) %>% 
  distinct()
```

Join to the unique `r params$unique_identifier_1`.

```{r}
prep_xx <- 
  prep_uniq %>% 
  left_join(prep_xx_04)

prep_xx %>% 
  skim()
```

Join onto existing data, while dropping the original colums.

```{r}
prep_building_missing_nulls <-
  prep_building_missing_nulls  %>% 
  select(-all_of(names_xx[-1])) %>% 
  left_join(prep_xx)

prep_building_missing_nulls %>% 
  skim(all_of(names_xx))
```

## Attaching locations data

Left join locations data to announcement data and then review.

```{r}
prep_xx <-
  prep_building_missing_nulls %>% 
  left_join(prep_location, by = "usajobs_control_number", suffix = c("_building", "_join")) %>% 
  select(usajobs_control_number,ends_with( c("_building", "_join")),position_location_city, 
         position_location_country, position_location_geo_location_code,
         position_location_state,position_location_zip)
prep_xx %>% 
  select(order(colnames(prep_xx))) %>% 
  skim()
```

Collect names.

```{r}
names_xx <- prep_location %>% names()

names_base <- prep_building_missing_nulls %>% names()

names_overlap <- intersect(names_xx,names_base)
```

Counting differences.

```{r}
names_xx[-1] %>% 
  map(
    ~counting_differences(in_data = prep_xx, in_variable_prefix = .)
  )
```

We want to take the values from the attached dataframe, unless they are null.

Exact matches.

```{r}
 prep_xx_01 <-
  prep_building_missing_nulls %>% 
  select(all_of(names_overlap)) %>% 
  inner_join(prep_location)
```

`r params$unique_identifier_1` not in `prep_location`.

```{r}
 prep_xx_02 <-
  prep_building_missing_nulls %>% 
  select(usajobs_control_number) %>% 
  anti_join(prep_location) %>% 
  inner_join(prep_building_missing_nulls) %>% 
  select(all_of(names_overlap)) 
```

In `prep_building_missing_nulls`, but not above two.

```{r}
prep_xx_03 <- 
  prep_building_missing_nulls %>% 
  anti_join(
    prep_xx_01 %>% 
      bind_rows(prep_xx_02) %>% 
      select(usajobs_control_number) %>% 
      distinct()

  ) %>% 
  select(all_of(names_overlap))  %>% 
      distinct()

```

Bind the above.

```{r}
prep_xx_04 <- 
  prep_xx_01 %>% 
  bind_rows(prep_xx_02) %>% 
  bind_rows(prep_xx_03) %>% 
  distinct()
```

Join to the unique `r params$unique_identifier_1`.

```{r}
prep_xx <- 
  prep_uniq %>% 
  left_join(prep_xx_04)

prep_xx %>% 
  skim()
```

# CONTINUE FROM HERE

Is this covering the whole year?

```{r}
prep_building_missing_nulls %>% 
  count(position_open_date) %>% 
  ggplot(aes(x = position_open_date, y = n)) + 
  geom_line()
```

```{r}
prep_building_missing_nulls %>% 
  count(position_open_date,maximum_grade) %>% 
  ggplot(aes(x = position_open_date, y = n, color = maximum_grade)) + 
  geom_line() 
+
  facet_wrap('maximum_grade', scales = "free")
```

```{r}
prep_building_missing_nulls %>% 
  count(position_open_date,minimum_grade) %>% 
  ggplot(aes(x = position_open_date, y = n, color = minimum_grade)) + 
  geom_line() 
```

```{r}
prep_building_missing_nulls %>% 
  count(position_close_date) %>% 
  ggplot(aes(x = position_close_date, y = n)) + 
  geom_line()
```

`telework_eligible_building` has no variable and `telework_eligible_telework` has lots of NULLs.

We will attach the data and drop `telework_eligible_building` from the `prep_building_missing_nulls` dataframe. We will transform this into indicators for a position being telework eligible and for the value being null.

```{r}
prep_building_missing_nulls <- 
  prep_building_missing_nulls %>% 
  select(-telework_eligible) %>% 
  left_join(prep_telework, by = "usajobs_control_number") %>% 
  mutate(
    telework_eligible = case_when(telework_eligible == 1 ~ "y",
                                  telework_eligible == 0 ~ "n")
  ) %>% 
  character_to_indicator(in_unique_identifier = params$unique_identifier_1, in_variable = telework_eligible, in_prefix = "telework_eligible") %>% 
  select(-telework_eligible_n)

prep_building_missing_nulls %>% 
  skim(starts_with("telework_eligible"))
```

join series, workschedule and appointment, locations

#### Join all deduplicated dataframes

```{r}
prep_xx <- 
  ls(envir=.GlobalEnv) %>% 
  as_tibble() %>% 
  filter(str_detect(value, "prep_[0-9]")) %>% 
  pull() %>% 
  mget(envir=.GlobalEnv) %>% 
  reduce(full_join)

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Rename dataframe.

```{r}
prep_dedup <- prep_xx
```

### Checking duplicating columns have been fixed

Created a dataframe of variable metadata

```{r}
column_meta_data_dedupe <- 
  prep_dedup %>% 
  contents()

column_meta_data_dedupe <- 
  column_meta_data_dedupe$contents %>% 
  as.data.frame()

column_meta_data_dedupe <-
  column_meta_data_dedupe %>% 
  rownames_to_column(var = "variable")

column_meta_data_dedupe
```

Checking duplicates for each variable in the dataframe.

```{r}
duplicates_by_variable <- 
  column_meta_data_dedupe %>% 
  pull(variable) %>% 
  map_lgl(~dupe_check(in_data = prep_dedup, .))

duplicates_by_variable
```

Add this to the column_meta_data dataframe.

```{r}
column_meta_data_dedupe <- 
  column_meta_data_dedupe %>% 
  add_column(
    duplicates_by_variable = duplicates_by_variable
  )

column_meta_data_dedupe
```

Create dataframe of only variables that cause duplicates.

```{r}
duplicating_variables <- 
  column_meta_data_dedupe %>% 
  filter(duplicates_by_variable)

duplicating_variables
```

## Addressing blanks

Created a dataframe of variable metadata for deduplicated data

```{r}
column_meta_data_dedupe <- 
  prep_dedup %>% 
  contents()

column_meta_data_dedupe <- 
  column_meta_data_dedupe$contents %>% 
  as.data.frame()

column_meta_data_dedupe <-
  column_meta_data_dedupe %>% 
  rownames_to_column(var = "variable")

column_meta_data_dedupe
```

Function that checks whether there are duplicates for different variables

```{r}
blank_check <- function(in_data,in_variable_name)
{
  check_01 <- 
  in_data %>% 
    filter(!!sym(in_variable_name) == "") %>% 
  distinct()

out <- nrow(check_01) > 0

return(out)
}
```

Checking duplicates for each variable in the dataframe.

```{r}
blanks_by_variable <- 
  column_meta_data_dedupe %>% 
  pull(variable) %>% 
  map_lgl(~blank_check(in_data = prep_dedup,.))

blanks_by_variable
```

Add this to the `column_meta_data_dedupe` dataframe.

```{r}
column_meta_data_dedupe <- 
  column_meta_data_dedupe %>% 
  add_column(
    blanks_by_variable = blanks_by_variable
  )

column_meta_data_dedupe
```

Create dataframe of only variables that have blanks.

```{r}
blanking_variables <- 
  column_meta_data_dedupe %>% 
  filter(blanks_by_variable)

blanking_variables
```

Grouping variables.

```{r}
blanking_variables <- 
  blanking_variables %>% 
  mutate(group = 
           case_when(variable == "position_open_date" ~ "dates",
                     variable == "position_close_date" ~ "dates",
                     variable == "announcement_closing_type" ~ "announcement_closing_type",
                     variable == "announcement_closing_type_description" ~ "announcement_closing_type",
                     variable == "application_limit" ~ "application_limit",
                     variable == "application_limit_set" ~ "application_limit",
                     variable == "agency_level" ~ "agency_level",
                     variable == "agency_level_sort" ~ "agency_level",
                     variable == "hiring_agency_code" ~ "agency_level",
                     variable == "hiring_agency_name" ~ "agency_level",
                     variable == "hiring_department_code" ~ "agency_level",
                     variable == "hiring_department_name" ~ "agency_level",
                     variable == "maximum_salary" ~ "maximum_salary",
                     variable == "minimum_salary" ~ "maximum_salary",
                     variable == "maximum_grade" ~ "maximum_salary",
                     variable == "minimum_grade" ~ "maximum_salary",
                     variable == "series" ~ "series",
                     variable == "series_title" ~ "series",
                     TRUE ~ variable
                     ))

blanking_variables
```

Create list of groups.

```{r}
blanking_groups <- 
  blanking_variables %>% 
  select(group) %>% 
  distinct() %>% 
  pull()

blanking_groups
```

### Blanks removal

```{r}
step_n <- 0
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing blanks produced by:

```{r}
blanking_variable <- 
  blanking_variables %>% 
  filter(group == blanking_groups[step_n])

blanking_variable
```

Reviewing blanks.

```{r}
blanks_01 <- 
  prep_dedup %>% 
  filter(if_any(all_of(blanking_variable %>% pull(variable)),  ~.x == "")) %>% 
  select(all_of(blanking_variable %>% pull(variable)),everything())

blanks_01
```

Blanks appear to be caused by a few columns:

1.    `hiring_agency_code` appears to be blank for agencies that don't fall into departments
2.    `hiring_agency_name` appears to be blank for the Air Force

Let's count the extent of the issues.

```{r}
blanking_variable %>% 
  pull(variable) %>% 
  map(
    ~blanks_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(!!sym(.x) == "")
  )
```

A lot of blanks in `hiring_agency_code`, some in `hiring_agency_name`, and very few in `hiring_department_name`.

Let's test filling blanks in `hiring_agency_name` with `hiring_department_name`.

```{r}
blanks_02 <- 
  blanks_01 %>% 
    select(starts_with("hiring_agency_"),starts_with("hiring_department_name"),everything()) %>%
  mutate(
    hiring_agency_name = case_when(
      hiring_agency_name == "" & !is.na(hiring_department_name) ~ hiring_department_name,
      TRUE ~ hiring_agency_name
      ),
    ) 
```

Reviewing blanks.

```{r}
  blanks_02 %>% 
  filter(if_any(all_of(blanking_variable %>% pull(variable)),  ~.x == "")) %>% 
  select(all_of(blanking_variable %>% pull(variable)),everything())
```

Reduced a lot.

Let's count the extent of the issues.

```{r}
blanking_variable %>% 
  pull(variable) %>% 
  map(
    ~blanks_02 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(!!sym(.x) == "")
  )
```

All the `hiring_agency_name` are gone.

Let's look into the issues with `hiring_department_name`.

```{r}
blanks_02 %>%
  filter(hiring_department_name == "")
```

Let's check what `hiring_department_name` this `hiring_agency_name` has in the larger data.

```{r}
prep_dedup %>% 
  filter(hiring_agency_name == "Public Defender Service for the District of Columbia") %>% 
  count(hiring_agency_name,hiring_department_name)
```

This suggests we should replace `hiring_department_name` this `hiring_agency_name`.

Let's do both steps.

```{r}
blanks_03 <- 
  blanks_01 %>% 
    select(starts_with("hiring_agency_"),starts_with("hiring_department_name"),everything()) %>%
  mutate(
    hiring_agency_name = case_when(
      hiring_agency_name == "" & !is.na(hiring_department_name) ~ hiring_department_name,
      TRUE ~ hiring_agency_name
      ),
    hiring_department_name = case_when(
      hiring_department_name == "" & !is.na(hiring_agency_name) ~ hiring_agency_name,
      TRUE ~ hiring_department_name
      ),
    ) 
```

Reviewing blanks.

```{r}
  blanks_03 %>% 
  filter(if_any(all_of(blanking_variable %>% pull(variable)),  ~.x == "")) %>% 
  select(all_of(blanking_variable %>% pull(variable)),everything())
```

Reduced a lot.

Let's count the extent of the issues.

```{r}
blanking_variable %>% 
  pull(variable) %>% 
  map(
    ~blanks_03 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(!!sym(.x) == "")
  )
```

Still have many issues with the `hiring_agency_code`. `hiring_agency_name` appears to have all of the important information and is more complete. So we will drop `hiring_agency_code`.

Now apply the steps above to the larger dataframe.

```{r}
prep_xx <- 
  prep_dedup %>% 
  mutate(
    hiring_agency_name = case_when(
      hiring_agency_name == "" & !is.na(hiring_department_name) ~ hiring_department_name,
      TRUE ~ hiring_agency_name
      ),
    hiring_department_name = case_when(
      hiring_department_name == "" & !is.na(hiring_agency_name) ~ hiring_agency_name,
      TRUE ~ hiring_department_name
      ),
    ) %>% 
  select(-hiring_agency_code)
```

Check for blanks.

```{r}
prep_xx %>% 
  filter(if_any(all_of(blanking_variable %>% filter(variable != 'hiring_agency_code') %>%  pull(variable)),  ~.x == ""))
```

Rename dataframe

```{r}
prep_blank <- prep_xx
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing blanks produced by:

```{r}
blanking_variable <- 
  blanking_variables %>% 
  filter(group == blanking_groups[step_n])

blanking_variable
```

Reviewing blanks.

```{r}
blanks_01 <- 
  prep_dedup %>% 
  filter(if_any(all_of(blanking_variable %>% pull(variable)),  ~.x == "")) %>% 
  select(all_of(blanking_variable %>% pull(variable)),everything())

blanks_01
```

Looks like `who_may_apply` is just blank.

Let's count the extent of the issues.

```{r}
blanking_variable %>% 
  pull(variable) %>% 
  map(
    ~blanks_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(!!sym(.x) == "")
  )
```

They are all blank. We will drop the `who_may_apply` column.

```{r}
prep_xx <- 
  prep_blank %>% 
  select(-who_may_apply)
```

Rename dataframe

```{r}
prep_blank <- prep_xx
```

```{r}
step_n <- step_n + 1
```

### Check that blanks are removed

Created a dataframe of variable metadata for deduplicated data

```{r}
column_meta_data_blank <- 
  prep_blank %>% 
  contents()

column_meta_data_blank <- 
  column_meta_data_blank$contents %>% 
  as.data.frame()

column_meta_data_blank <-
  column_meta_data_blank %>% 
  rownames_to_column(var = "variable")

column_meta_data_blank
```

Checking duplicates for each variable in the dataframe.

```{r}
blanks_by_variable <- 
  column_meta_data_blank %>% 
  pull(variable) %>% 
  map_lgl(~blank_check(in_data = prep_blank,.))

blanks_by_variable
```

Add this to the `column_meta_data_dedupe` dataframe.

```{r}
column_meta_data_blank <- 
  column_meta_data_blank %>% 
  add_column(
    blanks_by_variable = blanks_by_variable
  )

column_meta_data_blank
```

Create dataframe of only variables that have blanks.

```{r}
blanking_variables <- 
  column_meta_data_blank %>% 
  filter(blanks_by_variable)

blanking_variables
```

## Addressing nulls

Created a dataframe of variable metadata for blanks data

```{r}
column_meta_data_null <- 
  prep_blank %>% 
  contents()

column_meta_data_null <- 
  column_meta_data_null$contents %>% 
  as.data.frame()

column_meta_data_null <-
  column_meta_data_null %>% 
  rownames_to_column(var = "variable")

column_meta_data_null
```

Function that checks whether there are duplicates for different variables

```{r}
null_check <- function(in_data,in_variable_name)
{
  check_01 <- 
  in_data %>% 
    filter(is.na(!!sym(in_variable_name))) %>% 
  distinct()

out <- nrow(check_01) > 0

return(out)
}
```

Checking duplicates for each variable in the dataframe.

```{r}
nulls_by_variable <- 
  column_meta_data_null %>% 
  pull(variable) %>% 
  map_lgl(~null_check(in_data = prep_dedup,.))

nulls_by_variable
```

Add this to the `column_meta_data_null` dataframe.

```{r}
column_meta_data_null <- 
  column_meta_data_null %>% 
  add_column(
    nulls_by_variable = nulls_by_variable
  )

column_meta_data_null
```

Create dataframe of only variables that have blanks.

```{r}
null_variables <- 
  column_meta_data_null %>% 
  filter(nulls_by_variable)

null_variables
```

Grouping variables.

```{r}
null_variables <- 
  null_variables %>% 
  mutate(group = 
           case_when(variable == "position_open_date" ~ "dates",
                     variable == "position_close_date" ~ "dates",
                     variable == "announcement_closing_type" ~ "announcement_closing_type",
                     variable == "announcement_closing_type_description" ~ "announcement_closing_type",
                     variable == "application_limit" ~ "application_limit",
                     variable == "application_limit_set" ~ "application_limit",
                     variable == "agency_level" ~ "agency_level",
                     variable == "agency_level_sort" ~ "agency_level",
                     variable == "hiring_agency_code" ~ "agency_level",
                     variable == "hiring_agency_name" ~ "agency_level",
                     variable == "hiring_department_code" ~ "agency_level",
                     variable == "hiring_department_name" ~ "agency_level",
                     variable == "maximum_salary" ~ "maximum_salary",
                     variable == "minimum_salary" ~ "maximum_salary",
                     variable == "maximum_grade" ~ "maximum_salary",
                     variable == "minimum_grade" ~ "maximum_salary",
                     variable == "series" ~ "series",
                     variable == "series_title" ~ "series",
                     TRUE ~ variable
                     ))

null_variables
```

Create list of groups.

```{r}
null_groups <- 
  null_variables %>% 
  select(group) %>% 
  distinct() %>% 
  pull()

null_groups
```

### Nulls removal

```{r}
step_n <- 0
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing nulls produced by:

```{r}
null_variable <- 
  null_variables %>% 
  filter(group == null_groups[step_n])

null_variable
```

Reviewing blanks.

```{r}
null_01 <- 
  prep_blank %>% 
  filter(if_any(all_of(null_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(null_variable %>% pull(variable)),everything())

null_01
```

Looks like the `hiring_department_code` is null for some rows.

Let's count the extent of the issues.

```{r}
null_variable %>% 
  pull(variable) %>% 
  map(
    ~null_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's see how the hiring codes look in comparison to the department names.

```{r}
prep_blank %>% 
  count(hiring_department_code,hiring_department_name)
```

There are a lot of nulls, and 0s.

Check for any duplications in the codes.

```{r}
prep_blank %>% 
  count(hiring_department_code,hiring_department_name) %>% 
  get_dupes(hiring_department_code)
```

Lots of 0, nulls, and 5s.

The department name appears more complete. We will drop `hiring_department_code`.

```{r}
prep_xx <- 
  prep_blank %>% 
  select(-hiring_department_code)
```

Rename dataframe

```{r}
prep_null <- prep_xx
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing nulls produced by:

```{r}
null_variable <- 
  null_variables %>% 
  filter(group == null_groups[step_n])

null_variable
```

Reviewing blanks.

```{r}
null_01 <- 
  prep_blank %>% 
  filter(if_any(all_of(null_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(null_variable %>% pull(variable)),everything())

null_01
```

Looks like the `telework_eligibility_description` is null.

Let's count the extent of the issues.

```{r}
null_variable %>% 
  pull(variable) %>% 
  map(
    ~null_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

The whole column is null. We should drop.

```{r}
prep_xx <- 
  prep_null %>% 
  select(-telework_eligibility_description)
```

Rename dataframe

```{r}
prep_null <- prep_xx
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing nulls produced by:

```{r}
null_variable <- 
  null_variables %>% 
  filter(group == null_groups[step_n])

null_variable
```

Reviewing blanks.

```{r}
null_01 <- 
  prep_blank %>% 
  filter(if_any(all_of(null_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(null_variable %>% pull(variable)),everything())

null_01
```

Looks like the `second_announcement` is null.

Let's count the extent of the issues.

```{r}
null_variable %>% 
  pull(variable) %>% 
  map(
    ~null_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

The whole column is null. We should drop.

```{r}
prep_xx <- 
  prep_null %>% 
  select(-second_announcement)
```

Rename dataframe

```{r}
prep_null <- prep_xx
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing nulls produced by:

```{r}
null_variable <- 
  null_variables %>% 
  filter(group == null_groups[step_n])

null_variable
```

Reviewing blanks.

```{r}
null_01 <- 
  prep_blank %>% 
  filter(if_any(all_of(null_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(null_variable %>% pull(variable)),everything())

null_01
```

Looks like the `pay_scale` is null.

Let's count the extent of the issues.

```{r}
null_variable %>% 
  pull(variable) %>% 
  map(
    ~null_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's see how the `pay_scale` look in comparison to the `pay_scale_description`.

```{r}
prep_blank %>% 
  count(pay_scale,pay_scale_description)
```

Looks like the NAs in `pay_scale` are "Nonappropriated Funds--Nonsupervisory And Nonleader--Federal Wage System." We'll retain `pay_scale_description`and drop `pay_scale`.

```{r}
prep_xx <- 
  prep_null %>% 
  select(-pay_scale)
```

Rename dataframe

```{r}
prep_null <- prep_xx
```

```{r}
step_n <- step_n + 1
```

### Check that nulls are removed

Created a dataframe of variable metadata for blanks data

```{r}
column_meta_data_null <- 
  prep_null %>% 
  contents()

column_meta_data_null <- 
  column_meta_data_null$contents %>% 
  as.data.frame()

column_meta_data_null <-
  column_meta_data_null %>% 
  rownames_to_column(var = "variable")

column_meta_data_null
```

Checking duplicates for each variable in the dataframe.

```{r}
nulls_by_variable <- 
  column_meta_data_null %>% 
  pull(variable) %>% 
  map_lgl(~null_check(in_data = prep_null,.))

nulls_by_variable
```

Add this to the `column_meta_data_null` dataframe.

```{r}
column_meta_data_null <- 
  column_meta_data_null %>% 
  add_column(
    nulls_by_variable = nulls_by_variable
  )

column_meta_data_null
```

Create dataframe of only variables that have blanks.

```{r}
null_variables <- 
  column_meta_data_null %>% 
  filter(nulls_by_variable)

null_variables
```

## Other data cleaning

Review the data.

```{r}
prep_null %>% 
  skim()
```

Only two variables stand out: `telework_eligible` and `supervisory_status`.

First count them.

```{r}
prep_null %>% 
  count(telework_eligible,supervisory_status)
```

No positions appear to be telework eligible, which is wrong.

<!-- Talk to Melissa about this. -->

For supervisory status, it's preferable to transform this into an indicator.

```{r}
prep_xx <- 
  prep_null %>% 
  character_to_indicator(params$unique_identifier_1,supervisory_status,"supervisory_status") %>% 
  select(-supervisory_status_n)

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Rename dataframe.

```{r}
prep_12 <- prep_xx
```

## Creating final data set

Remove `supervisory_status` and attach `supervisory_status_y` the new data

```{r}
prep_xx <- 
  prep_null %>% 
  select(-supervisory_status) %>% 
  left_join(prep_12)

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Rename dataframe.

```{r}
joas <- prep_xx
```

## Save data {.tabset .tabset-pills}

```{r}
save(joas, 
     file = here::here("03. Data Collection",
                       "prepared-data",
                       glue::glue("joas-{today()}.rdata")))
```

## Extraction time {.tabset .tabset-pills}

```{r}
tictoc::toc()
```