### 2023-03-15 04. Analysis/scripts/01_data_preparation.Rmd
Update 01_data_preparation.Rmd
Added headers, include section comparing combined dataframe to originals, began to address nulls and blanks.

---
title: "Data preparation"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    code_folding: hide
params:
  author: "Ben Jaques-Leslie"
  project_number: 2312
  project_name: "Occupancy Patterns"
  google_drive_home: "G:/Shared drives/MSG Projects/1.0 Real Estate Solutions (LQ1)/2312 Work Patterns"
  data_folder_1: "Occupancy Data Pull"
  unique_identifier_1: !r c("employee_name","date")
  data_folder_2: "Rosters"
  unique_identifier_2: !r c("snapshot_date","employee_e_mail_address")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!("arsenal" %in% installed.packages()[,"Package"]))
{
  install.packages("arsenal")
}
if(!("oesrrr" %in% installed.packages()[,"Package"]))
{
  devtools::install_github(repo = "GSA/oesrrr")
}
library(oesrrr)
library(tidyverse)
library(janitor)
library(skimr)
library(DataExplorer)
library(readr)
library(arsenal)
library(Hmisc)
library(readxl)
library(flextable)
library(ggthemes)
```

# Data preparation {.tabset .tabset-pills}

**Project number**: `r params$project_number`

**Project name**: `r params$project_name`

**Author**: `r params$author`

**Data folder 1**: `r params$data_folder_`'

**Unique identifier for data folder 1**: `r params$unique_identifier_1`

## Load data {.tabset .tabset-pills}

### Data files

Find the list of .csv files in the `r params$data_folder_1` data folder.

```{r}
files_1 <- 
  list.files(file.path(params$google_drive_home,"03. Data Collection",params$data_folder_1), pattern = ".csv", full.names = TRUE)
files_1
```

Find the list of .csv files in the `r params$data_folder_2` data folder.

```{r}
files_2 <- 
  list.files(file.path(params$google_drive_home,"03. Data Collection",params$data_folder_2), pattern = ".xlsx", full.names = TRUE)
files_2
```

### Load data

Load all data files in the `r params$data_folder_1` data folder.

```{r}
prep_base_1 <- files_1 %>% 
  map_dfr(~mutate(.data = read_csv(., 
    col_types = cols(Date = col_date(format = "%m_%d_%Y"))), 
        file_name = str_remove(.,pattern = file.path(params$google_drive_home,"03. Data Collection"))
      )) %>% 
  clean_names() %>% 
  distinct() %>% 
  separate(supervisor_name,into = c("sup_last","sup_first"), sep = ",") %>% 
  mutate(
    across(
      .cols = starts_with("sup_"),
      .fns = ~word(.,1)
    )
  ) %>% 
  unite(col = "supervisor_display_name",sup_first,sup_last, sep = " ") %>% 
  rename(file_name_1 =file_name)%>% 
  mutate(
    date_month = floor_date(date,unit = "month"),
    date_month_rollback = rollback(date_month),
    date_month_forward = rollforward(date_month)
  )
```

Review.

```{r}
prep_base_1 %>% 
  skim()
```

Count number of unique employee_name.

```{r}
unique_name_base <- prep_base_1 %>% summarise(n_distinct(employee_name)) %>% pull()
unique_name_base
```

```{r}
prep_base_1_names <- names(prep_base_1)
```

Load all data files in the `r params$data_folder_2` data folder.

```{r}
prep_base_2 <- files_2 %>% 
  map_dfr(~mutate(.data = read_xlsx(path = ., 
                                    col_types = c("date", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "date", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text")), 
        file_name = str_remove(.,pattern = file.path(params$google_drive_home,"03. Data Collection"))
      )) %>% 
  clean_names() %>% 
  distinct() %>% 
  mutate(
    across(
      .cols = where(is.POSIXct),
      .fns = ~as_date(.)
    )
  )
```

Review.

```{r}
prep_base_2 %>% 
  skim()
```

Count number of unique email addresses.

```{r}
unique_email_base <- prep_base_2 %>% summarise(n_distinct(employee_e_mail_address)) %>% pull()
unique_email_base
```

## Joining data

### Join 1

First inner join by employee name, supervisor name, department id, and the last date, using the last day of the month before the daily checkin.

```{r}
prep_xx <- prep_base_1 %>% 
  inner_join(prep_base_2,
            by =  c("employee_name" = "full_name",
                    "supervisor_display_name" = "supervisor_display_name",
                    "dept_id" = "department_id",
                    "date_month_rollback" = "snapshot_date"
                    ),
            suffix = c("_daily","_roster")
  ) %>% 
  select(-starts_with("file_name")) %>% 
  distinct()
```

Review.

```{r}
prep_xx %>% 
  skim()
```

```{r}
prep_xx %>%
  get_dupes(date,employee_e_mail_address) %>%
  arrange(desc(dupe_count))
```

Create building dataframe.

```{r}
prep_building <- prep_xx
```

Count unique email addresses in building dataframe.

```{r}
unique_email_building <- prep_building %>% summarise(n_distinct(employee_e_mail_address)) %>% pull()
unique_email_building
```

Test: This does not differ from the base.

```{r}
unique_email_base == unique_email_building
```
Review.

```{r}
prep_building %>%  skim()
```

### Join 2

Second, remove data from building dataframe, then inner join by employee name, supervisor name, department id, and the last date, using the last day of the month after the daily checkin.

```{r}
prep_xx <- 
  prep_base_1 %>% 
  anti_join(prep_building) %>% 
  inner_join(prep_base_2,
            by =  c("employee_name" = "full_name",
                    "supervisor_display_name" = "supervisor_display_name",
                    "dept_id" = "department_id",
                    "date_month_forward" = "snapshot_date"
                    ),
            suffix = c("_daily","_roster")
  ) %>% 
  select(-starts_with("file_name")) %>% 
  distinct()
```

Review.

```{r}
prep_xx %>%  skim()
```

```{r}
prep_xx %>%
  get_dupes(date,employee_e_mail_address) %>%
  arrange(desc(dupe_count))
```

Bind to buidling dataframe.

```{r}
prep_building <- prep_building %>% 
  bind_rows(prep_xx) %>% 
  distinct()
```

Count unique email addresses

```{r}
unique_email_building <- prep_building %>% summarise(n_distinct(employee_e_mail_address)) %>% pull()
unique_email_building
```

Test: This does not differ from the base.

```{r}
unique_email_base == unique_email_building
```

Review.

```{r}
prep_building %>%  skim()
```

### Create unique person id

```{r}
prep_unique_id <- 
  prep_building %>% 
  select(employee_name,employee_e_mail_address) %>% 
  distinct() %>% 
  rownames_to_column(var = "person_id")
```

Review.

```{r}
prep_unique_id %>% 
  skim()
```

### Join to building dataframe

```{r}
prep_building <- 
  prep_building %>% 
  left_join(prep_unique_id) %>% 
  select(person_id,everything())
```

Review.

```{r}
prep_building %>% 
  skim()
```

Check for duplicates on `person_id` and `date`.

```{r}
prep_building %>% 
  get_dupes(person_id,date)
```

## Compare to original dataframes

```{r}
prep_base_1_summary <- 
  prep_base_1 %>% 
  unite(employee_name_date,employee_name,date,remove = FALSE) %>% 
  summarise(
    rows = n(),
    employee_name = n_distinct(employee_name),
    employee_name_date = n_distinct(employee_name_date),
  ) %>% 
  mutate(dataframe = "prep_base_1")
```

```{r}
prep_base_2_summary <- 
  prep_base_2 %>% 
  # unite(employee_name_date,employee_name,date,remove = FALSE) %>% 
  summarise(
    rows = n(),
    employee_name = n_distinct(full_name),
    employee_e_mail_address = n_distinct(employee_e_mail_address)
  ) %>% 
  mutate(dataframe = "prep_base_2")
```

```{r}
prep_building_summary <- 
  prep_building %>% 
  unite(employee_name_date,employee_name,date,remove = FALSE) %>%
  summarise(
    rows = n(),
    person_id = n_distinct(person_id),
    employee_name = n_distinct(employee_name),
    employee_e_mail_address = n_distinct(employee_e_mail_address),
    employee_name_date = n_distinct(employee_name_date),
  ) %>% 
  mutate(dataframe = "prep_building")
```

```{r}
prep_summary <- 
  prep_base_1_summary %>% 
  bind_rows(prep_base_2_summary) %>% 
  bind_rows(prep_building_summary) %>% 
  pivot_longer(-dataframe) %>% 
  pivot_wider(names_from = dataframe,values_from = value)
```

```{r}
prep_summary %>% 
  flextable() %>% 
  theme_zebra() %>% 
  autofit()
```

```{r}
prep_base_1_rows <- prep_summary %>% filter(name == 'rows') %>% pull(prep_base_1)

prep_base_1_employee_name <- prep_summary %>% filter(name == 'employee_name') %>% pull(prep_base_1)

diff_rows_count <- 
  (prep_base_1_rows) -
  (prep_summary %>% filter(name == 'rows') %>% pull(prep_building))

diff_rows_prop <- 
  diff_rows_count/prep_base_1_rows

diff_employee_name_count <- 
  (prep_base_1_employee_name) -
  (prep_summary %>% filter(name == 'employee_name') %>% pull(prep_building))

diff_employee_name_prop <- 
  diff_employee_name_count/prep_base_1_employee_name
```

From this table, we find:

1.    A loss of `r scales::comma(diff_rows_count)` rows between `prep_1_base` and `prep_building`, about `r scales::percent(diff_rows_prop)`
2.    A loss of `r scales::comma(diff_employee_name_count)` rows between `prep_1_base` and `prep_building`, about `r scales::percent(diff_employee_name_prop)`

These losses of data are reasonable.

Create a unique dataframe of just `date` and `person_id`.

```{r}
prep_uniq <- 
  prep_building %>% 
  select(date,person_id)
```

## Addressing blanks, nulls, and other values

Created a dataframe of variable metadata.

```{r}
column_meta_data <- 
  prep_building %>% 
  contents()

column_meta_data <- 
  column_meta_data$contents %>% 
  as.data.frame()

column_meta_data <-
  column_meta_data %>% 
  rownames_to_column(var = "variable")

column_meta_data <- 
  column_meta_data %>% 
  mutate(
    nulls_by_variable = case_when(NAs > 0 ~ TRUE,
                                  NAs == 0 ~ FALSE)
  )

column_meta_data
```

Counting top 10 values for all character or factor columns to identify other common values.

```{r}
values_count <- 
  prep_building %>% 
  select(where(is.character) | where(is.factor)) %>% 
  names()

values_count %>% 
  map(~count(prep_building,!!sym(.)) %>% 
  arrange(desc(n)) %>%
  slice_head(n = 10))

```

No obvious values.

Function that checks whether there are blanks for different variables

```{r}
blank_check <- function(in_data,in_variable_name)
{
  check_01 <- 
  in_data %>% 
    filter(!!sym(in_variable_name) == "") %>% 
  distinct()

out <- nrow(check_01) > 0

return(out)
}
```

Checking blanks for each variable in the dataframe.

```{r}
blanks_by_variable <- 
  column_meta_data %>% 
  pull(variable) %>% 
  map_lgl(~blank_check(in_data = prep_building,.))

blanks_by_variable
```

Add this to the `column_meta_data` dataframe.

```{r}
column_meta_data <- 
  column_meta_data %>% 
  add_column(
    blanks_by_variable = blanks_by_variable
  )

column_meta_data
```

Create dataframe of only variables that have blanks.

```{r}
problem_variables <- 
  column_meta_data %>% 
  filter(blanks_by_variable | nulls_by_variable)

problem_variables
```

Grouping variables.

```{r}
problem_variables <- 
  problem_variables %>% 
  mutate(group = 
           case_when(variable == "position_category_code" ~ "position_category_code",
                     variable == "position_category_description" ~ "position_category_code",
                     TRUE ~ variable
                     ))

problem_variables
```

Create list of groups.

```{r}
problem_groups <- 
  problem_variables %>% 
  select(group) %>% 
  distinct() %>% 
  arrange(group) %>% 
  pull()

problem_groups
```
Also create a building dataframe to add other dataframes where missing have been removed.

```{r}
prep_problem <- prep_uniq
```

### Problems removal

```{r}
step_n <- 0
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `business_line_2_letter_department_id`

Removing blanks produced by

```{r}
problem_variable <- 
  problem_variables %>% 
  filter(group == problem_groups[step_n])

problem_variable
```

Reviewing blanks.

```{r}
nulls_01 <- 
  prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(problem_variable %>% pull(variable)),everything())

nulls_01
```

Let's count the extent of the issues.

```{r}
problem_variable %>% 
  pull(variable) %>% 
  map(
    ~nulls_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's count the entries here.

```{r}
prep_building %>% 
  group_by(across(all_of(problem_variable %>% pull(variable)))) %>% 
  count() %>% 
  arrange(desc(n))
```

Lot's of entries.

Check relationship with time.

```{r}
prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  group_by(across(all_of(problem_variable %>% pull(variable))),date) %>% 
  count() %>% 
  ggplot(aes(x = date, y = n)) +
  geom_point() +
  geom_smooth()
```

Increase in July.

Is this an important variable?

No. We'll replace nulls with "(Missing)".

```{r}
prep_xx <- 
  prep_building %>% 
  select(date, person_id,all_of(problem_variable %>% pull(variable))) %>% 
  distinct() %>% 
  mutate(
    across(
      .cols = all_of(problem_variable %>% pull(variable)),
      .fns = ~case_when(is.na(.) ~ "(Missing)",
                        TRUE ~ .)
    )
  )

prep_xx %>% skim()
```

Join to building dataframe.

```{r}
prep_problem <- prep_problem %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `employee_e_mail_address`

Removing problems produced by

```{r}
problem_variable <- 
  problem_variables %>% 
  filter(group == problem_groups[step_n])

problem_variable
```

Reviewing blanks.

```{r}
nulls_01 <- 
  prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(problem_variable %>% pull(variable)),everything())

nulls_01
```

Let's count the extent of the issues.

```{r}
problem_variable %>% 
  pull(variable) %>% 
  map(
    ~nulls_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's count the entries here.

```{r}
prep_building %>% 
  group_by(across(all_of(problem_variable %>% pull(variable)))) %>% 
  count() %>% 
  arrange(desc(n))
```

Lot's of entries.

Check relationship with time.

```{r}
prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  group_by(across(all_of(problem_variable %>% pull(variable))),date) %>% 
  count() %>% 
  ggplot(aes(x = date, y = n)) +
  geom_jitter() +
  geom_smooth() +
  theme_fivethirtyeight()
```

Looks pretty random.

Is this an important variable?

Yes, but we won't be using it directly in the analysis. We'll replace nulls with "(Missing)".

```{r}
prep_xx <- 
  prep_building %>% 
  select(date, person_id,all_of(problem_variable %>% pull(variable))) %>% 
  distinct() %>% 
  mutate(
    across(
      .cols = all_of(problem_variable %>% pull(variable)),
      .fns = ~case_when(is.na(.) ~ "(Missing)",
                        TRUE ~ .)
    )
  )

prep_xx %>% skim()
```

Join to building dataframe.

```{r}
prep_problem <- prep_problem %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `location_city`

Removing problems produced by

```{r}
problem_variable <- 
  problem_variables %>% 
  filter(group == problem_groups[step_n])

problem_variable
```

Reviewing blanks.

```{r}
nulls_01 <- 
  prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(problem_variable %>% pull(variable)),everything())

nulls_01
```

Let's count the extent of the issues.

```{r}
problem_variable %>% 
  pull(variable) %>% 
  map(
    ~nulls_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's count the entries here.

```{r}
prep_building %>% 
  group_by(across(all_of(problem_variable %>% pull(variable)))) %>% 
  count() %>% 
  arrange(desc(n))
```

Lot's of entries.

Check relationship with time.

```{r}
prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  group_by(across(all_of(problem_variable %>% pull(variable))),date) %>% 
  count() %>% 
  ggplot(aes(x = date, y = n)) +
  geom_jitter() +
  geom_smooth() +
  theme_fivethirtyeight()
```

Weird that early in the year is low, then gets steady later.

Is this an important variable?

Yes, but we have other location variables to rely on. We'll replace nulls with "(Missing)".

```{r}
prep_xx <- 
  prep_building %>% 
  select(date, person_id,all_of(problem_variable %>% pull(variable))) %>% 
  distinct() %>% 
  mutate(
    across(
      .cols = all_of(problem_variable %>% pull(variable)),
      .fns = ~case_when(is.na(.) ~ "(Missing)",
                        TRUE ~ .)
    )
  )

prep_xx %>% skim()
```

Join to building dataframe.

```{r}
prep_problem <- prep_problem %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `location_state`

Removing problems produced by

```{r}
problem_variable <- 
  problem_variables %>% 
  filter(group == problem_groups[step_n])

problem_variable
```

Reviewing blanks.

```{r}
nulls_01 <- 
  prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(problem_variable %>% pull(variable)),everything())

nulls_01
```

Let's count the extent of the issues.

```{r}
problem_variable %>% 
  pull(variable) %>% 
  map(
    ~nulls_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's count the entries here.

```{r}
prep_building %>% 
  group_by(across(all_of(problem_variable %>% pull(variable)))) %>% 
  count() %>% 
  arrange(desc(n))
```

Lot's of entries.

Check relationship with time.

```{r}
prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  group_by(across(all_of(problem_variable %>% pull(variable))),date) %>% 
  count() %>% 
  ggplot(aes(x = date, y = n)) +
  geom_jitter() +
  geom_smooth() +
  theme_fivethirtyeight()
```

Looks pretty random, and low numbers.

Is this an important variable?

Yes, but we have other location variables to rely on. It appears from above that these are international locations. We'll replace nulls with "(International)".

```{r}
prep_xx <- 
  prep_building %>% 
  select(date, person_id,all_of(problem_variable %>% pull(variable))) %>% 
  distinct() %>% 
  mutate(
    across(
      .cols = all_of(problem_variable %>% pull(variable)),
      .fns = ~case_when(is.na(.) ~ "(International)",
                        TRUE ~ .)
    )
  )

prep_xx %>% skim()
```

Join to building dataframe.

```{r}
prep_problem <- prep_problem %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

# CONTINUE FROM HERE.

Third, remove data from building dataframe, then inner join by employee name.

```{r}
prep_xx <- 
  prep_base_1 %>% 
  anti_join(prep_building) %>%
  inner_join(prep_base_2,
            by =  c("employee_name" = "full_name"
                    ),
            suffix = c("_daily","_roster")
  ) %>% 
  select(-starts_with("file_name")) %>% 
  distinct()
```

Review.

```{r}
prep_xx %>%  skim()
```

Bind to building dataframe.

```{r}
prep_building <- prep_building %>% 
  bind_rows(prep_xx) %>% 
  distinct()
```

```{r}
unique_email_building <- prep_building %>% summarise(n_distinct(employee_e_mail_address)) %>% pull()
unique_email_building
```

Test: This does not differ from the base.

```{r}
unique_email_base == unique_email_building
```

```{r}
prep_building %>%  skim()
```

Load all data files in the `r params$data_folder_2` data folder.

```{r}
prep_base_2_1 <- files_2[1] %>% 
  map_dfr(~mutate(.data = read_xlsx(path = ., 
                                    col_types = c("date", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "date", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text")), 
        file_name = str_remove(.,pattern = file.path(params$google_drive_home,"03. Data Collection"))
      )) %>% 
  clean_names() %>% 
  distinct()
```

```{r}
prep_base_2_2 <- files_2[2] %>% 
  map_dfr(~mutate(.data = read_xlsx(path = ., 
                                    col_types = c("date", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "date", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text")), 
        file_name = str_remove(.,pattern = file.path(params$google_drive_home,"03. Data Collection"))
      )) %>% 
  clean_names() %>% 
  distinct()
```

```{r}
prep_base_2_3 <- files_2[3] %>% 
  map_dfr(~mutate(.data = read_xlsx(path = ., 
                                    col_types = c("date", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "date", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text")), 
        file_name = str_remove(.,pattern = file.path(params$google_drive_home,"03. Data Collection"))
      )) %>% 
  clean_names() %>% 
  distinct()
```

Joining data

```{r}
step_1 <- prep_base_1 %>% 
  # select(employee_name) %>% 
  # distinct() %>% 
   left_join(prep_base_2_1,
            by = c("employee_name" = "full_name",
                   "sso" = "gsa_sso",
                   "dept_id" = "department_id",
                   "supervisor_display_name" = "supervisor_display_name"
                   )) %>% 
  # filter(rollback(date) == snapshot_date |
  #          rollforward(date) == snapshot_date) %>%
  # select(-snapshot_date, -starts_with("file_name")) %>% 
  distinct()
  # skim()
  # get_dupes(employee_e_mail_address,date)
```

```{r}
step_1 %>% 
  filter(!is.na(employee_e_mail_address)) %>% 
  get_dupes(employee_name,date,sso, dept_id, supervisor_display_name)
```

```{r}
step_1 %>% 
  filter(!is.na(employee_e_mail_address)) %>% 
  get_dupes(employee_e_mail_address,date)
```

```{r}
step_2 <- 
  step_1 %>% 
  filter(!is.na(employee_e_mail_address)) %>% 
  get_dupes(employee_name) %>% 
  select(employee_name) %>% 
  distinct()
```

```{r}
step_3 <- 
  step_2 %>% 
  inner_join(prep_base_1 %>% 
  select(employee_name,sso) %>% 
    distinct()) %>% 
  left_join(prep_base_2_1,
            by = c("employee_name" = "full_name",
                   "sso" = "gsa_sso"
                   # ,
                   # "dept_id" = "department_id",
                   # "supervisor_display_name" = "supervisor_display_name"
                   )) %>% 
  # filter(rollback(date) == snapshot_date |
  #          rollforward(date) == snapshot_date) %>%
  # select(-snapshot_date, -starts_with("file_name")) %>% 
  distinct()
```

```{r}
step_3 %>% 
  filter(!is.na(employee_e_mail_address)) %>% 
  get_dupes(employee_name,employee_e_mail_address)
```

```{r}
step_2 <- 
  step_1 %>% 
  filter(is.na(employee_e_mail_address)) %>% 
  select(all_of(prep_base_1_names)) %>% 
  left_join(prep_base_2_2,
            by = c("employee_name" = "full_name",
                   "sso" = "gsa_sso",
                   "dept_id" = "department_id",
                   "supervisor_display_name" = "supervisor_display_name"
                   )) %>% 
  # filter(rollback(date) == snapshot_date |
  #          rollforward(date) == snapshot_date) %>%
  # select(-snapshot_date, -starts_with("file_name")) %>% 
  distinct()
```

```{r}
step_3 <- 
  step_2 %>% 
  filter(is.na(employee_e_mail_address)) %>% 
  select(all_of(prep_base_1_names)) %>% 
  left_join(prep_base_2_3,
            by = c("employee_name" = "full_name",
                   "sso" = "gsa_sso",
                   "dept_id" = "department_id",
                   "supervisor_display_name" = "supervisor_display_name"
                   )) %>% 
  # filter(rollback(date) == snapshot_date |
  #          rollforward(date) == snapshot_date) %>%
  # select(-snapshot_date, -starts_with("file_name")) %>% 
  distinct()
```

```{r}
step_3 %>% 
  filter(!is.na(employee_e_mail_address))
```

```{r}
rollback(ymd('2022-12-01'))
rollforward(ymd('2022-12-01'))
```

### Review data

```{r}
prep_base %>% skim()
```

## Addressing duplication

We are using `r params$unique_identifier_1` as the unique identifier.

### Check for duplication

```{r}
dupe_01 <- 
  prep_base %>%
  get_dupes(params$unique_identifier_1)
```

#### Data review

```{r}
dupe_01 %>% 
  skim()
```

### Dataframe of the unique identifier

```{r}
prep_uniq <- 
  prep_base %>% 
  select(params$unique_identifier_1) %>% 
  distinct()

prep_uniq %>% 
  count()
```

### Identifying and categorizing columns

Created a dataframe of variable metadata

```{r}
column_meta_data <- 
  prep_base %>% 
  contents()

column_meta_data <- 
  column_meta_data$contents %>% 
  as.data.frame()

column_meta_data <-
  column_meta_data %>% 
  rownames_to_column(var = "variable")

column_meta_data
```

Function that checks whether there are duplicates for different variables

```{r}
dupe_check <- function(in_data,in_variable_name)
{
  check_01 <- 
  in_data %>% 
  select(params$unique_identifier_1,all_of(in_variable_name)) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)

out <- nrow(check_01) > 0

return(out)
}
```

Checking duplicates for each variable in the dataframe.

```{r}
duplicates_by_variable <- 
  column_meta_data %>% 
  pull(variable) %>% 
  map_lgl(~dupe_check(in_data = dupe_01, .))

duplicates_by_variable
```

Add this to the column_meta_data dataframe.

```{r}
column_meta_data <- 
  column_meta_data %>% 
  add_column(
    duplicates_by_variable = duplicates_by_variable
  )

column_meta_data
```

Create dataframe of only variables that cause duplicates.

```{r}
duplicating_variables <- 
  column_meta_data %>% 
  filter(duplicates_by_variable)

duplicating_variables
```

Grouping variables.

```{r}
duplicating_variables <- 
  duplicating_variables %>% 
  mutate(group = 
           case_when(variable == "short_id" ~ "short_id",
                     variable == "dept_id" ~ "short_id",
                     variable == "supervisor_name" ~ "short_id",
                     variable == "location" ~ "short_id",
                    TRUE ~ variable
                     ))

duplicating_variables
```

Create list of groups.

```{r}
duplicating_groups <- 
  duplicating_variables %>% 
  select(group) %>% 
  distinct() %>% 
  pull()

duplicating_groups
```

### Duplication removal

```{r}
step_n <- 0
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)
```

Not too many duplicates here.

First count `sso` across the original data.

```{r}
prep_base %>% count(sso)
```

There are just a few, but we'll want to transform these into indicators.

We want to have this change to affect the whole dataframe so we address the duplicates off the original dataframe.

```{r}
prep_xx <- 
  prep_base %>% 
  character_to_indicator(params$unique_identifier_1,sso,"sso")

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Rename dataframe.

```{r}
prep_01 <- prep_xx
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)
```

Quite a few duplicates here and they are associated with supervisor and location.

First count `short_id` and `dept_id` across the original data.

```{r}
prep_base %>% count(short_id,dept_id)
```

Counting `short_id`.

```{r}
prep_base %>% count(short_id)
```

Counting `dept_id`.

```{r}
prep_base %>% count(dept_id)
```

Too many options to create useful indicators.

For lack better option, we'll select the rows with shorter ids.

```{r}
dedupe <- 
  dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  arrange(!!!params$unique_identifier_1,length(short_id),length(dept_id)) %>% 
  group_by(across(all_of(params$unique_identifier_1))) %>% 
  summarise(
    across(.fns = ~first(.))
  )

dedupe %>% 
  get_dupes(params$unique_identifier_1)
```

Creating a deduplicated dataframe of `r params$unique_identifier_1` and `r duplicating_groups[step_n]`.

First create a dataframe of distinct rows of the variables from the original data.

```{r}
working_01 <- 
  prep_base %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct()

working_01 %>% 
  glimpse()

working_01 %>% 
  skim()
```

Next remove rows that dataframe that were found to be duplicated and save that as a new dataframe.

```{r}
working_02 <- 
  working_01 %>% 
  anti_join(dedupe %>% select(params$unique_identifier_1))

working_02 %>% 
  glimpse()

working_02 %>% 
  skim()
```

Last bind this dataframe to the deduplicated set above.

```{r}
prep_xx <- 
  working_02 %>% 
  bind_rows(dedupe)

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Count rows.

```{r}
prep_xx %>% 
  count()
```

Do the number of in this dataframe match the number of unique identifiers?

```{r}
prep_xx %>% 
  count() %>% 
  pull() ==
  prep_uniq %>% 
  count() %>% 
  pull()
```

Rename dataframe

```{r}
prep_02 <- prep_xx
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`

Removing duplicates produced by:

```{r}
duplicating_group <- 
  duplicating_variables %>% 
  filter(group == duplicating_groups[step_n])
duplicating_group
```

Reviewing duplicates.

```{r}
dupe_01 %>% 
  select(params$unique_identifier_1,all_of(duplicating_group %>% pull(variable))) %>% 
  distinct() %>% 
  get_dupes(params$unique_identifier_1)
```

Create function to drop rows with null unique identifiers.

```{r}
filter_unique <- function(in_data, ...)
{
  in_data %>% 
  filter(if_all(c(...),~!is.na(.)))
}
```

Drop rows with null unique identifiers.

```{r}
prep_01 <- 
  prep_01 %>% 
  filter_unique(params$unique_identifier_1)
```

Review data.

```{r}
prep_01 %>% skim()
```

Identify columns with amounts unique values less than 30.

```{r}
low_unique_value_cols <- 
  prep_01 %>% 
  summarise(across(.fns = ~n_distinct(.))) %>% 
  rownames_to_column() %>% 
  pivot_longer(-rowname) %>% 
  filter(value < 30) %>% 
  pull(name)
```

**Unique low value columns**: `r low_unique_value_cols`

Convert low unique value variables to factors.

```{r}
prep_01 <- 
  prep_01 %>% 
  mutate(across(all_of(low_unique_value_cols),as.factor))
```

Review data.

```{r}
prep_01 %>% skim()
```

## Join/bind data {.tabset .tabset-pills}

No data to join or bind so far.

```{r}
d <- prep_01
```

Review data.

```{r}
d %>% skim()
```

## Save data {.tabset .tabset-pills}

```{r}
save(d, file = here::here("03. Data Collection",params$data_folder_1,"prepared_data.rdata"))
```