### 2023-04-19 04. Analysis/scripts/02_data_preparation_clusters_employee.Rmd
Create 02_data_preparation_clusters_employee.Rmd
initial commit

---
title: "Data preparation for employee category analysis"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    code_folding: hide
params:
  author: "Ben Jaques-Leslie"
  project_number: 2312
  project_name: "Occupancy Patterns"
  google_drive_home: "G:/Shared drives/MSG Projects/1.0 Real Estate Solutions (LQ1)/2312 Work Patterns"
  data_folder_1: "prepared-data"
  data_file_1: "check_in_panel_2023-04-12.rdata"
  unique_identifier_1: !r c("date","person_id")
  date_cohort_month: "2022-03-01"
  observation_window_length_months: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
tictoc::tic()
if(!("arsenal" %in% installed.packages()[,"Package"]))
{
  install.packages("arsenal")
}
if(!("oesrrr" %in% installed.packages()[,"Package"]))
{
  devtools::install_github(repo = "GSA/oesrrr")
}
library(oesrrr)
library(tidyverse)
library(janitor)
library(skimr)
library(DataExplorer)
library(readr)
library(arsenal)
library(Hmisc)
library(readxl)
library(flextable)
library(ggthemes)
```

```{r cohort dates}
date_cohort_month_start <- ymd(params$date_cohort_month)
date_cohort_month_end <- ceiling_date(date_cohort_month_start,unit = "month") %>% rollbackward()
date_observation_window_month_start <- date_cohort_month_start %m+% months(params$observation_window_length_months)
date_observation_window_month_end <- ceiling_date(date_observation_window_month_start,unit = "month") %>% rollbackward()
```

# Data preparation {.tabset .tabset-pills}

**Project number**: `r params$project_number`

**Project name**: `r params$project_name`

**Author**: `r params$author`

**Data folder 1**: `r params$data_folder_1`

**Data file 1**: `r params$data_file_1`

**Unique identifier for data file 1**: `r params$unique_identifier_1`

**Cohort start month**: `r nice_month_year(date_cohort_month_start)`

**Observation window**: `r nice_month_year(date_cohort_month_start)` to `r nice_month_year(date_observation_window_month_end)` (`r params$observation_window_length_months` months)

## Load data {.tabset .tabset-pills}

Load `r params$data_file_1` from `r params$data_folder_1` data folder.

```{r}
load(file = file.path(params$google_drive_home,"03. Data Collection",params$data_folder_1,params$data_file_1))
```

Review.

```{r}
check_in_panel %>% 
  skim()
```

## Creating cohort

Collect `person_id` of employees that worked in `r nice_month_year(date_cohort_month_start)`.

```{r}
cohort_ids <- 
  check_in_panel %>% 
  filter(date >= date_cohort_month_start & date <= date_cohort_month_end) %>% 
  select(person_id) %>% 
  distinct()
```

Review.

```{r}
cohort_ids %>% 
  skim()
```

Join with full panel.

```{r}
cohort_panel <- 
  cohort_ids %>% 
  inner_join(check_in_panel) %>% 
  filter(date >= date_cohort_month_start & date <= date_observation_window_month_end)
```

Review.

```{r}
cohort_panel %>% 
  skim()
```

Check: These have the same number of unique `person_id`.

```{r}
cohort_ids %>% 
  summarise(n_distinct(person_id)) %>% 
  pull() ==
  cohort_panel %>% 
  summarise(n_distinct(person_id)) %>% 
  pull()
```

Create dataframe of all initial characteristics.

```{r}
cohort_first <- 
  cohort_panel %>% 
  arrange(person_id,date) %>% 
  group_by(person_id) %>% 
  summarise(
    across(
      .cols = everything(),
      .fns = first
    )
  ) %>% 
  ungroup()
```

Review.

```{r}
cohort_first %>% 
  skim()
```

Check: These have the same number of unique `person_id`.

```{r}
cohort_ids %>% 
  summarise(n_distinct(person_id)) %>% 
  pull() ==
  cohort_first %>% 
  summarise(n_distinct(person_id)) %>% 
  pull()
````

## Creating aggregated variables

```{r}
cohort_panel %>% 
  summarise(working_days = n_distinct(date)) %>% 
  ungroup()
```

```{r}
cohort_panel %>% 
  group_by(person_id) %>% 
  filter(daily_status_imputed != "Leave") %>% 
  summarise(working_days = n_distinct(date)) %>% 
  ungroup()
```

```{r}
cohort_panel %>%
  mutate(daily_status_imputed = factor(daily_status_imputed)) %>% 
  group_by(person_id,daily_status_imputed,.drop = FALSE) %>% 
  summarise(working_days = n_distinct(date)) %>% 
  ungroup() %>% 
  pivot_wider(values_from = working_days, names_from = daily_status_imputed) %>% 
  clean_names() %>% 
  rename_with(~glue::glue("daily_status_imputed_{.}"), -person_id)
```

-   Proportion of days onsite

-   Proportion of days remote

-   Proportion of days remote by the day of the week

-   Correlation of work status between days of the week

-   Number of working days

-   Modal size of team over observation window

-   Proportion of working days co-located with at least one other team
    member

-   Proportion of working days co-located with supervisor

-   Proportion of working days co-located with entire team

-   Proportion of working days entire team is remote

Count number of unique employee_name.

```{r}
unique_name_base <- prep_base_1 %>% summarise(n_distinct(employee_name)) %>% pull()
unique_name_base
```

```{r}
prep_base_1_names <- names(prep_base_1)
```

Load all data files in the `r params$data_folder_2` data folder.

```{r}
prep_base_2 <- files_2 %>% 
  map_dfr(~mutate(.data = read_xlsx(path = ., 
                                    col_types = c("date", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "date", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text")), 
        file_name = str_remove(.,pattern = file.path(params$google_drive_home,"03. Data Collection"))
      )) %>% 
  clean_names() %>% 
  distinct() %>% 
  mutate(
    across(
      .cols = where(is.POSIXct),
      .fns = ~as_date(.)
    )
  )
```

Review.

```{r}
prep_base_2 %>% 
  skim()
```

Count number of unique email addresses.

```{r}
unique_email_base <- prep_base_2 %>% summarise(n_distinct(employee_e_mail_address)) %>% pull()
unique_email_base
```

## Joining data

### Join 1

First inner join by employee name, supervisor name, department id, and the last date, using the last day of the month before the daily checkin.

```{r}
prep_xx <- prep_base_1 %>% 
  inner_join(prep_base_2,
            by =  c("employee_name" = "full_name",
                    "supervisor_display_name" = "supervisor_display_name",
                    "dept_id" = "department_id",
                    "date_month_rollback" = "snapshot_date"
                    ),
            suffix = c("_daily","_roster")
  ) %>% 
  select(-starts_with("file_name")) %>% 
  distinct()
```

Review.

```{r}
prep_xx %>% 
  skim()
```

```{r}
prep_xx %>%
  get_dupes(date,employee_e_mail_address) %>%
  arrange(desc(dupe_count)) %>% 
  datatable_with_sampling(in_max_rows = 5000)
```

Create building dataframe.

```{r}
prep_building <- prep_xx
```

Count unique email addresses in building dataframe.

```{r}
unique_email_building <- prep_building %>% summarise(n_distinct(employee_e_mail_address)) %>% pull()
unique_email_building
```

Test: This does not differ from the base.

```{r}
unique_email_base == unique_email_building
```
Review.

```{r}
prep_building %>%  skim()
```

### Join 2

Second, remove data from building dataframe, then inner join by employee name, supervisor name, department id, and the last date, using the last day of the month after the daily checkin.

```{r}
prep_xx <- 
  prep_base_1 %>% 
  anti_join(prep_building) %>% 
  inner_join(prep_base_2,
            by =  c("employee_name" = "full_name",
                    "supervisor_display_name" = "supervisor_display_name",
                    "dept_id" = "department_id",
                    "date_month_forward" = "snapshot_date"
                    ),
            suffix = c("_daily","_roster")
  ) %>% 
  select(-starts_with("file_name")) %>% 
  distinct()
```

Review.

```{r}
prep_xx %>%  skim()
```

```{r}
prep_xx %>%
  get_dupes(date,employee_e_mail_address) %>%
  arrange(desc(dupe_count)) %>% 
  datatable_with_sampling(in_max_rows = 5000)
```

Bind to buidling dataframe.

```{r}
prep_building <- prep_building %>% 
  bind_rows(prep_xx) %>% 
  distinct()
```

Count unique email addresses

```{r}
unique_email_building <- prep_building %>% summarise(n_distinct(employee_e_mail_address)) %>% pull()
unique_email_building
```

Test: This does not differ from the base.

```{r}
unique_email_base == unique_email_building
```

Review.

```{r}
prep_building %>%  skim()
```

### Create unique person id

```{r}
prep_unique_id <- 
  prep_building %>% 
  select(employee_name,employee_e_mail_address) %>% 
  distinct() %>% 
  rownames_to_column(var = "person_id")
```

Review.

```{r}
prep_unique_id %>% 
  skim()
```

### Join to building dataframe

```{r}
prep_building <- 
  prep_building %>% 
  left_join(prep_unique_id) %>% 
  select(person_id,everything())
```

Review.

```{r}
prep_building %>% 
  skim()
```

Check for duplicates on `person_id` and `date`.

```{r}
prep_building %>% 
  get_dupes(person_id,date) %>% 
  datatable_with_sampling(in_max_rows = 5000)
```

## Compare to original dataframes

```{r}
prep_base_1_summary <- 
  prep_base_1 %>% 
  unite(employee_name_date,employee_name,date,remove = FALSE) %>% 
  summarise(
    rows = n(),
    employee_name = n_distinct(employee_name),
    employee_name_date = n_distinct(employee_name_date),
  ) %>% 
  mutate(dataframe = "prep_base_1")
```

```{r}
prep_base_2_summary <- 
  prep_base_2 %>% 
  # unite(employee_name_date,employee_name,date,remove = FALSE) %>% 
  summarise(
    rows = n(),
    employee_name = n_distinct(full_name),
    employee_e_mail_address = n_distinct(employee_e_mail_address)
  ) %>% 
  mutate(dataframe = "prep_base_2")
```

```{r}
prep_building_summary <- 
  prep_building %>% 
  unite(employee_name_date,employee_name,date,remove = FALSE) %>%
  summarise(
    rows = n(),
    person_id = n_distinct(person_id),
    employee_name = n_distinct(employee_name),
    employee_e_mail_address = n_distinct(employee_e_mail_address),
    employee_name_date = n_distinct(employee_name_date),
  ) %>% 
  mutate(dataframe = "prep_building")
```

```{r}
prep_summary <- 
  prep_base_1_summary %>% 
  bind_rows(prep_base_2_summary) %>% 
  bind_rows(prep_building_summary) %>% 
  pivot_longer(-dataframe) %>% 
  pivot_wider(names_from = dataframe,values_from = value)
```

```{r}
prep_summary %>% 
  flextable() %>% 
  theme_zebra() %>% 
  autofit()
```

```{r}
prep_base_1_rows <- prep_summary %>% filter(name == 'rows') %>% pull(prep_base_1)

prep_base_1_employee_name <- prep_summary %>% filter(name == 'employee_name') %>% pull(prep_base_1)

diff_rows_count <- 
  (prep_base_1_rows) -
  (prep_summary %>% filter(name == 'rows') %>% pull(prep_building))

diff_rows_prop <- 
  diff_rows_count/prep_base_1_rows

diff_employee_name_count <- 
  (prep_base_1_employee_name) -
  (prep_summary %>% filter(name == 'employee_name') %>% pull(prep_building))

diff_employee_name_prop <- 
  diff_employee_name_count/prep_base_1_employee_name
```

From this table, we find:

1.    A loss of `r scales::comma(diff_rows_count)` rows between `prep_1_base` and `prep_building`, about `r scales::percent(diff_rows_prop)`
2.    A loss of `r scales::comma(diff_employee_name_count)` rows between `prep_1_base` and `prep_building`, about `r scales::percent(diff_employee_name_prop)`

These losses of data are reasonable.

Create a unique dataframe of just `date` and `person_id`.

```{r}
prep_uniq <- 
  prep_building %>% 
  select(date,person_id)
```

## Addressing blanks, nulls, and other values

Created a dataframe of variable metadata.

```{r}
column_meta_data <- 
  prep_building %>% 
  contents()

column_meta_data <- 
  column_meta_data$contents %>% 
  as.data.frame()

column_meta_data <-
  column_meta_data %>% 
  rownames_to_column(var = "variable")

column_meta_data <- 
  column_meta_data %>% 
  mutate(
    nulls_by_variable = case_when(NAs > 0 ~ TRUE,
                                  NAs == 0 ~ FALSE)
  )

column_meta_data
```

Counting top 10 values for all character or factor columns to identify other common values.

```{r}
values_count <- 
  prep_building %>% 
  select(where(is.character) | where(is.factor)) %>% 
  names()

values_count %>% 
  map(~count(prep_building,!!sym(.)) %>% 
  arrange(desc(n)) %>%
  slice_head(n = 10))

```

No obvious values.

Function that checks whether there are blanks for different variables

```{r}
blank_check <- function(in_data,in_variable_name)
{
  check_01 <- 
  in_data %>% 
    filter(!!sym(in_variable_name) == "") %>% 
  distinct()

out <- nrow(check_01) > 0

return(out)
}
```

Checking blanks for each variable in the dataframe.

```{r}
blanks_by_variable <- 
  column_meta_data %>% 
  pull(variable) %>% 
  map_lgl(~blank_check(in_data = prep_building,.))

blanks_by_variable
```

Add this to the `column_meta_data` dataframe.

```{r}
column_meta_data <- 
  column_meta_data %>% 
  add_column(
    blanks_by_variable = blanks_by_variable
  )

column_meta_data
```

Create dataframe of only variables that have blanks.

```{r}
problem_variables <- 
  column_meta_data %>% 
  filter(blanks_by_variable | nulls_by_variable)

problem_variables
```

Grouping variables.

```{r}
problem_variables <- 
  problem_variables %>% 
  mutate(group = 
           case_when(variable == "position_category_code" ~ "position_category_code",
                     variable == "position_category_description" ~ "position_category_code",
                     TRUE ~ variable
                     ))

problem_variables
```

Create list of groups.

```{r}
problem_groups <- 
  problem_variables %>% 
  select(group) %>% 
  distinct() %>% 
  arrange(group) %>% 
  pull()

problem_groups
```

Also create a building dataframe to add other dataframes where missing have been removed.

```{r}
prep_problem <- prep_uniq
```

### Problems removal

```{r}
step_n <- 0
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `business_line_2_letter_department_id`

Removing blanks produced by

```{r}
problem_variable <- 
  problem_variables %>% 
  filter(group == problem_groups[step_n])

problem_variable
```

Reviewing blanks.

```{r}
nulls_01 <- 
  prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(problem_variable %>% pull(variable)),everything())

nulls_01
```

Let's count the extent of the issues.

```{r}
problem_variable %>% 
  pull(variable) %>% 
  map(
    ~nulls_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's count the entries here.

```{r}
prep_building %>% 
  group_by(across(all_of(problem_variable %>% pull(variable)))) %>% 
  count() %>% 
  arrange(desc(n))
```

Lot's of entries.

Check relationship with time.

```{r}
prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  group_by(across(all_of(problem_variable %>% pull(variable))),date) %>% 
  count() %>% 
  ggplot(aes(x = date, y = n)) +
  geom_point() +
  geom_smooth()
```

Increase in July.

Is this an important variable?

No. We'll replace nulls with "(Missing)".

```{r}
prep_xx <- 
  prep_building %>% 
  select(date, person_id,all_of(problem_variable %>% pull(variable))) %>% 
  distinct() %>% 
  mutate(
    across(
      .cols = all_of(problem_variable %>% pull(variable)),
      .fns = ~case_when(is.na(.) ~ "(Missing)",
                        TRUE ~ .)
    )
  )

prep_xx %>% skim()
```

Join to building dataframe.

```{r}
prep_problem <- prep_problem %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `employee_e_mail_address`

Removing problems produced by

```{r}
problem_variable <- 
  problem_variables %>% 
  filter(group == problem_groups[step_n])

problem_variable
```

Reviewing blanks.

```{r}
nulls_01 <- 
  prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(problem_variable %>% pull(variable)),everything())

nulls_01
```

Let's count the extent of the issues.

```{r}
problem_variable %>% 
  pull(variable) %>% 
  map(
    ~nulls_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's count the entries here.

```{r}
prep_building %>% 
  group_by(across(all_of(problem_variable %>% pull(variable)))) %>% 
  count() %>% 
  arrange(desc(n))
```

Lot's of entries.

Check relationship with time.

```{r}
prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  group_by(across(all_of(problem_variable %>% pull(variable))),date) %>% 
  count() %>% 
  ggplot(aes(x = date, y = n)) +
  geom_jitter() +
  geom_smooth() +
  theme_fivethirtyeight()
```

Looks pretty random.

Is this an important variable?

Yes, but we won't be using it directly in the analysis. We'll replace nulls with "(Missing)".

```{r}
prep_xx <- 
  prep_building %>% 
  select(date, person_id,all_of(problem_variable %>% pull(variable))) %>% 
  distinct() %>% 
  mutate(
    across(
      .cols = all_of(problem_variable %>% pull(variable)),
      .fns = ~case_when(is.na(.) ~ "(Missing)",
                        TRUE ~ .)
    )
  )

prep_xx %>% skim()
```

Join to building dataframe.

```{r}
prep_problem <- prep_problem %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `location_city`

Removing problems produced by

```{r}
problem_variable <- 
  problem_variables %>% 
  filter(group == problem_groups[step_n])

problem_variable
```

Reviewing blanks.

```{r}
nulls_01 <- 
  prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(problem_variable %>% pull(variable)),everything())

nulls_01
```

Let's count the extent of the issues.

```{r}
problem_variable %>% 
  pull(variable) %>% 
  map(
    ~nulls_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's count the entries here.

```{r}
prep_building %>% 
  group_by(across(all_of(problem_variable %>% pull(variable)))) %>% 
  count() %>% 
  arrange(desc(n))
```

Lot's of entries.

Check relationship with time.

```{r}
prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  group_by(across(all_of(problem_variable %>% pull(variable))),date) %>% 
  count() %>% 
  ggplot(aes(x = date, y = n)) +
  geom_jitter() +
  geom_smooth() +
  theme_fivethirtyeight()
```

Weird that early in the year is low, then gets steady later.

Is this an important variable?

Yes, but we have other location variables to rely on. We'll replace nulls with "(Missing)".

```{r}
prep_xx <- 
  prep_building %>% 
  select(date, person_id,all_of(problem_variable %>% pull(variable))) %>% 
  distinct() %>% 
  mutate(
    across(
      .cols = all_of(problem_variable %>% pull(variable)),
      .fns = ~case_when(is.na(.) ~ "(Missing)",
                        TRUE ~ .)
    )
  )

prep_xx %>% skim()
```

Join to building dataframe.

```{r}
prep_problem <- prep_problem %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `location_state`

Removing problems produced by

```{r}
problem_variable <- 
  problem_variables %>% 
  filter(group == problem_groups[step_n])

problem_variable
```

Reviewing blanks.

```{r}
nulls_01 <- 
  prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(problem_variable %>% pull(variable)),everything())

nulls_01
```

Let's count the extent of the issues.

```{r}
problem_variable %>% 
  pull(variable) %>% 
  map(
    ~nulls_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's count the entries here.

```{r}
prep_building %>% 
  group_by(across(all_of(problem_variable %>% pull(variable)))) %>% 
  count() %>% 
  arrange(desc(n))
```

Lot's of entries.

Check relationship with time.

```{r}
prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  group_by(across(all_of(problem_variable %>% pull(variable))),date) %>% 
  count() %>% 
  ggplot(aes(x = date, y = n)) +
  geom_jitter() +
  geom_smooth() +
  theme_fivethirtyeight()
```

Looks pretty random, and low numbers.

Is this an important variable?

Yes, but we have other location variables to rely on. It appears from above that these are international locations. We'll replace nulls with "(International)".

```{r}
prep_xx <- 
  prep_building %>% 
  select(date, person_id,all_of(problem_variable %>% pull(variable))) %>% 
  distinct() %>% 
  mutate(
    across(
      .cols = all_of(problem_variable %>% pull(variable)),
      .fns = ~case_when(is.na(.) ~ "(International)",
                        TRUE ~ .)
    )
  )

prep_xx %>% skim()
```

Join to building dataframe.

```{r}
prep_problem <- prep_problem %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `position_category_code`

Removing problems produced by

```{r}
problem_variable <- 
  problem_variables %>% 
  filter(group == problem_groups[step_n])

problem_variable
```

Reviewing blanks.

```{r}
nulls_01 <- 
  prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(problem_variable %>% pull(variable)),everything())

nulls_01
```

Let's count the extent of the issues.

```{r}
problem_variable %>% 
  pull(variable) %>% 
  map(
    ~nulls_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's count the entries here.

```{r}
prep_building %>% 
  group_by(across(all_of(problem_variable %>% pull(variable)))) %>% 
  count() %>% 
  arrange(desc(n))
```

Lot's of entries.

Check relationship with time.

```{r}
prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  group_by(across(all_of(problem_variable %>% pull(variable))),date) %>% 
  count() %>% 
  ggplot(aes(x = date, y = n)) +
  geom_jitter() +
  geom_smooth() +
  theme_fivethirtyeight()
```

Looks pretty random, and low numbers.

Let's investigate whether these employees ever have other values for `r problem_variable %>% pull(variable)`.

```{r}
prep_01 <-
  prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(person_id) %>% 
  distinct() %>% 
  inner_join(prep_building) %>% 
  arrange(person_id,date) %>% 
  select(person_id,date,all_of(problem_variable %>% pull(variable)))

prep_01 %>% 
  datatable_with_sampling(in_max_rows = 5000)
```

They do have some.

Now let's look at how they change.

```{r}
prep_02 <- 
  prep_01 %>% 
  group_by(across(all_of(problem_variable %>% pull(variable))),person_id) %>% 
  summarise(
    n = n(),
    date_first = first(date),
    date_last = last(date)
  ) %>% 
  arrange(person_id,n)

prep_02 %>% 
  datatable_with_sampling(in_max_rows = 5000)
```

It appears that most start with nulls then became a category.

Let's check their position category at the earliest date and count.

```{r}
prep_02 %>% 
  group_by(person_id) %>% 
  filter(date_first == min(date_first)) %>% 
  ungroup() %>% 
  count(position_category_code,position_category_description)
```

Most of this group are null at the beginning. Some not.

Let's check the position category at the latest date and count.

```{r}
prep_02 %>% 
  group_by(person_id) %>% 
  filter(date_first == max(date_first)) %>% 
  ungroup() %>% 
  count(position_category_code,position_category_description)
```

Is this an important variable?

Yes, but we want to use this in the analysis. We will fill the variable the most proximate value and create an indicator that these values were imputed.

```{r}
prep_xx <- 
  prep_building %>% 
  select(date, person_id,all_of(problem_variable %>% pull(variable))) %>% 
  distinct() %>% 
  mutate(
    position_category_imputed = case_when(is.na(position_category_code) ~ 1,
                                          is.na(position_category_description) ~ 1,
                                          TRUE ~ 0) 
  ) %>% 
  arrange(person_id, date) %>% 
  group_by(person_id) %>% 
  fill(position_category_code, .direction = "downup") %>% 
  fill(position_category_description, .direction = "downup") %>% 
  ungroup() %>% 
  mutate(
    across(
      .cols = all_of(problem_variable %>% pull(variable)),
      .fns = ~case_when(is.na(.) ~ "(Missing)",
                        TRUE ~ .)
    )
  )

prep_xx %>%
  skim()
```

Compare against the identified cases above.

```{r}
prep_xx %>% 
  inner_join(prep_01,
            by =  c("person_id", "date"),
            suffix = c("_xx","_01")
  ) %>% 
  datatable_with_sampling(in_max_rows = 5000)
```

Looks good.

Join to building dataframe.

```{r}
prep_problem <- prep_problem %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `supervisor_correction`

Removing problems produced by

```{r}
problem_variable <- 
  problem_variables %>% 
  filter(group == problem_groups[step_n])

problem_variable
```

Reviewing blanks.

```{r}
nulls_01 <- 
  prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(problem_variable %>% pull(variable)),everything())

nulls_01 %>% 
  datatable_with_sampling(in_max_rows = 5000)
```

Let's count the extent of the issues.

```{r}
problem_variable %>% 
  pull(variable) %>% 
  map(
    ~nulls_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Let's count the entries here.

```{r}
prep_building %>% 
  group_by(across(all_of(problem_variable %>% pull(variable)))) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
   flextable() %>% 
  theme_zebra() %>% 
  autofit()
```

Lot's of entries.

Check relationship with time.

```{r}
prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  group_by(across(all_of(problem_variable %>% pull(variable))),date) %>% 
  count() %>% 
  ggplot(aes(x = date, y = n)) +
  geom_jitter() +
  geom_smooth() +
  theme_fivethirtyeight()
```

Some increase in the use of this over time.

Let's count this along side `daily_status`.

```{r}
prep_building %>% 
  group_by(across(all_of(problem_variable %>% pull(variable))),daily_status) %>% 
  count() %>% 
   flextable() %>% 
  theme_zebra() %>% 
  autofit()
```

Is this an important variable?

Yes, but we want to use this in the analysis. We will use with `daily_status` to define status.

```{r}
prep_xx <- 
  prep_building %>% 
  select(date, person_id,all_of(problem_variable %>% pull(variable)),daily_status) %>% 
  distinct() %>% 
  mutate(
    daily_status_imputed = case_when(is.na(supervisor_correction) &
                                       daily_status == "Supervisor Corrected" ~ "(Missing)",
                                     is.na(supervisor_correction) ~ daily_status,
                                     daily_status == "Supervisor Corrected" ~ supervisor_correction
                                     ) 
  )

prep_xx %>%
  skim()
```

Let's count these variables to make sure they are coming out correctly.

```{r}
prep_xx %>% 
  count(supervisor_correction,daily_status,daily_status_imputed) %>% 
  flextable() %>% 
  theme_zebra() %>% 
  autofit()
```

Looks good.

Let's remove `supervisor_correction`.

```{r}
prep_xx <- 
  prep_xx %>% 
  select(-supervisor_correction)
```

Join to building dataframe.

```{r}
prep_problem <- prep_problem %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: `supervisor_e_mail_address`

Removing problems produced by

```{r}
problem_variable <- 
  problem_variables %>% 
  filter(group == problem_groups[step_n])

problem_variable
```

Reviewing blanks.

```{r}
nulls_01 <- 
  prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(problem_variable %>% pull(variable)),everything())

nulls_01 %>% 
  datatable_with_sampling(in_max_rows = 5000)
```

Let's count the extent of the issues.

```{r}
problem_variable %>% 
  pull(variable) %>% 
  map(
    ~nulls_01 %>% 
  count(!!sym(.x)) %>% 
  arrange(desc(n)) %>% 
    filter(is.na(!!sym(.x)))
  )
```

Not too many.

Check relationship with time.

```{r}
prep_building %>% 
  filter(if_any(all_of(problem_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  group_by(across(all_of(problem_variable %>% pull(variable))),date) %>% 
  count() %>% 
  ggplot(aes(x = date, y = n)) +
  geom_jitter() +
  geom_smooth() +
  theme_fivethirtyeight()
```

Let's see if we can find these names in the roster data.

```{r}
nulls_01 %>% 
  select(supervisor_display_name) %>% 
  distinct() %>% 
  separate(supervisor_display_name,c("supervisor_first_name","supervisor_last_name"), sep = " ") %>% 
  mutate(
    full_name = glue::glue("{supervisor_last_name},{supervisor_first_name}")
  ) %>% 
  inner_join(prep_base_2)
```

Let's see if anyone has the same last name.

```{r}
prep_01 <- 
  nulls_01 %>% 
  select(supervisor_display_name) %>% 
  distinct() %>% 
  separate(supervisor_display_name,c("supervisor_first_name","supervisor_last_name"), sep = " ",
           remove = FALSE) %>% 
  mutate(
    full_name = glue::glue("{supervisor_last_name},{supervisor_first_name}")
  ) %>% 
  inner_join(prep_base_2 %>% 
               separate(full_name,c("supervisor_last_name","supervisor_first_name"), sep = ","),
             by = "supervisor_last_name") %>% 
  select(starts_with("supervisor_"), employee_e_mail_address,everything())

prep_01 %>% 
  datatable_with_sampling(in_max_rows = 5000)
```

This looks like the right person. Let's collect her information.

```{r}
prep_02 <- 
  prep_01 %>% 
  select(supervisor_display_name.x,employee_e_mail_address) %>% 
  distinct() %>% 
  filter(!is.na(employee_e_mail_address)) %>% 
  rename(supervisor_display_name = supervisor_display_name.x,
         supervisor_e_mail_address = employee_e_mail_address)

prep_02 %>% 
  datatable_with_sampling(in_max_rows = 5000)
```

```{r}
prep_xx <- 
  prep_building %>% 
  select(date, person_id,all_of(problem_variable %>% pull(variable)),supervisor_display_name) %>% 
  distinct() %>% 
  left_join(prep_02,by = "supervisor_display_name") %>% 
  mutate(
    supervisor_e_mail_address.z = 
      case_when(is.na(supervisor_e_mail_address.x) ~ supervisor_e_mail_address.y,
                TRUE ~supervisor_e_mail_address.x
                                     )
  )

prep_xx %>%
  skim()
```

Let's make sure it only effected this case.

```{r}
prep_xx %>% 
  filter(supervisor_e_mail_address.y == supervisor_e_mail_address.z) %>% 
  count(supervisor_display_name)
```

Looks good. Let's only use `person_id`, `date`, and rename `supervisor_e_mail_address.z`.

```{r}
prep_xx <- 
  prep_xx %>% 
  select(person_id,date,supervisor_e_mail_address.z) %>% 
  rename(supervisor_e_mail_address = supervisor_e_mail_address.z)

prep_xx %>%
  skim()
```

Join to building dataframe.

```{r}
prep_problem <- prep_problem %>% 
  left_join(prep_xx)
```

```{r}
step_n <- step_n + 1
```

#### Step `r step_n`: Done!

Removing problems produced by

```{r}
problem_variable <- 
  problem_variables %>% 
  filter(group == problem_groups[step_n])

problem_variable
```

Done.

### Combining  data

#### Create dataframe off non-null columns

Collect column names of variables not creating duplicates.

```{r}
non_problem_variables <- 
  column_meta_data  %>% 
  filter(!(blanks_by_variable | nulls_by_variable)) %>% 
  pull(variable)

non_problem_variables
```

Create dataframe of only these variables.

```{r}
prep_xx <- 
  prep_building %>% 
  select(any_of(non_problem_variables)) %>% 
  distinct()

prep_xx %>% 
  glimpse()

prep_xx %>% 
  skim()
```

Join to building dataframe.

```{r}
prep_building <- prep_problem %>% 
  left_join(prep_xx)
```

### Checking problems

Created a dataframe of variable metadata.

```{r}
column_meta_data <- 
  prep_building %>% 
  contents()

column_meta_data <- 
  column_meta_data$contents %>% 
  as.data.frame()

column_meta_data <-
  column_meta_data %>% 
  rownames_to_column(var = "variable")

if("NAs" %in% names(column_meta_data))
{
  column_meta_data <- 
  column_meta_data %>% 
  mutate(
    nulls_by_variable = case_when(NAs > 0 ~ TRUE,
                                  NAs == 0 ~ FALSE)
  )
}

if(!("NAs" %in% names(column_meta_data)))
{
  column_meta_data <- 
  column_meta_data %>% 
  mutate(
    nulls_by_variable = FALSE
  )
}

column_meta_data
```

Counting top 10 values for all character or factor columns to identify other common values.

```{r}
values_count <- 
  prep_building %>% 
  select(where(is.character) | where(is.factor)) %>% 
  names()

values_count %>% 
  map(~count(prep_building,!!sym(.)) %>% 
  arrange(desc(n)) %>%
  slice_head(n = 10))

```

No obvious values.

Function that checks whether there are blanks for different variables

```{r}
blank_check <- function(in_data,in_variable_name)
{
  check_01 <- 
  in_data %>% 
    filter(!!sym(in_variable_name) == "") %>% 
  distinct()

out <- nrow(check_01) > 0

return(out)
}
```

Checking blanks for each variable in the dataframe.

```{r}
blanks_by_variable <- 
  column_meta_data %>% 
  pull(variable) %>% 
  map_lgl(~blank_check(in_data = prep_building,.))

blanks_by_variable
```

Add this to the `column_meta_data` dataframe.

```{r}
column_meta_data <- 
  column_meta_data %>% 
  add_column(
    blanks_by_variable = blanks_by_variable
  )

column_meta_data
```

Create dataframe of only variables that have blanks.

```{r}
problem_variables <- 
  column_meta_data %>% 
  filter(blanks_by_variable | nulls_by_variable)

problem_variables
```

## Other data issues

Review the data.

```{r}
prep_building %>% 
  skim()
```

## Creating final dataframe

```{r}
check_in_panel <- prep_building
```

```{r}
rm(prep_building)
```

## Save data {.tabset .tabset-pills}

```{r}
save(check_in_panel, file = file.path(params$google_drive_home,"03. Data Collection","prepared-data",
          glue::glue("check_in_panel_{Sys.Date()}.rdata")))
```

Remove all data elements.

```{r}
rm(list = ls())
gc()
```

## Extraction time {.tabset .tabset-pills}

```{r}
tictoc::toc()
```