### 2023-05-11 04. Analysis/scripts/02_data_preparation_clusters_employee.Rmd
Update 02_data_preparation_clusters_employee.Rmd
Completing rule based categories and saving data.

---
title: "Data preparation for employee category analysis"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    code_folding: hide
params:
  author: "Ben Jaques-Leslie"
  project_number: 2312
  project_name: "Occupancy Patterns"
  google_drive_home: "G:/Shared drives/MSG Projects/1.0 Real Estate Solutions (LQ1)/2312 Work Patterns"
  data_folder_1: "prepared-data"
  data_file_1: "check_in_panel_2023-04-12.rdata"
  unique_identifier_1: !r c("date","person_id")
  date_cohort_month: "2022-03-01"
  observation_window_length_months: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
tictoc::tic()
if(!("arsenal" %in% installed.packages()[,"Package"]))
{
  install.packages("arsenal")
}
if(!("oesrrr" %in% installed.packages()[,"Package"]))
{
  devtools::install_github(repo = "GSA/oesrrr")
}
library(oesrrr)
library(tidyverse)
library(janitor)
library(skimr)
library(DataExplorer)
library(readr)
library(arsenal)
library(Hmisc)
library(readxl)
library(flextable)
library(ggthemes)
```

```{r cohort dates}
date_cohort_month_start <- ymd(params$date_cohort_month)
date_cohort_month_end <- ceiling_date(date_cohort_month_start,unit = "month") %>% rollbackward()
date_observation_window_month_start <- date_cohort_month_start %m+% months(params$observation_window_length_months)
date_observation_window_month_end <- ceiling_date(date_observation_window_month_start,unit = "month") %>% rollbackward()
```

# Data preparation {.tabset .tabset-pills}

**Project number**: `r params$project_number`

**Project name**: `r params$project_name`

**Author**: `r params$author`

**Data folder 1**: `r params$data_folder_1`

**Data file 1**: `r params$data_file_1`

**Unique identifier for data file 1**: `r params$unique_identifier_1`

**Cohort start month**: `r nice_month_year(date_cohort_month_start)`

**Observation window**: `r nice_month_year(date_cohort_month_start)` to `r nice_month_year(date_observation_window_month_end)` (`r params$observation_window_length_months` months)

## Load data {.tabset .tabset-pills}

Load `r params$data_file_1` from `r params$data_folder_1` data folder.

```{r}
load(file = file.path(params$google_drive_home,"03. Data Collection",params$data_folder_1,params$data_file_1))
```

Review.

```{r}
check_in_panel %>% 
  skim()
```

Updated data type for check-in status.

```{r}
check_in_panel <- 
  check_in_panel %>% 
  mutate(
    daily_status_imputed = factor(daily_status_imputed),
    daily_status_imputed =
      fct_recode(daily_status_imputed,
                 permanent_departure = "This employee is no longer with the agency (retired, resigned, or other form of permanent departure)")
  )
```

Create days of the week for dates.

```{r}
check_in_panel <- 
  check_in_panel %>% 
  mutate(date_day_of_week = wday(date, label = TRUE))
```

## Creating cohort {.tabset .tabset-pills}

Collect `person_id` of employees that worked in `r nice_month_year(date_cohort_month_start)`.

```{r}
cohort_ids <- 
  check_in_panel %>% 
  filter(date >= date_cohort_month_start & date <= date_cohort_month_end) %>% 
  select(person_id) %>% 
  distinct()
```

Review.

```{r}
cohort_ids %>% 
  skim()
```

Join with full panel.

```{r}
cohort_panel <- 
  cohort_ids %>% 
  inner_join(check_in_panel) %>% 
  filter(date >= date_cohort_month_start & date <= date_observation_window_month_end)
```

Review.

```{r}
cohort_panel %>% 
  skim()
```

Check: These have the same number of unique `person_id`.

```{r}
cohort_ids %>% 
  summarise(n_distinct(person_id)) %>% 
  pull() ==
  cohort_panel %>% 
  summarise(n_distinct(person_id)) %>% 
  pull()
```

Create dataframe of all initial characteristics.

```{r}
cohort_first <- 
  cohort_panel %>% 
  arrange(person_id,date) %>% 
  group_by(person_id) %>% 
  summarise(
    across(
      .cols = everything(),
      .fns = first
    )
  ) %>% 
  ungroup()
```

Review.

```{r}
cohort_first %>% 
  skim()
```

Check: These have the same number of unique `person_id`.

```{r}
cohort_ids %>% 
  summarise(n_distinct(person_id)) %>% 
  pull() ==
  cohort_first %>% 
  summarise(n_distinct(person_id)) %>% 
  pull()
```

Set up `prep_building`.

```{r}
prep_building <- 
  cohort_ids
```

## Creating aggregated variables {.tabset .tabset-pills}

### Employee, whole period variables

For following variables are aggregated for the whole period for each employee:

-   Proportion of days onsite

-   Proportion of days remote

-   Number of working days

First, calculating the number of days in the observation window.

```{r}
observation_window_days <- 
  cohort_panel %>% 
  summarise(observation_window_days = n_distinct(date)) %>% 
  ungroup() %>% 
  pull()

observation_window_days
```

Now calculating the number of days per employee by `daily_status_imputed`. Proportions are caluculated for onsite and telework only. The calculations are made in two ways in two ways. First as a proportion of employee working days:

$$
\frac{DailyStatusDays}{WorkingDays}
$$

where

$$
WorkingDays = OnsiteDays + TeleworkDays
$$

Second as a proportion of days in the observations window:

$$
\frac{DailyStatusDays}{ObservationWindowDays}
$$

```{r}
prep_xx <- 
  cohort_panel %>%
  group_by(person_id,daily_status_imputed,.drop = FALSE) %>% 
  summarise(working_days = n_distinct(date)) %>% 
  ungroup() %>% 
  pivot_wider(values_from = working_days, names_from = daily_status_imputed) %>% 
  clean_names() %>% 
  rename_with(~glue::glue("daily_status_{.}"), -person_id) %>% 
  mutate(
    working_days = daily_status_reporting_to_job_site + daily_status_teleworking,
    daily_status_reporting_to_job_site_working_prop = daily_status_reporting_to_job_site/working_days,
    daily_status_teleworking_working_prop = daily_status_teleworking/working_days,
    daily_status_reporting_to_job_site_observation_prop = daily_status_reporting_to_job_site/observation_window_days,
    daily_status_teleworking_observation_prop = daily_status_teleworking/observation_window_days
  ) %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  )
```

Review data.

```{r}
prep_xx %>% 
  skim()
```

Check: These have the same number of unique `person_id`.

```{r}
cohort_ids %>% 
  summarise(n_distinct(person_id)) %>% 
  pull() ==
  prep_xx %>% 
  summarise(n_distinct(person_id)) %>% 
  pull()
```

Join to `prep_building`.

```{r}
prep_building <- 
  prep_building %>% 
  left_join(prep_xx)
```

### Employee, day of week variables

For following variables are aggregated for the whole period by day of the week for each employee:

-   Proportion of days remote by the day of the week

-   Correlation of work status between days of the week

#### Proportion of days remote by day of the week

First, calculating the number of days in the observation window.

```{r}
observation_window_days_day_of_week <- 
  cohort_panel %>% 
  group_by(date_day_of_week) %>% 
  summarise(observation_window_days = n_distinct(date)) %>% 
  ungroup()

observation_window_days_day_of_week
```

We make the same calculations as above, but by the day of the week.

```{r}
prep_day_of_week <-
  cohort_panel %>%
  group_by(person_id,daily_status_imputed,date_day_of_week,) %>% 
  summarise(working_days = n_distinct(date)) %>% 
  ungroup() %>% 
  pivot_wider(values_from = working_days, names_from = c(daily_status_imputed)) %>% 
  left_join(observation_window_days_day_of_week) %>% 
  clean_names() %>% 
  rename_with(~glue::glue("daily_status_{.}"), c(-person_id,-date_day_of_week,-observation_window_days))  %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  ) %>% 
  mutate(
    working_days = daily_status_reporting_to_job_site + daily_status_teleworking,
    daily_status_reporting_to_job_site_working_prop = daily_status_reporting_to_job_site/working_days,
    daily_status_teleworking_working_prop = daily_status_teleworking/working_days,
    daily_status_reporting_to_job_site_observation_prop = daily_status_reporting_to_job_site/observation_window_days,
    daily_status_teleworking_observation_prop = daily_status_teleworking/observation_window_days
  )  %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  )
```

Review data.

```{r}
prep_day_of_week %>% 
  skim()
```

Now collect only the results for telework.

```{r}
prep_xx <- 
  prep_day_of_week %>% 
  select(person_id,date_day_of_week,daily_status_teleworking_working_prop,daily_status_teleworking_observation_prop) %>% 
  pivot_longer(cols = starts_with("daily_status")) %>% 
  mutate(date_day_of_week = str_to_lower(date_day_of_week)) %>% 
  unite("name",date_day_of_week,name) %>% 
  pivot_wider(values_from = value, names_from = name) %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  )
```

Review data.

```{r}
prep_xx %>% 
  skim()
```

Check: These have the same number of unique `person_id`.

```{r}
cohort_ids %>% 
  summarise(n_distinct(person_id)) %>% 
  pull() ==
  prep_xx %>% 
  summarise(n_distinct(person_id)) %>% 
  pull()
```

Join to `prep_building`.

```{r}
prep_building <- 
  prep_building %>% 
  left_join(prep_xx)
```

#### Correlation of work status between days of the week

Three strategies to captured relationships between day are below:

1.  Correlation in telework between days of week

2.  Correlation in status (as an ordered factor) between days of week

3.  Proportion of telework on other days of week, if the employ teleworked on a specific day

##### Correlation in telework between days of week

The script below calculates the correlation in telework between days of the week.

```{r correlation remote only}
prep_day_of_week_cor_telework <- 
  cohort_panel %>% 
  select(person_id,date,date_day_of_week,daily_status_imputed) %>% 
  mutate(ind = 1) %>% 
  pivot_wider(names_from = daily_status_imputed, values_from = ind) %>% 
  clean_names() %>% 
  select(person_id,date,date_day_of_week,teleworking) %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  ) %>% 
  mutate(
    date_week_start = floor_date(date,unit = "week")) %>% 
  select(-date) %>% 
  pivot_wider(values_from = teleworking,names_from = date_day_of_week) %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  ) %>% 
  clean_names() %>% 
  group_by(person_id) %>%
  # filter(person_id == "10070")
  summarise(
    mon_tue_cor_telework = cor(mon,tue,method = "kendall", use = "complete.obs"),
    mon_wed_cor_telework = cor(mon,wed,method = "kendall", use = "complete.obs"),
    mon_thu_cor_telework = cor(mon,thu,method = "kendall", use = "complete.obs"),
    mon_fri_cor_telework = cor(mon,fri,method = "kendall", use = "complete.obs"),
    tue_wed_cor_telework = cor(tue,wed,method = "kendall", use = "complete.obs"),
    tue_thu_cor_telework = cor(tue,thu,method = "kendall", use = "complete.obs"),
    tue_fri_cor_telework = cor(tue,fri,method = "kendall", use = "complete.obs"),
    wed_thu_cor_telework = cor(wed,thu,method = "kendall", use = "complete.obs"),
    wed_fri_cor_telework = cor(wed,fri,method = "kendall", use = "complete.obs"),
    thu_fri_cor_telework = cor(thu,fri,method = "kendall", use = "complete.obs")
    ) %>% 
  ungroup()
```

Now rename to attach to `prep_building`.

```{r}
prep_xx <- 
  prep_day_of_week_cor_telework
```

Review data.

```{r}
prep_xx %>% 
  skim()
```

Check: These have the same number of unique `person_id`.

```{r}
cohort_ids %>% 
  summarise(n_distinct(person_id)) %>% 
  pull() ==
  prep_xx %>% 
  summarise(n_distinct(person_id)) %>% 
  pull()
```

Join to `prep_building`.

```{r}
prep_building <- 
  prep_building %>% 
  left_join(prep_xx)
```

##### Correlation in status between days of week

The script below calculates the correlation between work status.

```{r correlation ordered status}
prep_day_of_week_cor_status <- 
  cohort_panel %>% 
  select(person_id,date,date_day_of_week,daily_status_imputed) %>% 
  mutate(
    daily_status_imputed = fct_lump_prop(daily_status_imputed, prop = .1),
    daily_status_imputed = fct_infreq(daily_status_imputed),
    daily_status_imputed = factor(daily_status_imputed, ordered = TRUE)
  )  %>% 
  mutate(
    date_week_start = floor_date(date,unit = "week")) %>% 
  select(-date) %>% 
  pivot_wider(values_from = daily_status_imputed,names_from = date_day_of_week) %>% 
  mutate(
    across(
      .cols = where(is.factor),
      .fns = ~replace_na(.,"Other")
    )
  ) %>% 
  clean_names() %>% 
  group_by(person_id) %>%
  summarise(
    mon_tue_cor_status = cor(as.numeric(mon),as.numeric(tue),method = "spearman", use = "complete.obs"),
    mon_wed_cor_status = cor(as.numeric(mon),as.numeric(wed),method = "spearman", use = "complete.obs"),
    mon_thu_cor_status = cor(as.numeric(mon),as.numeric(thu),method = "spearman", use = "complete.obs"),
    mon_fri_cor_status = cor(as.numeric(mon),as.numeric(fri),method = "spearman", use = "complete.obs"),
    tue_wed_cor_status = cor(as.numeric(tue),as.numeric(wed),method = "spearman", use = "complete.obs"),
    tue_thu_cor_status = cor(as.numeric(tue),as.numeric(thu),method = "spearman", use = "complete.obs"),
    tue_fri_cor_status = cor(as.numeric(tue),as.numeric(fri),method = "spearman", use = "complete.obs"),
    wed_thu_cor_status = cor(as.numeric(wed),as.numeric(thu),method = "spearman", use = "complete.obs"),
    wed_fri_cor_status = cor(as.numeric(wed),as.numeric(fri),method = "spearman", use = "complete.obs"),
    thu_fri_cor_status = cor(as.numeric(thu),as.numeric(fri),method = "spearman", use = "complete.obs")
    ) %>% 
  ungroup()
```

Now rename to attach to `prep_building`.

```{r}
prep_xx <- 
  prep_day_of_week_cor_status
```

Review data.

```{r}
prep_xx %>% 
  skim()
```

Check: These have the same number of unique `person_id`.

```{r}
cohort_ids %>% 
  summarise(n_distinct(person_id)) %>% 
  pull() ==
  prep_xx %>% 
  summarise(n_distinct(person_id)) %>% 
  pull()
```

Join to `prep_building`.

```{r}
prep_building <- 
  prep_building %>% 
  left_join(prep_xx)
```

##### Proportion of telework week day pairs

First, create a dataframe of telework by days of the week.

```{r proportion remote only prep dataframe}
prep_a <- 
  cohort_panel %>% 
  select(person_id,date,date_day_of_week,daily_status_imputed) %>% 
  mutate(ind = 1) %>% 
  pivot_wider(names_from = daily_status_imputed, values_from = ind) %>% 
  clean_names() %>% 
  select(person_id,date,date_day_of_week,teleworking) %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  ) %>% 
  mutate(
    date_week_start = floor_date(date,unit = "week")) %>% 
  select(-date) %>% 
  pivot_wider(values_from = teleworking,names_from = date_day_of_week) %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  ) %>% 
  clean_names()
```

Now, a function to calculate the proportion of telework on a different day, if the employee teleworked on a specific day.

```{r}
day_to_day_proportions <- function(in_data,in_var)
{
  var_string <- rlang::as_name(enquo(in_var))

  # print(person_id)

  out <- 
    in_data %>% 
  filter({{in_var}} == 1) %>% 
  group_by(person_id) %>% 
  summarise(
    across(
      .cols = where(is.numeric),
      .fns = ~sum(.,na.rm = TRUE)
  )) %>% 
  mutate(
    across(
      .cols = c(-person_id,-{{in_var}}),
      .fns = ~./{{in_var}},
      .names = "{var_string}_{.col}"
    )
  ) %>% 
  ungroup() %>% 
    select(person_id,starts_with(var_string),-{{in_var}})

  return(out)
}
```

Iterated over the days of the week and join results.

```{r}
prep_day_of_week_proportion_pairs <- 
  prep_a %>% 
  select(thu:fri) %>%
  names() %>% 
  map(~day_to_day_proportions(prep_a,!!sym(.))) %>% 
  reduce(full_join, by = "person_id")

```

Now rename to attach to `prep_building`.

```{r}
prep_xx <- 
  prep_day_of_week_proportion_pairs %>% 
  rename_with(-person_id,
              .fn = ~glue::glue("{.}_prop_pairs")
  )
```

Review data.

```{r}
prep_xx %>% 
  skim()
```

Check: These have the same number of unique `person_id`.

```{r}
cohort_ids %>% 
  summarise(n_distinct(person_id)) %>% 
  pull() ==
  prep_xx %>% 
  summarise(n_distinct(person_id)) %>% 
  pull()
```

Join to `prep_building`.

```{r}
prep_building <- 
  prep_building %>% 
  left_join(prep_xx)
```

#### Modal size of team over observation window

Calculate number of team members per day by supervisor email.

```{r}
prep_a <- 
  cohort_panel %>% 
  group_by(date,supervisor_e_mail_address) %>% 
  summarise(team_size = n_distinct(person_id)) %>% 
  ungroup()
```

Review data.

```{r}
prep_a %>% 
  skim()
```

Join to `cohort_panel` only selecting, `person_id`, `date`, and `supervisor_e_mail_address`.

```{r}
prep_b <- 
  cohort_panel %>% 
  select(date,person_id,supervisor_e_mail_address) %>% 
  left_join(prep_a)
```

Review data.

```{r}
prep_b %>% 
  skim()
```

Calculate the mode team size for each employee.

```{r}
prep_xx <- 
  prep_b %>% 
  group_by(person_id) %>% 
  summarise(team_size_mode = getmode(team_size)) %>% 
  ungroup()
```

Review data.

```{r}
prep_xx %>% 
  skim()
```

Check: These have the same number of unique `person_id`.

```{r}
cohort_ids %>% 
  summarise(n_distinct(person_id)) %>% 
  pull() ==
  prep_xx %>% 
  summarise(n_distinct(person_id)) %>% 
  pull()
```

Join to `prep_building`.

```{r}
prep_building <- 
  prep_building %>% 
  left_join(prep_xx)
```

#### Proportion of working days co-located with at least one other team member

Identify employees who report to the site and collect `person_id`, `dates`, `location`, and `supervisor_e_mail_address`.

```{r}
prep_a <-
  cohort_panel %>%
  filter(daily_status_imputed == "Reporting to Job Site") %>%
  select(person_id, date, location, supervisor_e_mail_address) %>%
  distinct()
```

Review data.

```{r}
prep_a %>% 
  skim()
```

Self left join on `dates`, `location`, and `supervisor_e_mail_address` and filter to rows where the `person_id`s do not match.

```{r}
prep_b <- 
  prep_a %>% 
  left_join(prep_a,
            by = c("date", "location", "supervisor_e_mail_address"),
            suffix = c("","_joined")) %>%
  filter(person_id != person_id_joined)
```

Review data.

```{r}
prep_b %>% 
  skim()
```

Count dates where co-location occurs by `person_id`.

```{r}
prep_c <- 
  prep_b %>% 
  group_by(person_id) %>% 
  summarise(days_colocated_1 = n_distinct(date)) %>% 
  ungroup()
```

Review data.

```{r}
prep_c %>% 
  skim()
```

Full join `working_days` from `prep_building`, calculate proportion of co-located days with at least one team member by `working_days` and `observation_window_days`, replace nulls with 0, and remove `working_days`.

```{r}
prep_xx <- 
  prep_c %>% 
  full_join(prep_building %>% 
              select(person_id,working_days)) %>% 
  mutate(days_colocated_1_working_prop = days_colocated_1/working_days,
         days_colocated_1_observation_prop = days_colocated_1/observation_window_days) %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  ) %>% 
  select(-working_days)
```

Review data.

```{r}
prep_xx %>% 
  skim()
```

Check: These have the same number of unique `person_id`.

```{r}
cohort_ids %>% 
  summarise(n_distinct(person_id)) %>% 
  pull() ==
  prep_xx %>% 
  summarise(n_distinct(person_id)) %>% 
  pull()
```

Join to `prep_building`.

```{r}
prep_building <- 
  prep_building %>% 
  left_join(prep_xx)
```

#### Proportion of working days co-located with supervisor

Identify employees who report to the site and collect `person_id`, `dates`, `location`, `employee_e_mail_address`, and `supervisor_e_mail_address`.

```{r}
prep_a <-
  cohort_panel %>%
  filter(daily_status_imputed == "Reporting to Job Site") %>%
  select(person_id, date, location, employee_e_mail_address, supervisor_e_mail_address) %>%
  distinct()
```

Review data.

```{r}
prep_a %>% 
  skim()
```

Self inner join on `dates`, `location`, and `supervisor_e_mail_address` equal to `employee_e_mail_address` in the right table, and filter to rows where the `person_id`s do not match.

```{r}
prep_b <- 
  prep_a %>% 
  inner_join(prep_a,
            by = c("date", "location", "supervisor_e_mail_address" = "employee_e_mail_address"),
            suffix = c("","_joined")) 
```

Review data.

```{r}
prep_b %>% 
  skim()
```

Count dates where co-location with supervisor occurs by `person_id`.

```{r}
prep_c <- 
  prep_b %>% 
  group_by(person_id) %>% 
  summarise(days_colocated_supervisor = n_distinct(date)) %>% 
  ungroup()
```

Review data.

```{r}
prep_c %>% 
  skim()
```

Full join `working_days` from `prep_building`, calculate proportion of co-located days with at least one team member by `working_days` and `observation_window_days`, replace nulls with 0, and remove `working_days`.

```{r}
prep_xx <- 
  prep_c %>% 
  full_join(prep_building %>% 
              select(person_id,working_days)) %>% 
  mutate(days_colocated_supervisor_working_prop = days_colocated_supervisor/working_days,
         days_colocated_supervisor_observation_prop = days_colocated_supervisor/observation_window_days) %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  ) %>% 
  select(-working_days)
```

Review data.

```{r}
prep_xx %>% 
  skim()
```

Check: These have the same number of unique `person_id`.

```{r}
cohort_ids %>% 
  summarise(n_distinct(person_id)) %>% 
  pull() ==
  prep_xx %>% 
  summarise(n_distinct(person_id)) %>% 
  pull()
```

Join to `prep_building`.

```{r}
prep_building <- 
  prep_building %>% 
  left_join(prep_xx)
```

#### Proportion of working days co-located with entire team

Identify employees who report to the site and collect `person_id`, `dates`, `location`, and `supervisor_e_mail_address`.

```{r}
prep_a <-
  cohort_panel %>%
  filter(daily_status_imputed == "Reporting to Job Site") %>%
  select(person_id, date, location, supervisor_e_mail_address) %>%
  distinct()
```

Review data.

```{r}
prep_a %>% 
  skim()
```

Self left join on `dates`, `location`, and `supervisor_e_mail_address`.

```{r}
prep_b <- 
  prep_a %>% 
  left_join(prep_a,
            by = c("date", "location", "supervisor_e_mail_address"),
            suffix = c("","_joined"))
```

Review data.

```{r}
prep_b %>% 
  skim()
```

Count number of team members co-located by `person_id` and `date` by counting unique `person_id_joined`.

```{r}
prep_c <- 
  prep_b %>% 
  group_by(person_id,date,supervisor_e_mail_address) %>% 
  summarise(colocated_team_members = n_distinct(person_id_joined)) %>% 
  ungroup()
```

Review data.

```{r}
prep_c %>% 
  skim()
```

Calculate number of team members per day by supervisor email.

```{r}
prep_d <- 
  cohort_panel %>% 
  group_by(date,supervisor_e_mail_address) %>% 
  summarise(team_size = n_distinct(person_id)) %>% 
  ungroup()
```

Review data.

```{r}
prep_d %>% 
  skim()
```

Left join co-located team members and teams size, but `date` and `supervisor_e_mail_address`, then filter to when co-located team members is the same size as the team

```{r}
prep_e <- 
  prep_c %>% 
  left_join(prep_d) %>% 
  filter(colocated_team_members == team_size)
```

Review data.

```{r}
prep_e %>% 
  skim()
```

Count dates where co-location with whole team occurs by `person_id`.

```{r}
prep_f <- 
  prep_e %>% 
  group_by(person_id) %>% 
  summarise(days_colocated_team = n_distinct(date)) %>% 
  ungroup()
```

Review data.

```{r}
prep_f %>% 
  skim()
```

Full join `working_days` from `prep_building`, calculate proportion of co-located days with team member by `working_days` and `observation_window_days`, replace nulls with 0, and remove `working_days`.

```{r}
prep_xx <- 
  prep_f %>% 
  full_join(prep_building %>% 
              select(person_id,working_days)) %>% 
  mutate(days_colocated_team_working_prop = days_colocated_team/working_days,
         days_colocated_team_observation_prop = days_colocated_team/observation_window_days) %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  ) %>% 
  select(-working_days)
```

Review data.

```{r}
prep_xx %>% 
  skim()
```

Check: These have the same number of unique `person_id`.

```{r}
cohort_ids %>% 
  summarise(n_distinct(person_id)) %>% 
  pull() ==
  prep_xx %>% 
  summarise(n_distinct(person_id)) %>% 
  pull()
```

Join to `prep_building`.

```{r}
prep_building <- 
  prep_building %>% 
  left_join(prep_xx)
```

#### Proportion of working days entire team is remote

Identify employees who telework and collect `person_id`, `dates`, `location`, and `supervisor_e_mail_address`.

```{r}
prep_a <-
  cohort_panel %>% 
  filter(daily_status_imputed == "Teleworking") %>%
  select(person_id, date, location, supervisor_e_mail_address) %>%
  distinct()
```

Review data.

```{r}
prep_a %>% 
  skim()
```

Self left join on `dates`, `location`, and `supervisor_e_mail_address`.

```{r}
prep_b <- 
  prep_a %>% 
  left_join(prep_a,
            by = c("date", "location", "supervisor_e_mail_address"),
            suffix = c("","_joined"))
```

Review data.

```{r}
prep_b %>% 
  skim()
```

Count number of team members co-located by `person_id` and `date` by counting unique `person_id_joined`.

```{r}
prep_c <- 
  prep_b %>% 
  group_by(person_id,date,supervisor_e_mail_address) %>% 
  summarise(telework_team_members = n_distinct(person_id_joined)) %>% 
  ungroup()
```

Review data.

```{r}
prep_c %>% 
  skim()
```

Calculate number of team members per day by supervisor email.

```{r}
prep_d <- 
  cohort_panel %>% 
  group_by(date,supervisor_e_mail_address) %>% 
  summarise(team_size = n_distinct(person_id)) %>% 
  ungroup()
```

Review data.

```{r}
prep_d %>% 
  skim()
```

Left join telework team members and teams size, but `date` and `supervisor_e_mail_address`, then filter to when telework team members is the same size as the team

```{r}
prep_e <- 
  prep_c %>% 
  left_join(prep_d) %>% 
  filter(telework_team_members == team_size)
```

Review data.

```{r}
prep_e %>% 
  skim()
```

Count dates where telework with whole team occurs by `person_id`.

```{r}
prep_f <- 
  prep_e %>% 
  group_by(person_id) %>% 
  summarise(days_telework_team = n_distinct(date)) %>% 
  ungroup()
```

Review data.

```{r}
prep_f %>% 
  skim()
```

Full join `working_days` from `prep_building`, calculate proportion of telework days with team member by `working_days` and `observation_window_days`, replace nulls with 0, and remove `working_days`.

```{r}
prep_xx <- 
  prep_f %>% 
  full_join(prep_building %>% 
              select(person_id,working_days)) %>% 
  mutate(days_telework_team_working_prop = days_telework_team/working_days,
         days_telework_team_observation_prop = days_telework_team/observation_window_days) %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  ) %>% 
  select(-working_days)
```

Review data.

```{r}
prep_xx %>% 
  skim()
```

Check: These have the same number of unique `person_id`.

```{r}
cohort_ids %>% 
  summarise(n_distinct(person_id)) %>% 
  pull() ==
  prep_xx %>% 
  summarise(n_distinct(person_id)) %>% 
  pull()
```

Join to `prep_building`.

```{r}
prep_building <- 
  prep_building %>% 
  left_join(prep_xx)
```

## Define rule-based groups {.tabset .tabset-pills}

The analysis plan defines nine rule-based groups for analysis:

1.  **All remote** - A policy definition for staff that are 100% remote
2.  **Routine teleworker** - Employees with position_category_description's *Onsite Flexible* or *Offsite* that work at least two two days per pay period onsite.
    per pay period onsite.
3.  **Onsite Flexible** - Defined by `position_category_description` in
    the roster data
4.  **Offsite** - Defined by `position_category_description` in the
    roster data
5.  **Onsite Required** - Defined by `position_category_description` in
    the roster data
6.  **Primarily remote** - Employees who work remotely above the 95th
    percentile <!-- could be 75th -->
7.  **Mostly remote** - Employees who work remotely between the median
    and the 95th percentile <!-- could be 75th -->
8.  **Mostly onsite** - Employees who work remotely between 5th
    percentile and the median <!-- could be 25th -->
9.  **Primarily onsite** - Employees who work remotely below the 5th
    percentile <!-- could be 75th -->

Below the variables for the groups are created.

### All remote

Define workers with 100% of their time as **all remote**.

```{r}
prep_xx <- 
  prep_building %>% 
  mutate(
    rule_based_all_remote = case_when(
      daily_status_teleworking_working_prop == 1 ~ 1,
                                      TRUE ~ 0)
  ) %>% 
  select(person_id,starts_with("rule_based_"))
```

Review data.

```{r}
prep_xx %>% 
  skim()
```

Check: These have the same number of unique `person_id`.

```{r}
cohort_ids %>% 
  summarise(n_distinct(person_id)) %>% 
  pull() ==
  prep_xx %>% 
  summarise(n_distinct(person_id)) %>% 
  pull()
```

Join to `prep_building`.

```{r}
prep_building <- 
  prep_building %>% 
  left_join(prep_xx)
```

### Primarily, mostly onsite and remote

First, calculate the quantiles for probabilities 0.50 and 0.95.

```{r}
prep_a <- 
  prep_building %>% 
  summarise(
    across(
      .cols = c(daily_status_reporting_to_job_site_working_prop,daily_status_teleworking_working_prop,
                daily_status_reporting_to_job_site_observation_prop,daily_status_teleworking_observation_prop),
      .fns = list(q_95 = ~quantile(.,probs = .95),
                  q_50 = ~quantile(.,probs = .50))
    )
  )
```

Review data.

```{r}
prep_a %>% 
  skim()
```

Use the quantiles above to create new variables for `primarily_` and `mostly_` working remote or onsite.

```{r}
prep_xx <-
  prep_building %>%
  select(
    person_id,
    daily_status_reporting_to_job_site_working_prop,
    daily_status_teleworking_working_prop,
    daily_status_reporting_to_job_site_observation_prop,
    daily_status_teleworking_observation_prop
  ) %>%
  mutate(
    rule_based_primarily_reporting_to_job_site_working =
      case_when(
        daily_status_reporting_to_job_site_working_prop >= prep_a$daily_status_reporting_to_job_site_working_prop_q_95 ~ 1,
        TRUE ~ 0
      ),
    rule_based_primarily_teleworking_working =
      case_when(
        daily_status_teleworking_working_prop >= prep_a$daily_status_teleworking_working_prop_q_95 ~ 1,
        TRUE ~ 0
      ),
    rule_based_primarily_reporting_to_job_site_observation =
      case_when(
        daily_status_reporting_to_job_site_observation_prop >= prep_a$daily_status_reporting_to_job_site_observation_prop_q_95 ~ 1,
        TRUE ~ 0
      ),
    rule_based_primarily_teleworking_observation =
      case_when(
        daily_status_teleworking_observation_prop >= prep_a$daily_status_teleworking_observation_prop_q_95 ~ 1,
        TRUE ~ 0
      ),
    rule_based_mostly_reporting_to_job_site_working =
      case_when(
        daily_status_reporting_to_job_site_working_prop >= prep_a$daily_status_reporting_to_job_site_working_prop_q_50 ~ 1,
        TRUE ~ 0
      ),
    rule_based_mostly_teleworking_working =
      case_when(
        daily_status_teleworking_working_prop >= prep_a$daily_status_teleworking_working_prop_q_50 ~ 1,
        TRUE ~ 0
      ),
    rule_based_mostly_reporting_to_job_site_observation =
      case_when(
        daily_status_reporting_to_job_site_observation_prop >= prep_a$daily_status_reporting_to_job_site_observation_prop_q_50 ~ 1,
        TRUE ~ 0
      ),
    rule_based_mostly_teleworking_observation =
      case_when(
        daily_status_teleworking_observation_prop >= prep_a$daily_status_teleworking_observation_prop_q_50 ~ 1,
        TRUE ~ 0
      ),
  ) %>%
  select(person_id, starts_with("rule_based_"))
```

Review data.

```{r}
prep_xx %>% 
  skim()
```

Check: These have the same number of unique `person_id`.

```{r}
cohort_ids %>% 
  summarise(n_distinct(person_id)) %>% 
  pull() ==
  prep_xx %>% 
  summarise(n_distinct(person_id)) %>% 
  pull()
```

Join to `prep_building`.

```{r}
prep_building <- 
  prep_building %>% 
  left_join(prep_xx)
```

### Position description categories

```{r, eval=FALSE}
prep_b <- 
  cohort_panel %>% 
  group_by(person_id) %>% 
  summarise(
    across(
      .cols = position_category_description,
      .fns = getmode
    )
  )

prep_c <- 
  cohort_first %>% 
  select(person_id,position_category_description)

prep_c %>% 
  inner_join(prep_a, by = "person_id") %>% 
  filter(position_category_description.x != position_category_description.y)
```

Identify the `position_category_description` at the beginning of the observation window and convert the categories into indicators.

```{r}
prep_xx <- 
  cohort_first %>% 
  select(person_id,position_category_description) %>% 
  character_to_indicator(in_unique_identifier = "person_id",
                         position_category_description,
                         in_prefix = "rule_based_position_category_")
```

Review data.

```{r}
prep_xx %>% 
  skim()
```

Check: These have the same number of unique `person_id`.

```{r}
cohort_ids %>% 
  summarise(n_distinct(person_id)) %>% 
  pull() ==
  prep_xx %>% 
  summarise(n_distinct(person_id)) %>% 
  pull()
```

Join to `prep_building`.

```{r}
prep_building <- 
  prep_building %>% 
  left_join(prep_xx)
```

### Teleworker

First define pay periods based on this: https://www.gsa.gov/cdnstatic/GSA_Payroll_Calendar_2022_Revised.pdf. File is also saved in `r file.path(params$google_drive_home,"04. Analysis","documents")`.

```{r}
prep_a <- 
  tibble(
  date = seq.Date(from = ymd("2022-01-01"),
         to = ymd("2022-12-31"),
         by = "2 weeks")
) %>% 
  rownames_to_column(var = "period_number") 
```

Review data.

```{r}
prep_a %>% 
  skim()
```

Second, create a dataframe of all the days of the year.

```{r}
prep_b <- 
  tibble(
  date = seq.Date(from = ymd("2022-01-01"),
         to = ymd("2022-12-31"),
         by = "days")
)
```

Review data.

```{r}
prep_b %>% 
  skim()
```

Third, left join the pay periods dataframe (`prep_a`) to the days of the year (`prep_b`), then fill the null `period_number` values with the entry above.

```{r}
prep_c <- 
  prep_b %>% 
  left_join(prep_a) %>% 
  fill(period_number, .direction = "down")
```
Review data.

```{r}
prep_c %>% 
  skim()
```

Fourth, take the cohort panel (`cohort_panel`) select `person_id`, `date`, and `daily_status_imputed` and left join to calendar dataframe with filled `period_number` (`prep_c`) on `date`.

```{r}
prep_d <- 
  cohort_panel %>% 
  select(person_id,date, daily_status_imputed) %>% 
  left_join(prep_c)
```

Review data.

```{r}
prep_d %>% 
  skim()
```

Fifth, count the number of unique days by `person_id`, `daily_status_imputed`, and `period_number`.

```{r}
prep_e <- 
      prep_d %>% 
  group_by(person_id,daily_status_imputed,period_number) %>% 
  summarise(days = n_distinct(date)) %>% 
  ungroup()
```

Review data.

```{r}
prep_e %>% 
  skim()
```

Sixth, filter to only see observations where `daily_status_imputed` is "Reporting to Job Site" and create `telework_onsite_minimum` to be 1 when `days` is greater than or equal to 2, that is employees have at least two days in the office in that pay period. Then, calculate the mean `telework_onsite_minimum` (`telework_onsite_minimum_mean`) per employee. Those where this is 1 are in the office at least two days per pay period for the whole observation window. Lastly, create `telework_onsite_minimum` to be 1 when `telework_onsite_minimum_mean` is 1.

```{r}
prep_f <- 
  prep_e %>% 
  filter(daily_status_imputed == "Reporting to Job Site") %>% 
  mutate(
    telework_onsite_minimum = case_when(days >= 2 ~ 1,
                                        TRUE ~ 0)
  ) %>% 
  group_by(person_id) %>% 
  summarise(
    telework_onsite_minimum_mean = mean(telework_onsite_minimum)
  ) %>% 
  ungroup() %>% 
  mutate(
    telework_onsite_minimum = case_when(telework_onsite_minimum_mean == 1 ~ 1,
                                        TRUE ~ 0)
  )
```

Review data.

```{r}
prep_f %>% 
  skim()
```

Seventh, select `person_id`, and `position_category_description` from the beginning of the observation window (`cohort_first`) and left join the above (`prep_f`). Then, create `rule_based_telework_onsite_minimum` when `telework_onsite_minimum` is 1 and `position_category_description` is *Onsite Flexible* or *Offsite*.

```{r}
prep_xx <- 
  cohort_first %>% 
  select(person_id,position_category_description) %>% 
  left_join(prep_f) %>% 
  mutate(
    rule_based_telework_onsite_minimum = 
      case_when(telework_onsite_minimum == 1 & position_category_description == "Onsite Flexible" ~ 1,
                telework_onsite_minimum == 1 & position_category_description == "Offsite" ~ 1,
                TRUE ~ 0)
  ) %>%
  select(person_id, starts_with("rule_based_"))
```

Review data.

```{r}
prep_xx %>% 
  skim()
```

Check: These have the same number of unique `person_id`.

```{r}
cohort_ids %>% 
  summarise(n_distinct(person_id)) %>% 
  pull() ==
  prep_xx %>% 
  summarise(n_distinct(person_id)) %>% 
  pull()
```

Join to `prep_building`.

```{r}
prep_building <- 
  prep_building %>% 
  left_join(prep_xx)
```

## Attach characteristics

Define `cohort_first` as `prep_xx`.

```{r}
prep_xx <- cohort_first
```

Review data.

```{r}
prep_xx %>% 
  skim()
```

Check: These have the same number of unique `person_id`.

```{r}
cohort_ids %>% 
  summarise(n_distinct(person_id)) %>% 
  pull() ==
  prep_xx %>% 
  summarise(n_distinct(person_id)) %>% 
  pull()
```

Join to `prep_building`.

```{r}
prep_building <- 
  prep_building %>% 
  left_join(prep_xx)
```

## Addressing blanks, nulls, and other values

Created a dataframe of variable metadata.

```{r}
column_meta_data <- 
  prep_building %>% 
  contents()

column_meta_data <- 
  column_meta_data$contents %>% 
  as.data.frame()

column_meta_data <-
  column_meta_data %>% 
  rownames_to_column(var = "variable")

column_meta_data <- 
  column_meta_data %>% 
  mutate(
    nulls_by_variable = case_when(NAs > 0 ~ TRUE,
                                  NAs == 0 ~ FALSE)
  )

column_meta_data
```

Counting top 10 values for all character or factor columns to identify other common values.

```{r}
values_count <- 
  prep_building %>% 
  select(where(is.character) | where(is.factor)) %>% 
  names()

values_count %>% 
  map(~count(prep_building,!!sym(.)) %>% 
  arrange(desc(n)) %>%
  slice_head(n = 10))

```

No obvious values.

Function that checks whether there are blanks for different variables

```{r}
blank_check <- function(in_data,in_variable_name)
{
  check_01 <- 
  in_data %>% 
    filter(!!sym(in_variable_name) == "") %>% 
  distinct()

out <- nrow(check_01) > 0

return(out)
}
```

Checking blanks for each variable in the dataframe.

```{r}
blanks_by_variable <- 
  column_meta_data %>% 
  pull(variable) %>% 
  map_lgl(~blank_check(in_data = prep_building,.))

blanks_by_variable
```

Add this to the `column_meta_data` dataframe.

```{r}
column_meta_data <- 
  column_meta_data %>% 
  add_column(
    blanks_by_variable = blanks_by_variable
  )

column_meta_data
```

Create dataframe of only variables that have blanks.

```{r}
problem_variables <- 
  column_meta_data %>% 
  filter(blanks_by_variable | nulls_by_variable)

problem_variables
```

All of these are null correlations, which will address in analysis.

## Creating final dataframe

```{r}
prep_employees <- prep_building
```

```{r}
rm(prep_building)
```

## Save data {.tabset .tabset-pills}

```{r}
save(prep_employees, file = file.path(params$google_drive_home,"03. Data Collection","prepared-data",
          glue::glue("prep_employees_{Sys.Date()}.rdata")))
```

Remove all data elements.

```{r}
rm(list = ls())
gc()
```

## Extraction time {.tabset .tabset-pills}

```{r}
tictoc::toc()
```