### 2023-05-31 04. Analysis/scripts/03-01-data-analysis-employee-categories.Rmd
Update 03-01-data-analysis-employee-categories.Rmd
Scaled variables for clusters and began testing cluster analysis

---
title: "Data analysis for employee categories"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    code_folding: hide
params:
  author: "Ben Jaques-Leslie"
  project_number: 2312
  project_name: "Occupancy Patterns"
  google_drive_home: "G:/Shared drives/MSG Projects/1.0 Real Estate Solutions (LQ1)/2312 Work Patterns"
  data_folder_1: "prepared-data"
  data_file_1: "prep_employees_2023-05-24.rdata"
  unique_identifier_1: !r c("date","person_id")
  date_cohort_month: "2022-03-01"
  observation_window_length_months: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
tictoc::tic()
if(!("arsenal" %in% installed.packages()[,"Package"]))
{
  install.packages("arsenal")
}
if(!("oesrrr" %in% installed.packages()[,"Package"]))
{
  devtools::install_github(repo = "GSA/oesrrr")
}
library(oesrrr)
library(tidyverse)
library(janitor)
library(skimr)
library(DataExplorer)
library(readr)
library(arsenal)
library(Hmisc)
library(readxl)
library(flextable)
library(ggthemes)
library(tidymodels)
```

```{r cohort dates}
date_cohort_month_start <- ymd(params$date_cohort_month)
date_cohort_month_end <- ceiling_date(date_cohort_month_start,unit = "month") %>% rollbackward()
date_observation_window_month_start <- date_cohort_month_start %m+% months(params$observation_window_length_months)
date_observation_window_month_end <- ceiling_date(date_observation_window_month_start,unit = "month") %>% rollbackward()
```

```{r}
set.seed(27)
```

# Data preparation {.tabset .tabset-pills}

**Project number**: `r params$project_number`

**Project name**: `r params$project_name`

**Author**: `r params$author`

**Data folder 1**: `r params$data_folder_1`

**Data file 1**: `r params$data_file_1`

**Unique identifier for data file 1**: `r params$unique_identifier_1`

**Cohort start month**: `r nice_month_year(date_cohort_month_start)`

**Observation window**: `r nice_month_year(date_cohort_month_start)` to `r nice_month_year(date_observation_window_month_end)` (`r params$observation_window_length_months` months)

## Load data {.tabset .tabset-pills}

Load `r params$data_file_1` from `r params$data_folder_1` data folder.

```{r}
load(file = file.path(params$google_drive_home,"03. Data Collection",params$data_folder_1,params$data_file_1))
```

Review.

```{r}
prep_employees %>% 
  skim()
```

## Scale data {.tabset .tabset-pills}

```{r}
d <- 
  prep_employees %>% 
  mutate(
    across(
      .cols = c(daily_status_reporting_to_job_site_working_prop,
                daily_status_teleworking_working_prop,
                daily_status_reporting_to_job_site_observation_prop,
                daily_status_teleworking_observation_prop,
                starts_with("mon_"),
                starts_with("tue_"),
                starts_with("wed_"),
                starts_with("thu_"),
                starts_with("fri_"),
                working_days,
                team_size_mode,
                starts_with("days_colocated_"),
                starts_with("days_telework_"),
                -days_colocated_1,
                -days_colocated_supervisor,
                -days_colocated_team,
                -days_telework_team
                ),
      .fns = list(scale = scale,rank = rank),
      .names = "{.col}_{.fn}"
    )
  )
```

Review data.

```{r}
d %>% 
  skim()
```

## Create clusters {.tabset .tabset-pills}

### Scale

#### Observation

##### Correlation telework

```{r}
prep_a <-
  d %>%
  select(
    person_id,
    ends_with("_scale"),
    -contains("_working_prop_"),
    -contains("_cor_status_"),
    -contains("_prop_pairs_"),
    ) %>% 
  drop_na()
```

```{r}
prep_a %>% 
  skim()
```

```{r}
prep_b <- 
  prep_a %>% 
  select(-person_id)
```

Review data.

```{r}
prep_b %>% 
  skim()
```

```{r}
kclusts <- 
  tibble(k = 1:9) %>%
  mutate(
    kclust = map(k, ~kmeans(prep_b, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, prep_b)
  )

```

```{r}
clusters <- 
  kclusts %>%
  unnest(cols = c(tidied))

assignments <- 
  kclusts %>% 
  unnest(cols = c(augmented))

clusterings <- 
  kclusts %>%
  unnest(cols = c(glanced))

```

```{r}
p1 <- 
  ggplot(assignments, aes(x = days_colocated_1_observation_prop_scale, y = days_colocated_supervisor_observation_prop_scale)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p1

```

```{r}
ggplot(clusterings, aes(k, tot.withinss)) +
  geom_line() +
  geom_point()

```

```{r}
clusterings %>% 
  select(k,tot.withinss) %>% 
  mutate(
    tot.withinss_lag = lag(tot.withinss),
    diff = tot.withinss_lag - tot.withinss,
    change = diff/tot.withinss_lag,
    change_lag = lag(change),
    diff_2 = change_lag - change
  )
```

```{r}
kclust <- kmeans(points, centers = 3)
kclust
#> K-means clustering with 3 clusters of sizes 148, 51, 101
#> 
#> Cluster means:
#>        x1    x2
#> 1  0.0885  1.05
#> 2 -3.1429 -2.00
#> 3  5.0040 -1.05
#> 
#> Clustering vector:
#>   [1] 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3
#>  [38] 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3
#>  [75] 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 3 1 1 1 1 1 1 1 1 1 1 1
#> [112] 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1
#> [149] 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1
#> [186] 1 1 1 1 1 1 1 1 1 1 1 1 1 3 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1
#> [223] 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 1 2 1 1 1 1 1 1 2 2 2 2 2 2 2 2 2
#> [260] 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2
#> [297] 2 2 2 2
#> 
#> Within cluster sum of squares by cluster:
#> [1] 299 109 243
#>  (between_SS / total_SS =  82.5 %)
#> 
#> Available components:
#> 
#> [1] "cluster"      "centers"      "totss"        "withinss"     "tot.withinss"
#> [6] "betweenss"    "size"         "iter"         "ifault"
summary(kclust)
#>              Length Class  Mode   
#> cluster      300    -none- numeric
#> centers        6    -none- numeric
#> totss          1    -none- numeric
#> withinss       3    -none- numeric
#> tot.withinss   1    -none- numeric
#> betweenss      1    -none- numeric
#> size           3    -none- numeric
#> iter           1    -none- numeric
#> ifault         1    -none- numeric

```

##### Correlation status

##### Proportion pairs

#### Working

##### Correlation telework

##### Correlation status

##### Proportion pairs

### Rank

#### Observation

##### Correlation telework

##### Correlation status

##### Proportion pairs

#### Working

##### Correlation telework

##### Correlation status

##### Proportion pairs