### 2023-08-22 04. Analysis/scripts/03-03-data-analysis-employee-cluster-descriptive-statistics.Rmd
Update 03-03-data-analysis-employee-cluster-descriptive-statistics.Rmd
Adding summaries for GSA regions, position categories and grades.

---
title: "Data analysis for employee cluster descriptive statistics"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    code_folding: hide
params:
  author: "Ben Jaques-Leslie"
  project_number: 2312
  project_name: "Occupancy Patterns"
  google_drive_home: "G:/Shared drives/GSA Evaluation Division/Projects/1.0 Real Estate Solutions (LQ1)/2312 Work Patterns"
  data_folder_1: "prepared-data"
  data_file_1: "clusters-employee-2023-08-09.rdata"
  data_file_2: "prep_employees_2023-08-09.rdata"
  in_number_of_clusters_1: 2
  in_number_of_clusters_2: 5
  in_specification: "scale_observation_none_1_20_TRUE"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
tictoc::tic()
if(!("arsenal" %in% installed.packages()[,"Package"]))
{
  install.packages("arsenal")
}
if(!("beepr" %in% installed.packages()[,"Package"]))
{
  install.packages("beepr")
}
if(!("geomtextpath" %in% installed.packages()[,"Package"]))
{
  install.packages("geomtextpath")
}
if(!("oesrrr" %in% installed.packages()[,"Package"]))
{
  devtools::install_github(repo = "GSA/oesrrr")
}
if(!("oesrrr" %in% installed.packages()[,"Package"]))
{
  devtools::install_github(repo = "GSA/oesrrr")
}
library(oesrrr)
library(tidyverse)
library(janitor)
library(skimr)
library(DataExplorer)
library(readr)
library(arsenal)
library(Hmisc)
library(readxl)
library(flextable)
library(ggthemes)
library(tidymodels)
```

```{r cohort dates}
# date_cohort_month_start <- ymd(params$date_cohort_month)
# date_cohort_month_end <- ceiling_date(date_cohort_month_start,unit = "month") %>% rollbackward()
# date_observation_window_month_start <- date_cohort_month_start %m+% months(params$observation_window_length_months)
# date_observation_window_month_end <- ceiling_date(date_observation_window_month_start,unit = "month") %>% rollbackward()
```

# Data preparation {.tabset .tabset-pills}

**Project number**: `r params$project_number`

**Project name**: `r params$project_name`

**Author**: `r params$author`

**Data folder 1**: `r params$data_folder_1`

**Data file 1**: `r params$data_file_1`

**Data file 2**: `r params$data_file_2`

**Number of clusters**: `r params$in_number_of_clusters`

**Specification** `r params$in_specification`

## Load data {.tabset .tabset-pills}

Load `r params$data_file_1` from `r params$data_folder_1` data folder.

```{r}
load(file = file.path(params$google_drive_home,"03. Data Collection",params$data_folder_1,params$data_file_1))
```

Review.

```{r}
kclusts %>% 
  skim()
```

Load `r params$data_file_2` from `r params$data_folder_1` data folder.

```{r}
load(file = file.path(params$google_drive_home,"03. Data Collection",params$data_folder_1,params$data_file_2))
```

Review.

```{r}
prep_employees %>% 
  skim()
```

## Select and unnest data {.tabset .tabset-pills}

### Filter to specification and cluster

Filter to data with `r params$in_number_of_clusters_1` clusters with specification `r params$in_specification`.

```{r}
clust_1 <- 
  kclusts %>% 
  filter(
    k == params$in_number_of_clusters_1 &
      specification == params$in_specification
  )
```

Filter to data with `r params$in_number_of_clusters_2` clusters with specification `r params$in_specification`.

```{r}
clust_2 <- 
  kclusts %>% 
  filter(
    k == params$in_number_of_clusters_2 &
      specification == params$in_specification
  )
```

### Unnest

Filter to data with `r params$in_number_of_clusters_2` clusters with specification `r params$in_specification`.

```{r}
clusters_1 <- 
  clust_1 %>%
  unnest(cols = c(tidied))

assignments_1 <- 
  clust_1 %>% 
  unnest(cols = c(augmented))

clusterings_1 <- 
  clust_1 %>%
  unnest(cols = c(glanced))

```

Check data.

```{r}
assignments_1 %>% 
  glimpse()
```

Filter to data with `r params$in_number_of_clusters_2` clusters with specification `r params$in_specification`.

```{r}
clusters_2 <- 
  clust_2 %>%
  unnest(cols = c(tidied))

assignments_2 <- 
  clust_2 %>% 
  unnest(cols = c(augmented))

clusterings_2 <- 
  clust_2 %>%
  unnest(cols = c(glanced))

```

Check data.

```{r}
assignments_2 %>% 
  glimpse()
```

### Join assignments to employee data

Function to rename `.cluster` column to include the number of clusters.

```{r}
rename_to_k <- function(in_data)
{
  tot_cluster <- in_data %>% select(k) %>% distinct() %>% pull()
  out_data <-
    in_data %>% rename_with( ~ glue::glue("{.x}_{tot_cluster}"), .cluster)
  return(out_data)
}
```

Apply function to assignments and join.

```{r}
assignments <-
  assignments_1 %>%
  rename_to_k() %>%
  select(person_id, starts_with(".cluster")) %>%
  full_join(assignments_2 %>%
              rename_to_k() %>%
              select(person_id, starts_with(".cluster")))
```
Check data.

```{r}
assignments %>% 
  glimpse()
```

Join to `prep_employees`

```{r}
d <-
  assignments %>%
  select(person_id, starts_with(".cluster")) %>%
  full_join(prep_employees) %>% 
  mutate(
    across(
      .cols = starts_with(".cluster"),
      .fns = ~fct_na_value_to_level(., level = "(Missing)")
    )
  )
```
Review data.

```{r}
d %>% 
  skim()
```

## Cluster descriptions {.tabset .tabset-pills}

Clusters 2(1): **Teleworkers**

Very low number of days reporting to the job site. Majority of time telework. Very few days co-located with at least one team member, a supervisor, or the whole team. Many days where the whole team is co-located, but not the majority. High number of working days.

Clusters 2(2): **Everyone else**

Higher number of days reporting to job site, but with large variance. Lower proportion of days teleworking, but still about a quarter of days. Again, higher variance of telework. Some days co-located with at least one team member and the majority of days working at the site. Many days working onsite with supervisor, but very few days where the whole team is co-located. Very few days where the whole team is co-located. Lower number of working days.

Clusters 5(1): **Teleworkers**

Very low number of days (~3) reporting to the job site. Most days (~119) employees telework. Hardly any days co-located with at least one team member (<1), supervisor (<1), or whole team (<1). Many days (~54) where entire team works remotely. Many working days (~122). Team is average size (~6).

Clusters 5(2): **Teleworkers that go to the office sometimes**

A few days (~14) working at the job site, but high variation. Majority of days(~98) employees telework. Some days co-located with at least one team member (~6) or supervisor (~3). Few days co-located with entire team (~1).  Some days (~19) where the entire team works remotely. Many working days (~112). Team is large (~9).

Clusters 5(3): **Hybrid workers**

Some days (~69) with employees working to the job site. Staff often telework (~48), but high variance. Many days co-located with at least one team member (~29) or with supervisor (~13). Few days co-located with entire team (~1). Few days (~5) where the whole team is remote.  Many working days (~115). Team is average size (~6).

Clusters 5(4): **Onsite workers**

Employees that report to the job site for most days (~97). Some days (~22) telework. Many days co-located with at least one team member (~74), supervisor (~60), or whole team (~17), but high variance. Few days (~4) where the whole team is remote, but high variance. Many working days (~119). Team is smaller (~5)

Clusters 5(5): **Non-workers**
Low number of days (~5) reporting to the site, but high variance. Some days (~27) telework, but high variance. Few days co-located with at least one team member (2)  or with supervisor (~1). Almost no days (<1) with the whole team co-located. Several days (~9) where the entire team works remotely. Few working days (~34) and high variance. Team is large (~9) and high variance.

## Work pattern statistics {.tabset .tabset-pills}

### Summary statistics

```{r}
d %>% 
  group_by(.cluster_2) %>% 
  skim(daily_status_reporting_to_job_site,
       daily_status_reporting_to_job_site_observation_prop,
       daily_status_teleworking,
       daily_status_teleworking_observation_prop,
       days_colocated_1,
       days_colocated_1_observation_prop,
       days_colocated_supervisor,
       days_colocated_supervisor_observation_prop,
       days_colocated_team,
       days_colocated_team_observation_prop,
       days_telework_team,
       days_telework_team_observation_prop,
       working_days,
       team_size_mode) %>% 
  as.data.frame()
```

```{r}
d %>% 
  group_by(.cluster_5) %>% 
  skim(daily_status_reporting_to_job_site,
       daily_status_reporting_to_job_site_observation_prop,
       daily_status_teleworking,
       daily_status_teleworking_observation_prop,
       days_colocated_1,
       days_colocated_1_observation_prop,
       days_colocated_supervisor,
       days_colocated_supervisor_observation_prop,
       days_colocated_team,
       days_colocated_team_observation_prop,
       days_telework_team,
       days_telework_team_observation_prop,
       working_days,
       team_size_mode) %>% 
  as.data.frame()
```

### Cluster summary statistics

#### Rule-based groups

##### Two clusters

```{r}
d %>% 
  group_by(.cluster_2) %>% 
  skim(starts_with("rule_based_"),
       -ends_with("working")) %>% 
  as.data.frame()
```

##### Five clusters

```{r}
d %>% 
  group_by(.cluster_5) %>% 
  skim(starts_with("rule_based_"),
       -ends_with("working")) %>% 
  as.data.frame()
```

```{r}
d %>% names()
```

#### Exiting employment and multiple supervisors

##### Two clusters

```{r}
d %>%
  group_by(.cluster_2) %>%
  skim(exit_employment,multiple_supervisors)
```

##### Five clusters

```{r}
d %>%
  group_by(.cluster_5) %>%
  skim(exit_employment,multiple_supervisors)
```

#### SSO

##### Two clusters

```{r}
d %>% 
  group_by(.cluster_2) %>% 
  count(sso) %>% 
  ungroup() %>% 
  left_join(
    d %>% 
  count(.cluster_2, name = "tot")
  ) %>% 
  mutate(
    prop = n/tot
  ) %>% 
  ggplot(aes(x = .cluster_2, y =  prop,fill = .cluster_2)) +
  geom_col(position = "dodge")  +
  facet_wrap(facets = vars(sso), scales = "free_y")
```

##### Five clusters

```{r}
d %>% 
  group_by(.cluster_5) %>% 
  count(sso) %>% 
  ungroup() %>% 
  left_join(
    d %>% 
  count(.cluster_5, name = "tot")
  ) %>% 
  mutate(
    prop = n/tot
  ) %>% 
  ggplot(aes(x = .cluster_5, y =  prop,fill = .cluster_5)) +
  geom_col(position = "dodge") +
  facet_wrap(facets = vars(sso)
             , scales = "free_y"
             )
```

#### GSA Region

##### Two clusters

```{r}
level_10 <- 
  d %>% 
  count(gsa_region) %>% 
  arrange(desc(n)) %>% 
  slice(10) %>% 
  pull(n)

d %>%
  # mutate(gsa_region = fct_lump_min(gsa_region, min = level_10)) %>%
  group_by(.cluster_2) %>%
  count(gsa_region) %>%
  ungroup() %>%
  left_join(d %>%
              count(.cluster_2, name = "tot")) %>%
  mutate(prop = n / tot) %>%
  ggplot(aes(x = .cluster_2, y =  prop, fill = .cluster_2)) +
  geom_col(position = "dodge")  +
  facet_wrap(facets = vars(gsa_region), scales = "free_y")
```

##### Five clusters

```{r}
level_10 <- 
  d %>% 
  count(gsa_region) %>% 
  arrange(desc(n)) %>% 
  slice(10) %>% 
  pull(n)

d %>%
  # mutate(gsa_region = fct_lump_min(gsa_region, min = level_10)) %>%
  group_by(.cluster_5) %>%
  count(gsa_region) %>%
  ungroup() %>%
  left_join(d %>%
              count(.cluster_5, name = "tot")) %>%
  mutate(prop = n / tot) %>%
  ggplot(aes(x = .cluster_5, y =  prop, fill = .cluster_5)) +
  geom_col(position = "dodge")  +
  facet_wrap(facets = vars(gsa_region), scales = "free_y")
```

#### Position category

##### Two clusters

```{r}
level_10 <- 
  d %>% 
  count(position_category_description) %>% 
  arrange(desc(n)) %>% 
  slice(10) %>% 
  pull(n)

d %>%
  # mutate(position_category_description = fct_lump_min(position_category_description, min = level_10)) %>%
  group_by(.cluster_2) %>%
  count(position_category_description) %>%
  ungroup() %>%
  left_join(d %>%
              count(.cluster_2, name = "tot")) %>%
  mutate(prop = n / tot) %>%
  ggplot(aes(x = .cluster_2, y =  prop, fill = .cluster_2)) +
  geom_col(position = "dodge")  +
  facet_wrap(facets = vars(position_category_description), scales = "free_y")
```

##### Five clusters

```{r}
level_10 <- 
  d %>% 
  count(position_category_description) %>% 
  arrange(desc(n)) %>% 
  slice(10) %>% 
  pull(n)

d %>%
  # mutate(position_category_description = fct_lump_min(position_category_description, min = level_10)) %>%
  group_by(.cluster_5) %>%
  count(position_category_description) %>%
  ungroup() %>%
  left_join(d %>%
              count(.cluster_5, name = "tot")) %>%
  mutate(prop = n / tot) %>%
  ggplot(aes(x = .cluster_5, y =  prop, fill = .cluster_5)) +
  geom_col(position = "dodge")  +
  facet_wrap(facets = vars(position_category_description), scales = "free_y")
```

#### Position category

##### Two clusters

```{r}
level_10 <- 
  d %>% 
  count(grade) %>% 
  arrange(desc(n)) %>% 
  slice(10) %>% 
  pull(n)

d %>%
  mutate(grade = fct_lump_min(grade, min = level_10)) %>%
  group_by(.cluster_2) %>%
  count(grade) %>%
  ungroup() %>%
  left_join(d %>%
              count(.cluster_2, name = "tot")) %>%
  mutate(prop = n / tot) %>%
  ggplot(aes(x = .cluster_2, y =  prop, fill = .cluster_2)) +
  geom_col(position = "dodge")  +
  facet_wrap(facets = vars(grade), scales = "free_y")
```

##### Five clusters

```{r}
level_10 <- 
  d %>% 
  count(grade) %>% 
  arrange(desc(n)) %>% 
  slice(10) %>% 
  pull(n)

d %>%
  mutate(grade = fct_lump_min(grade, min = level_10)) %>%
  group_by(.cluster_5) %>%
  count(grade) %>%
  ungroup() %>%
  left_join(d %>%
              count(.cluster_5, name = "tot")) %>%
  mutate(prop = n / tot) %>%
  ggplot(aes(x = .cluster_5, y =  prop, fill = .cluster_5)) +
  geom_col(position = "dodge")  +
  facet_wrap(facets = vars(grade), scales = "free_y")
```

## Cluster descriptions {.tabset .tabset-pills}

Clusters 2(1): **Teleworkers**

Very low number of days reporting to the job site. Majority of time telework. Very few days co-located with at least one team member, a supervisor, or the whole team. Many days where the whole team is co-located, but not the majority. High number of working days.

Position categories tend to be Offsite, Onsite Flexible, and Position Has Changed. A very small proportion of Onsite Required.

Few people exit employment (~3%) and some have multiple supervisors (~26%).

Disproportionately from FAS, GSA IT, OCFO, OGP, and OHRM, and fewer in PBS.

Disproportionately in Region 0.

Clusters 2(2): **Everyone else**

Higher number of days reporting to job site, but with large variance. Lower proportion of days teleworking, but still about a quarter of days. Again, higher variance of telework. Some days co-located with at least one team member and the majority of days working at the site. Many days working onsite with supervisor, but very few days where the whole team is co-located. Very few days where the whole team is co-located. Lower number of working days.

Position categories tend to be Onsite Required. All (Missing) and most are in this cluster.

Some people exit employment (~15%) and some have multiple supervisors (~22%).

Majority are from PBS.

Disproportionately in all regions apart from 0.

Clusters 5(1): **Teleworkers**

Very low number of days (~3) reporting to the job site. Most days (~119) employees telework. Hardly any days co-located with at least one team member (<1), supervisor (<1), or whole team (<1). Many days (~54) where entire team works remotely. Many working days (~122). Team is average size (~6).

Position categories tend to be Offsite and Position Has Changed. A small proportion of Onsite Required.

Almost no one (<.1%) exit employment. Some have multiple employers (20-30%).

High proportion in FAS, GSAIT, OCFO, OGP, OHRM, OSDBU and lower in PBS, OCE, IOA, and OMA.

Disproportionately in Region 0 and Region 2.

Clusters 5(2): **Teleworkers that go to the office sometimes**

A few days (~14) working at the job site, but high variation. Majority of days(~98) employees telework. Some days co-located with at least one team member (~6) or supervisor (~3). Few days co-located with entire team (~1).  Some days (~19) where the entire team works remotely. Many working days (~112). Team is large (~9).

Position categories tend to be Offsite and Position Has Changed. A small proportion of Onsite Required.

Very few people exit employment (~3%). Some have multiple employers (20-30%).

High proportion in FAS, GSAIT, OCFO, OCR, OGP, OHRM, OSDBU and lower in PBS, OCE, OSC, OSDBU, IOA, and OMA.

Disproportionately in Region 0.

Clusters 5(3): **Hybrid workers**

Some days (~69) with employees working to the job site. Staff often telework (~48), but high variance. Many days co-located with at least one team member (~29) or with supervisor (~13). Few days co-located with entire team (~1). Few days (~5) where the whole team is remote.  Many working days (~115). Team is average size (~6).

Position categories tend to be Onsite Flexible.

Very few people exit employment (~2%). Some have multiple employers (20-30%).

High proportion in IOA, OGC, OMA, OSC, and PBS and lower in FAS, GSA IT, OCE, OCFO, OCR, OGP, OHRM, and OSDBU.

Disproportionately in Region 1, Region 4, Region 5, Region 8, Region 9, and Region 10.

Clusters 5(4): **Onsite workers**

Employees that report to the job site for most days (~97). Some days (~22) telework. Many days co-located with at least one team member (~74), supervisor (~60), or whole team (~17), but high variance. Few days (~4) where the whole team is remote, but high variance. Many working days (~119). Team is smaller (~5).

Position categories tend to be Onsite Required.

Very few people exit employment (~2%). Some have multiple employers (20-30%).

High proportion in OAS, and PBS, and lower in FAS, GSA IT, IOA, OCE, OCFO, OCIA, OCR, OGC, OGP, OHRM, OMA, OSC, and OSDBU.

Disproportionately in Region 1, Region 2, Region 3, Region 6, Region 7, Region 8, Region 9, Region 10, and Region 11. Very low proportion in Region 0.

Clusters 5(5): **Non-workers**

Low number of days (~5) reporting to the site, but high variance. Some days (~27) telework, but high variance. Few days co-located with at least one team member (2)  or with supervisor (~1). Almost no days (<1) with the whole team co-located. Several days (~9) where the entire team works remotely. Few working days (~34) and high variance. Team is large (~9) and high variance.

Appear across all position categories and make up all of those that are (Missing).

Many people exit employment (~54%). Some have multiple employers (20-30%).

High proportion in FAS, GSA IT, OCE, OCIA, OCR, OSC, and OSDBU and lower in OMA.

Fairly even dispersion across regions.

```{r}
names(d)
```

# CONTINUE FROM HERE

```{r}
beepr::beep(sound = 5)
```

```{r}

```

```{r}
summaries <- function(in_data,...)
{
  in_data %>% 
  group_by(...) %>% 
  summarise(
    across(
      .cols = c(daily_status_teleworking,daily_status_reporting_to_job_site,
                daily_status_teleworking_working_prop,daily_status_reporting_to_job_site_working_prop,
                daily_status_teleworking_observation_prop,daily_status_reporting_to_job_site_observation_prop,
                starts_with("mon_daily_"),
                starts_with("tue_daily_"),
                starts_with("wed_daily_"),
                starts_with("thu_daily_"),
                starts_with("fri_daily_"),
                ends_with("_cor_telework"),
                ends_with("_cor_status"),
                ends_with("prop_pairs"),
                team_size_mode,
                starts_with("days_colocated"),
                starts_with("days_telework"),
                working_days
                ),
      .fns = ~mean(.,na.rm = TRUE)
    )
  ) %>% 
    mutate(
      statistic = "mean"
    )
}

filter_rule_base <- function(in_summaries_data,in_filter)
{
  filt_var_mod <- enquo(in_filter) 
  in_summaries_data %>% 
     dplyr::filter((!!filt_var_mod) == 1) %>% 
    pivot_longer(cols = !!filt_var_mod) %>% 
    mutate(
      temp = str_remove(name,"rule_based_"),
      temp = str_replace_all(temp,"_"," "),
      temp = str_to_sentence(temp),
      group = glue::glue("Rule-based: {temp}")
    ) %>% 
    select(-name, -value,-temp)
}

summaries_rule_based <- function(in_data,in_rule_based_var)
{
  d %>% 
    summaries({{in_rule_based_var}}) %>% 
    filter_rule_base({{in_rule_based_var}})
}

prep_a <- bind_rows(
  d %>%
    summaries() %>%
    mutate(group = "Whole cohort"),
  d %>%
    summaries(.cluster_2) %>%
    mutate(group = glue::glue("Cluster 2: {.cluster_2}")) %>%
    select(-.cluster_2),
  d %>%
    summaries(.cluster_5) %>%
    mutate(group = glue::glue("Cluster 5: {.cluster_5}")) %>%
    select(-.cluster_5),
  d %>% 
    summaries_rule_based(rule_based_all_remote),
  d %>% 
    summaries_rule_based(rule_based_primarily_teleworking_observation),
  d %>% 
    summaries_rule_based(rule_based_primarily_teleworking_working),
  d %>% 
    summaries_rule_based(rule_based_primarily_reporting_to_job_site_observation),
  d %>% 
    summaries_rule_based(rule_based_primarily_reporting_to_job_site_working),
  d %>% 
    summaries_rule_based(rule_based_mostly_teleworking_observation),
  d %>% 
    summaries_rule_based(rule_based_mostly_teleworking_working),
  d %>% 
    summaries_rule_based(rule_based_mostly_reporting_to_job_site_observation),
  d %>% 
    summaries_rule_based(rule_based_mostly_reporting_to_job_site_working),
  d %>% 
    summaries_rule_based(rule_based_position_category_missing),
  d %>% 
    summaries_rule_based(rule_based_position_category_onsite_flexible),
  d %>% 
    summaries_rule_based(rule_based_position_category_offsite),
  d %>% 
    summaries_rule_based(rule_based_position_category_onsite_required),
  d %>% 
    summaries_rule_based(rule_based_position_category_position_has_changed),
  d %>% 
    summaries_rule_based(rule_based_position_category_not_specified),
  d %>% 
    summaries_rule_based(rule_based_telework_onsite_minimum),
)
```

### Daily telework

```{r}
prep_a %>% 
  select(group,statistic,starts_with("daily_status_teleworking")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

### Daily telework by day of week

#### Monday

```{r}
prep_a %>% 
  select(group,statistic,starts_with("mon_daily_status_teleworking")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

#### Tuesday

```{r}
prep_a %>% 
  select(group,statistic,starts_with("tue_daily_status_teleworking")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

#### Wednesday

```{r}
prep_a %>% 
  select(group,statistic,starts_with("wed_daily_status_teleworking")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

#### Thursday

```{r}
prep_a %>% 
  select(group,statistic,starts_with("thu_daily_status_teleworking")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

#### Friday

```{r}
prep_a %>% 
  select(group,statistic,starts_with("fri_daily_status_teleworking")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

### Daily reporting to site

```{r}
prep_a %>% 
  select(group,statistic,starts_with("daily_status_reporting_to_job_site")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

### Team size

```{r}
prep_a %>% 
  select(group,statistic,starts_with("team_size")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

### Working days

```{r}
prep_a %>% 
  select(group,statistic,starts_with("working_days")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

### Co-located days with at least one team member

```{r}
prep_a %>% 
  select(group,statistic,starts_with("days_colocated_1")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

### Co-located days with at whole team

```{r}
prep_a %>% 
  select(group,statistic,starts_with("days_colocated_team_")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

### Co-located days with supervisor

```{r}
prep_a %>% 
  select(group,statistic,starts_with("days_colocated_supervisor_")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

### Days team teleworks

```{r}
prep_a %>% 
  select(group,statistic,starts_with("days_telework_team_")) %>% 
  rename_all(.funs = ~str_to_sentence(str_replace_all(.,"_"," "))) %>% 
  flextable() %>% 
   colformat_double(
    big.mark = ",", decimal.mark = ".",
    digits = 2,
    na_str = "na") %>% 
  autofit() %>% 
  theme_zebra()
```

The principle work pattern statistics are as follows:

-   Employee work pattern statistics:

    -   Proportion in employee time onsite by the day of the week

    -   Correlations of by employee time between the days of the week

    -   Number exiting employment

    -   Number changing supervisors

## Descriptive statistics {.tabset .tabset-pills}

 Employee descriptive statistics

    -   Number and proportion of employees by grade

    -   Number and proportion of employees by occupational series

    -   Number and proportion of employees by position category

    -   Number and proportion of employees by GSA region

    -   Number and proportion of employees by SSO

    -   Number and proportion of employees by state

    -   Number and proportion of employees by city

    -   Number and proportion of employees by supervisor level

## Distributions by cluster {.tabset .tabset-pills}

### Function

Function to create a density plot for a specific numeric variable.

```{r}
gg_density <- function(in_data,in_variable,in_cluster_number)
{
  in_variable_sym <- sym(in_variable)

  # spec <- unique(in_data$specification)
  # 
  # clus <- unique(in_data$k)

  if(in_cluster_number==2)
  {
    step_01 <-
      in_data %>%
      ggplot(aes(
        x = !!in_variable_sym,
        color = .cluster_2,
        fill = .cluster_2
      )) +
      geom_density(alpha = .25) +
      geomtextpath::geom_textdensity(
        aes(label = .cluster_2),
        hjust = "ymax",
        vjust = 1,
        text_only = TRUE,
        text_smoothing = 0,
        straight = TRUE
      )
  }

  if(in_cluster_number==5)
  {
    step_01 <-
      in_data %>%
      ggplot(aes(
        x = !!in_variable_sym,
        color = .cluster_5,
        fill = .cluster_5
      )) +
      geom_density(alpha = .25) +
      geomtextpath::geom_textdensity(
        aes(label = .cluster_5),
        hjust = "ymax",
        vjust = 1,
        text_only = TRUE,
        text_smoothing = 0,
        straight = TRUE
      )
  }

  step_01 +
    theme_fivethirtyeight() +
    labs(
      title = glue::glue("Histogram of `{in_variable}`"),
      caption = glue::glue("For {in_cluster_number} clusters.")
    ) +
    theme(
      legend.position = "none",
      panel.background = element_rect(fill = "white"),
      plot.background = element_rect(fill = "white"),
      # panel.border = element_blank()
    )
}
```

### List of numeric variables

```{r}
num_vars <- 
  d %>% 
  select_if(is.numeric) %>% 
  names()

num_vars
```

### Distributions

```{r}
num_vars[1:20] %>% 
  map(~gg_density(d,in_variable = .,in_cluster_number = 2))
```

### Telework

#### Number of days

```{r}
gg_density(d,"daily_status_teleworking", in_cluster_number = 2)
```

```{r}
gg_density(d,"daily_status_teleworking", in_cluster_number = 5)
```

#### Proportion of observation window

```{r}
gg_density(d,"daily_status_teleworking_observation_prop", in_cluster_number = 2)
```

```{r}
gg_density(d,"daily_status_teleworking_observation_prop", in_cluster_number = 5)
```

### Reporting to job site

#### Number of days

```{r}
gg_density(d,"daily_status_reporting_to_job_site", in_cluster_number = 2)
```

```{r}
gg_density(d,"daily_status_reporting_to_job_site", in_cluster_number = 5)
```

#### Proportion of observation window

```{r}
gg_density(d,"daily_status_reporting_to_job_site_observation_prop", in_cluster_number = 2)
```

```{r}
gg_density(d,"daily_status_reporting_to_job_site_observation_prop", in_cluster_number = 5)
```

### Collocation with at least one team member

#### Number of days

```{r}
gg_density(d,"days_colocated_1", in_cluster_number = 2)
```

```{r}
gg_density(d,"days_colocated_1", in_cluster_number = 5)
```

#### Proportion of observation window

```{r}
gg_density(d,"days_colocated_1_observation_prop", in_cluster_number = 2)
```

```{r}
gg_density(d,"days_colocated_1_observation_prop", in_cluster_number = 5)
```

### Collocation with supervisor

#### Number of days

```{r}
gg_density(d,"days_colocated_supervisor", in_cluster_number = 2)
```

```{r}
gg_density(d,"days_colocated_supervisor", in_cluster_number = 5)
```

#### Proportion of observation window

```{r}
gg_density(d,"days_colocated_supervisor_observation_prop", in_cluster_number = 2)
```

```{r}
gg_density(d,"days_colocated_supervisor_observation_prop", in_cluster_number = 5)
```

### Collocation with teamn

#### Number of days

```{r}
gg_density(d,"days_colocated_team", in_cluster_number = 2)
```

```{r}
gg_density(d,"days_colocated_team", in_cluster_number = 5)
```

#### Proportion of observation window

```{r}
gg_density(d,"days_colocated_team_observation_prop", in_cluster_number = 2)
```

```{r}
gg_density(d,"days_colocated_team_observation_prop", in_cluster_number = 5)
```

### Whole team telewords

#### Number of days

```{r}
gg_density(d,"days_colocated_team", in_cluster_number = 2)
```

```{r}
gg_density(d,"days_colocated_team", in_cluster_number = 5)
```

#### Proportion of observation window

```{r}
gg_density(d,"days_colocated_team_observation_prop", in_cluster_number = 2)
```

```{r}
gg_density(d,"days_colocated_team_observation_prop", in_cluster_number = 5)
```

### Working days

#### Number of days

```{r}
gg_density(d,"working_days", in_cluster_number = 2)
```

```{r}
gg_density(d,"working_days", in_cluster_number = 5)
```

### Team size

#### Number of employees

```{r}
gg_density(d,"team_size_mode", in_cluster_number = 2)
```

```{r}
gg_density(d,"team_size_mode", in_cluster_number = 5)
```

#### Proportion of observation window

```{r}
gg_density(d,"daily_status_reporting_to_job_site_observation_prop", in_cluster_number = 2)
```

```{r}
gg_density(d,"daily_status_reporting_to_job_site_observation_prop", in_cluster_number = 5)
```

## Descriptive

The descriptive statistics are:

-  
The employees statistics will be calculated for the following groups:

1.  Whole cohort of employees
2.  By rule-based categories of work pattern
3.  By cluster defined categories of work pattern

## Scale data {.tabset .tabset-pills}

Scaling the data in two ways:

1.  Normalize - Data will have the mean that that element subtracted and
    then divided by the standard deviation, using the `scale()`
    function[^2]

2.  Rank - The rank of each element will be used instead of the element.

[^2]: [Documentation for the scale
    function.](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/scale)

```{r}
d <- 
  prep_employees %>% 
  mutate(
    across(
      .cols = c(daily_status_reporting_to_job_site_working_prop,
                daily_status_teleworking_working_prop,
                daily_status_reporting_to_job_site_observation_prop,
                daily_status_teleworking_observation_prop,
                starts_with("mon_"),
                starts_with("tue_"),
                starts_with("wed_"),
                starts_with("thu_"),
                starts_with("fri_"),
                working_days,
                team_size_mode,
                starts_with("days_colocated_"),
                starts_with("days_telework_"),
                -days_colocated_1,
                -days_colocated_supervisor,
                -days_colocated_team,
                -days_telework_team
                ),
      .fns = list(scale = scale,rank = rank),
      .names = "{.col}_{.fn}"
    )
  )
```

Review data.

```{r}
d %>% 
  skim()
```

## Create clusters {.tabset .tabset-pills}

### Functions

#### Selecting data for cluster analysis

```{r}
#' Selects the correct columns for employee cluster analysis
#'
#' @param in_data Data to use
#' @param in_scale_or_rank Option for which scaling procedure: "scale" (normalization) or "rank" (rank order)
#' @param in_working_or_observation_window Option for proportion denominator: "observation" (proportion of entire observation window) or "working" (proportion of period employ was working onsite or via telework)
#' @param in_weekday_correlation Option for inter weekday relationship: "correlation telework" (telework correlation between days of the week), "correlation status" (work status correlation based on ordered work status by days of the week), or "proportion pairs" (proportion of telework in pairs)
#'
#' @return Dataframe with only selected columns and dropped null observations
#' @export
#'
#' @examples
select_cluster_cols <-
  function(in_data,
           in_scale_or_rank,
           in_working_or_observation_window,
           in_weekday_correlation)
  {
    stopifnot(in_scale_or_rank %in% c("scale", "rank"))
    stopifnot(in_working_or_observation_window %in% c("observation", "working"))
    stopifnot(in_weekday_correlation %in% c("correlation telework", "correlation status","proportion pairs"))

    if (in_scale_or_rank == "scale" &
        in_working_or_observation_window == "observation" &
        in_weekday_correlation == "correlation telework")
    {
      step_01 <-
        in_data %>%
        select(
          person_id,
          ends_with("_scale"),
          -contains("_working_prop_"),
          -contains("_cor_status_"),
          -contains("_prop_pairs_"),
        ) %>%
        drop_na()
    }

    if (in_scale_or_rank == "rank" &
        in_working_or_observation_window == "observation" &
        in_weekday_correlation == "correlation telework")
    {
      step_01 <-
        in_data %>%
        select(
          person_id,
          ends_with("_rank"),
          -contains("_working_prop_"),
          -contains("_cor_status_"),
          -contains("_prop_pairs_"),
        ) %>%
        drop_na()
    }

    if (in_scale_or_rank == "scale" &
        in_working_or_observation_window == "working" &
        in_weekday_correlation == "correlation telework")
    {
      step_01 <-
        in_data %>%
        select(
          person_id,
          ends_with("_scale"),
          -contains("_observation_"),
          -contains("_cor_status_"),
          -contains("_prop_pairs_"),
        ) %>%
        drop_na()
    }

    if (in_scale_or_rank == "rank" &
        in_working_or_observation_window == "working" &
        in_weekday_correlation == "correlation telework")
    {
      step_01 <-
        in_data %>%
        select(
          person_id,
          ends_with("_rank"),
          -contains("_observation_"),
          -contains("_cor_status_"),
          -contains("_prop_pairs_"),
        ) %>%
        drop_na()
    }

    if (in_scale_or_rank == "scale" &
        in_working_or_observation_window == "observation" &
        in_weekday_correlation == "correlation status")
    {
      step_01 <-
        in_data %>%
        select(
          person_id,
          ends_with("_scale"),
          -contains("_working_prop_"),
          -contains("_cor_telework_"),
          -contains("_prop_pairs_"),
        ) %>%
        drop_na()
    }

    if (in_scale_or_rank == "rank" &
        in_working_or_observation_window == "observation" &
        in_weekday_correlation == "correlation status")
    {
      step_01 <-
        in_data %>%
        select(
          person_id,
          ends_with("_rank"),
          -contains("_working_prop_"),
          -contains("_cor_telework_"),
          -contains("_prop_pairs_"),
        ) %>%
        drop_na()
    }

    if (in_scale_or_rank == "scale" &
        in_working_or_observation_window == "working" &
        in_weekday_correlation == "correlation status")
    {
      step_01 <-
        in_data %>%
        select(
          person_id,
          ends_with("_scale"),
          -contains("_observation_"),
          -contains("_cor_telework_"),
          -contains("_prop_pairs_"),
        ) %>%
        drop_na()
    }

    if (in_scale_or_rank == "rank" &
        in_working_or_observation_window == "working" &
        in_weekday_correlation == "correlation status")
    {
      step_01 <-
        in_data %>%
        select(
          person_id,
          ends_with("_rank"),
          -contains("_observation_"),
          -contains("_cor_telework_"),
          -contains("_prop_pairs_"),
        ) %>%
        drop_na()
    }

    if (in_scale_or_rank == "scale" &
        in_working_or_observation_window == "observation" &
        in_weekday_correlation == "proportion pairs")
    {
      step_01 <-
        in_data %>%
        select(
          person_id,
          ends_with("_scale"),
          -contains("_working_prop_"),
          -contains("_cor_status_"),
          -contains("cor_telework"),
        ) %>%
        drop_na()
    }

    if (in_scale_or_rank == "rank" &
        in_working_or_observation_window == "observation" &
        in_weekday_correlation == "proportion pairs")
    {
      step_01 <-
        in_data %>%
        select(
          person_id,
          ends_with("_rank"),
          -contains("_working_prop_"),
          -contains("_cor_status_"),
          -contains("cor_telework"),
        ) %>%
        drop_na()
    }

    if (in_scale_or_rank == "scale" &
        in_working_or_observation_window == "working" &
        in_weekday_correlation == "proportion pairs")
    {
      step_01 <-
        in_data %>%
        select(
          person_id,
          ends_with("_scale"),
          -contains("_observation_"),
          -contains("_cor_status_"),
          -contains("cor_telework"),
        ) %>%
        drop_na()
    }

    if (in_scale_or_rank == "rank" &
        in_working_or_observation_window == "working" &
        in_weekday_correlation == "proportion pairs")
    {
      step_01 <-
        in_data %>%
        select(
          person_id,
          ends_with("_rank"),
          -contains("_observation_"),
          -contains("_cor_status_"),
          -contains("cor_telework"),
        ) %>%
        drop_na()
    }

    out <- 
      step_01 %>% 
      as_tibble()

    return(out)
  }
```

#### Making clusters of different sizes

```{r}
#' Make employee clusters of a variety of sizes and different specifications
#'
#' @param in_data Data to use
#' @param in_scale_or_rank Option for which scaling procedure: "scale" (normalization) or "rank" (rank order)
#' @param in_working_or_observation_window Option for proportion denominator: "observation" (proportion of entire observation window) or "working" (proportion of period employ was working onsite or via telework)
#' @param in_weekday_correlation Option for inter weekday relationship: "correlation telework" (telework correlation between days of the week), "correlation status" (work status correlation based on ordered work status by days of the week), or "proportion pairs" (proportion of telework in pairs)
#' @param in_number_of_clusters_start Minimum number of clusters to test
#' @param in_number_of_clusters_end Maximum number of clusters to test
#'
#' @return Dataframe of clusters and associatied statistics
#' @export
#'
#' @examples
make_clusters <-
  function(in_data,
           in_scale_or_rank,
           in_working_or_observation_window,
           in_weekday_correlation,
           in_number_of_clusters_start,
           in_number_of_clusters_end)
  {
    stopifnot(in_scale_or_rank %in% c("scale", "rank"))
    stopifnot(in_working_or_observation_window %in% c("observation", "working"))
    stopifnot(
      in_weekday_correlation %in% c(
        "correlation telework",
        "correlation status",
        "proportion pairs"
      )
    )
    stopifnot(in_number_of_clusters_start >= 1)
    stopifnot(in_number_of_clusters_start <= in_number_of_clusters_end)

    step_01 <-
      in_data %>%
      select_cluster_cols(
        in_scale_or_rank = in_scale_or_rank,
        in_working_or_observation_window = in_working_or_observation_window,
        in_weekday_correlation = in_weekday_correlation
      )

    print(glue::glue("Review data for step 1. {Sys.time()}"))

    step_01 %>%
      skim() %>%
      print()

    print(glue::glue("Remove person_id. {Sys.time()}"))

    step_02 <-
      step_01 %>%
      select(-person_id)

    print(glue::glue("Review data for step 2. {Sys.time()}"))

    step_02 %>%
      skim() %>%
      print()

    print(
      glue::glue(
        "Create clusters of the data from {in_number_of_clusters_start} to {in_number_of_clusters_end}. {Sys.time()}"
      )
    )

    out <-
      tibble(k = in_number_of_clusters_start:in_number_of_clusters_end) %>%
      mutate(
        kclust = map(k, ~ kmeans(step_02, .x)),
        tidied = map(kclust, tidy),
        glanced = map(kclust, glance),
        augmented = map(kclust, augment, step_01)
      )

    return(out)
  }
```

#### Making clusters of different sizes, but only for `d` and adding specification

```{r}
make_clusters_d <- function(in_specification,
                            in_scale_or_rank,
                            in_working_or_observation_window,
                            in_weekday_correlation,
                            in_number_of_clusters_start,
                            in_number_of_clusters_end)
{
  make_clusters(
    in_data = d,
    in_scale_or_rank = in_scale_or_rank,
    in_working_or_observation_window = in_working_or_observation_window,
    in_weekday_correlation = in_weekday_correlation,
    in_number_of_clusters_start = in_number_of_clusters_start,
    in_number_of_clusters_end = in_number_of_clusters_end
  ) %>%
    mutate(specification = in_specification)
}
```

### Create clusters

#### Create dataframe of all permutations

This table shows all permutations of scaling (normalization, or rank), denominators for proportions (observation window or working days), and types of inter day correlations.

```{r}
specs <- 
  tibble(
  in_scale_or_rank = c(rep("scale", 6), rep("rank", 6)),
  in_working_or_observation_window = rep(c(rep("observation", 3), rep("working", 3)), 2),
  in_weekday_correlation = rep(
    c(
      "correlation telework",
      "correlation status",
      "proportion pairs"
    ),
    4
  ),
  in_number_of_clusters_start = rep(params$in_number_of_clusters_start, 12),
  in_number_of_clusters_end = rep(params$in_number_of_clusters_end, 12)
) %>% 
  unite(col = "in_specification",everything(),remove = FALSE)
```

Review data.

```{r}
specs
```

#### Create clusters

Iterate over `specs` creating and collecting dataframe of clusters from all of the above specifications.

```{r}
kclusts <- 
  specs %>%
  pmap_dfr(make_clusters_d)
```

## Save data {.tabset .tabset-pills}

File name.

```{r}
file_name <- glue::glue("clusters-employee-{Sys.Date()}.rdata")
file_name
```

```{r}
save(kclusts, file = file.path(params$google_drive_home,"03. Data Collection","prepared-data",file_name))
```

Remove all data elements.

```{r}
rm(list = ls())
gc()
```

## Extraction time {.tabset .tabset-pills}

```{r}
tictoc::toc()
```