### 2023-08-22 04. Analysis/scripts/04-01-reporting-presentation.Rmd
Create 04-01-reporting-presentation.Rmd

---
title: "slides for lab"
author: "Ben Jaques-Leslie"
date: "2023-08-22"
output: powerpoint_presentation
params:
  author: "Ben Jaques-Leslie"
  project_number: 2312
  project_name: "Occupancy Patterns"
  google_drive_home: "G:/Shared drives/GSA Evaluation Division/Projects/1.0 Real Estate Solutions (LQ1)/2312 Work Patterns"
  data_folder_1: "prepared-data"
  data_file_1: "clusters-employee-2023-08-09.rdata"
  data_file_2: "prep_employees_2023-08-09.rdata"
  in_number_of_clusters_1: 2
  in_number_of_clusters_2: 5
  in_specification: "scale_observation_none_1_20_TRUE"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      fig.width = 8,
                      fig.height = 4,
                      dpi = 300)
options(scipen=999)
tictoc::tic()
if(!("arsenal" %in% installed.packages()[,"Package"]))
{
  install.packages("arsenal")
}
if(!("beepr" %in% installed.packages()[,"Package"]))
{
  install.packages("beepr")
}
if(!("geomtextpath" %in% installed.packages()[,"Package"]))
{
  install.packages("geomtextpath")
}
if(!("oesrrr" %in% installed.packages()[,"Package"]))
{
  devtools::install_github(repo = "GSA/oesrrr")
}
if(!("oesrrr" %in% installed.packages()[,"Package"]))
{
  devtools::install_github(repo = "GSA/oesrrr")
}
library(oesrrr)
library(tidyverse)
library(janitor)
library(skimr)
library(DataExplorer)
library(readr)
library(arsenal)
library(Hmisc)
library(readxl)
library(flextable)
library(ggthemes)
library(tidymodels)
library(patchwork)
```

```{r include=FALSE}
load(file = file.path(params$google_drive_home,"03. Data Collection",params$data_folder_1,params$data_file_1))
```

```{r include=FALSE}
load(file = file.path(params$google_drive_home,"03. Data Collection",params$data_folder_1,params$data_file_2))
```

```{r include=FALSE}
clust_1 <- 
  kclusts %>% 
  filter(
    k == params$in_number_of_clusters_1 &
      specification == params$in_specification
  )
```

```{r include=FALSE}
clust_2 <- 
  kclusts %>% 
  filter(
    k == params$in_number_of_clusters_2 &
      specification == params$in_specification
  )
```

```{r include=FALSE}
clusters_1 <- 
  clust_1 %>%
  unnest(cols = c(tidied))

assignments_1 <- 
  clust_1 %>% 
  unnest(cols = c(augmented))

clusterings_1 <- 
  clust_1 %>%
  unnest(cols = c(glanced))

```

```{r include=FALSE}
clusters_2 <- 
  clust_2 %>%
  unnest(cols = c(tidied))

assignments_2 <- 
  clust_2 %>% 
  unnest(cols = c(augmented))

clusterings_2 <- 
  clust_2 %>%
  unnest(cols = c(glanced))

```

```{r include=FALSE}
rename_to_k <- function(in_data)
{
  tot_cluster <- in_data %>% select(k) %>% distinct() %>% pull()
  out_data <-
    in_data %>% rename_with( ~ glue::glue("{.x}_{tot_cluster}"), .cluster)
  return(out_data)
}
```

```{r include=FALSE}
assignments <-
  assignments_1 %>%
  rename_to_k() %>%
  select(person_id, starts_with(".cluster")) %>%
  full_join(assignments_2 %>%
              rename_to_k() %>%
              select(person_id, starts_with(".cluster")))
```

```{r include=FALSE}
d <-
  assignments %>%
  select(person_id, starts_with(".cluster")) %>%
  full_join(prep_employees) %>% 
  mutate(
    across(
      .cols = starts_with(".cluster"),
      .fns = ~fct_na_value_to_level(., level = "(Missing)")
    )
  )
```

```{r include=FALSE}
d <- d %>% 
  mutate(
    .cluster_5 = fct_recode(.cluster_5,
                            "Teleworkers" = "1",
                            "Office visitors" = "2",
                            "Hybrid workers" = "3",
                            "Onsite workers" = "4",
                            "Disconnected workers" = "5"),
    .cluster_5 = fct_relabel(.cluster_5, ~str_wrap(.x,width = 10))
  )
```

# How much telework?

```{r include=FALSE}
prep_01 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  summarise(num = mean(daily_status_teleworking)) %>% 
  ungroup()

prep_02 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  summarise(xPos = max(daily_status_teleworking),
            yPos = max((density(daily_status_teleworking))$y))

prep_02 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  summarise(y = max(density(daily_status_teleworking)$y),
            x = density(daily_status_teleworking)$x[density(daily_status_teleworking)$y == max(density(daily_status_teleworking)$y)]
  )
```

```{r}
p_1 <- 
  prep_01 %>% 
  ggplot(
    aes(x = .cluster_5, y = num, fill = .cluster_5)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::comma) +
  ylab("Mean days of telework") +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_2 <- 
  d %>% 
  ggplot(
    aes(x = daily_status_teleworking, color = .cluster_5, fill = .cluster_5)
  ) +
  geom_density(alpha = .5) +
  # geom_vline(data = prep_01, aes(xintercept = num, color = .cluster_5)) +
  geom_text(data=prep_02, aes(x=x, y=y, label=.cluster_5), size = 3, hjust = "inward") +
  xlab("Days of telework") +
  ylab("Density") +
  theme_light() +
  theme(legend.position = "None")

p_1|p_2
```

# How much onsite work?

```{r include=FALSE}
prep_01 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  summarise(num = mean(daily_status_reporting_to_job_site)) %>% 
  ungroup()

prep_02 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  summarise(xPos = max(daily_status_reporting_to_job_site),
            yPos = max((density(daily_status_reporting_to_job_site))$y))

prep_02 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  summarise(y = max(density(daily_status_reporting_to_job_site)$y),
            x = density(daily_status_reporting_to_job_site)$x[density(daily_status_reporting_to_job_site)$y == max(density(daily_status_reporting_to_job_site)$y)]
  )
```

```{r}
p_1 <- 
  prep_01 %>% 
  ggplot(
    aes(x = .cluster_5, y = num, fill = .cluster_5)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::comma) +
  ylab("Mean days of onsite work") +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_2 <- 
  d %>% 
  ggplot(
    aes(x = daily_status_reporting_to_job_site, color = .cluster_5, fill = .cluster_5)
  ) +
  geom_density(alpha = .5) +
  # geom_vline(data = prep_01, aes(xintercept = num, color = .cluster_5)) +
  geom_text(data=prep_02, aes(x=x, y=y, label=.cluster_5), size = 3, hjust = "inward") +
  xlab("Days of onsite work") +
  ylab("Density") +
  theme_light() +
  theme(legend.position = "None")

p_1|p_2
```

# Disconnected workers: Working days

```{r include=FALSE}
prep_01 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  summarise(num = mean(working_days)) %>% 
  ungroup()

prep_02 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  summarise(xPos = max(working_days),
            yPos = max((density(working_days))$y))

prep_02 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  summarise(y = max(density(working_days)$y),
            x = density(working_days)$x[density(working_days)$y == max(density(working_days)$y)]
  )
```

```{r}
p_1 <- 
  prep_01 %>% 
  ggplot(
    aes(x = .cluster_5, y = num, fill = .cluster_5)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::comma) +
  ylab("Mean working days") +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_2 <- 
  d %>% 
  ggplot(
    aes(x = working_days, color = .cluster_5, fill = .cluster_5)
  ) +
  geom_density(alpha = .5) +
  # geom_vline(data = prep_01, aes(xintercept = num, color = .cluster_5)) +
  geom_text(data=prep_02, aes(x=x, y=y, label=.cluster_5), size = 3, hjust = "inward") +
  xlab("Working days") +
  ylab("Density") +
  theme_light() +
  theme(legend.position = "None")

p_1|p_2
```

# Disconnected workers: Exiting employment

```{r include=FALSE}
prep_01 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  summarise(num = sum(exit_employment)) %>% 
  ungroup()

prep_02 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  count() %>% 
  ungroup()

prep_03 <- 
  prep_01 %>% 
  full_join(prep_02) %>% 
  mutate(
    num = num/n
  )
```

```{r}
p_1 <-
  prep_03 %>%
  ggplot(aes(x = .cluster_5, y = num, fill = .cluster_5)) +
  geom_col() +
  ylab("Percent exiting employment") +
  scale_y_continuous(labels = scales::percent) +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8)
  )

p_1
```

# Position categories

```{r}
prep_d_01 <- 
  d %>% 
  group_by(position_category_description) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n))

prep_v_01 <- 
  prep_d_01 %>% 
  slice(3) %>% 
  pull(n)

prep_d_02 <-
  d %>%
  mutate(position_category_description = fct_lump_min(position_category_description, min = prep_v_01),
         position_category_description = fct_infreq(position_category_description)) %>%
  group_by(position_category_description) %>%
  count() %>%
  ungroup() %>%
  arrange(desc(n)) %>% 
  rename(num = n)
```

```{r}
p_1 <- 
  prep_d_02 %>% 
  ggplot(
    aes(x = position_category_description, y = num, fill = position_category_description)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::comma) +
  ylab("Employees") +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_1
```

# Rule-based groups

```{r}
prep_d_01 <- 
  d %>%
  summarise(across(
    .cols = starts_with("rule_based"),
    .fns = ~ sum(., rm.na = T)
  )) %>%
  select(-contains("_position_category_")) %>%
  select(-ends_with("observation")) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  select(-rowname) %>%
  mutate(
    name = str_remove(name, "rule_based_"),
    name = str_remove(name, "_working"),
    name = str_replace_all(name, "_", " "),
    name = str_to_sentence(name),
    name = fct_relevel(
      name,
      "Primarily teleworking",
      "Mostly teleworking",
      "Mostly reporting to job site",
      "Primarily reporting to job site",
      "All remote",
      "Telework onsite minimum"
    ),
    name = fct_relabel(name, ~str_wrap(.,width = 20))
  ) %>% 
  arrange(name)

prep_d_02 <- 
  prep_d_01 %>% 
  mutate(tot = d %>% count() %>% pull(),
         prop = value/tot,
         text = scales::percent(prop, accuracy = .1))
```

```{r}
p_1 <- 
  prep_d_01 %>% 
  ggplot(
    aes(x = name, y = value, fill = name)
  ) +
  geom_col() +
  geom_text(data = prep_d_02, aes(x = name, y = .5*value, label = text), color = "white") +
  scale_y_continuous(labels = scales::comma) +
  ylab("Employees") +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_1
```

# Clusters

```{r}
prep_d_01 <- 
  d %>%
  group_by(.cluster_5) %>% 
  count() %>% 
  ungroup()

prep_d_02 <- 
  prep_d_01 %>% 
  mutate(tot = d %>% count() %>% pull(),
         prop = n/tot,
         text = scales::percent(prop, accuracy = .1))
```

```{r}
p_1 <- 
  prep_d_01 %>% 
  ggplot(
    aes(x = .cluster_5, y = n, fill = .cluster_5)
  ) +
  geom_col() +
  geom_text(data = prep_d_02, aes(x = .cluster_5, y = .5*n, label = text), color = "white") +
  scale_y_continuous(labels = scales::comma) +
  ylab("Employees") +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_1
```