### 2023-09-06 04. Analysis/scripts/04-01-reporting-presentation.Rmd
Updates for the abstract

---
title: "slides for lab"
author: "Ben Jaques-Leslie"
date: "2023-08-22"
output: powerpoint_presentation
params:
  author: "Ben Jaques-Leslie"
  project_number: 2312
  project_name: "Occupancy Patterns"
  google_drive_home: "G:/Shared drives/GSA Evaluation Division/Projects/1.0 Real Estate Solutions (LQ1)/2312 Work Patterns"
  data_folder_1: "prepared-data"
  data_file_1: "clusters-employee-2023-09-06.rdata"
  data_file_2: "prep_employees_2023-09-06.rdata"
  in_number_of_clusters_1: 2
  in_number_of_clusters_2: 5
  in_specification: "scale_observation_none_1_20_TRUE"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      fig.width = 8,
                      fig.height = 4,
                      dpi = 300)
options(scipen=999)
tictoc::tic()
if(!("ggalluvial" %in% installed.packages()[,"Package"]))
{
  install.packages("ggalluvial")
}
if(!("arsenal" %in% installed.packages()[,"Package"]))
{
  install.packages("arsenal")
}
if(!("beepr" %in% installed.packages()[,"Package"]))
{
  install.packages("beepr")
}
if(!("geomtextpath" %in% installed.packages()[,"Package"]))
{
  install.packages("geomtextpath")
}
if(!("oesrrr" %in% installed.packages()[,"Package"]))
{
  devtools::install_github(repo = "GSA/oesrrr")
}
if(!("oescolorrrs" %in% installed.packages()[,"Package"]))
{
  devtools::install_github(repo = "GSA/oescolorrrs")
}
if(!("ggrain" %in% installed.packages()[,"Package"]))
{
  install.packages("ggrain")
}
# install.packages("tigris")
library(oesrrr)
library(tidyverse)
library(janitor)
library(skimr)
library(DataExplorer)
library(readr)
library(arsenal)
library(Hmisc)
library(readxl)
library(flextable)
library(ggthemes)
library(tidymodels)
library(patchwork)
library(sf)
library(tigris)
library(ggrepel)
library(ggalluvial)
library(ggrain)
options(tigris_class = "sf")
options(tigris_use_cache = TRUE)
```

```{r include=FALSE}
load(file = file.path(params$google_drive_home,"03. Data Collection",params$data_folder_1,params$data_file_1))
```

```{r include=FALSE}
load(file = file.path(params$google_drive_home,"03. Data Collection",params$data_folder_1,params$data_file_2))
```

```{r include=FALSE}
clust_1 <- 
  kclusts %>% 
  filter(
    k == params$in_number_of_clusters_1 &
      specification == params$in_specification
  )
```

```{r include=FALSE}
clust_2 <- 
  kclusts %>% 
  filter(
    k == params$in_number_of_clusters_2 &
      specification == params$in_specification
  )
```

```{r include=FALSE}
clusters_1 <- 
  clust_1 %>%
  unnest(cols = c(tidied))

assignments_1 <- 
  clust_1 %>% 
  unnest(cols = c(augmented))

clusterings_1 <- 
  clust_1 %>%
  unnest(cols = c(glanced))

```

```{r include=FALSE}
clusters_2 <- 
  clust_2 %>%
  unnest(cols = c(tidied))

assignments_2 <- 
  clust_2 %>% 
  unnest(cols = c(augmented))

clusterings_2 <- 
  clust_2 %>%
  unnest(cols = c(glanced))

```

```{r include=FALSE}
rename_to_k <- function(in_data)
{
  tot_cluster <- in_data %>% select(k) %>% distinct() %>% pull()
  out_data <-
    in_data %>% rename_with( ~ glue::glue("{.x}_{tot_cluster}"), .cluster)
  return(out_data)
}
```

```{r include=FALSE}
assignments <-
  assignments_1 %>%
  rename_to_k() %>%
  select(person_id, starts_with(".cluster")) %>%
  full_join(assignments_2 %>%
              rename_to_k() %>%
              select(person_id, starts_with(".cluster")))
```

```{r include=FALSE}
d <-
  assignments %>%
  select(person_id, starts_with(".cluster")) %>%
  full_join(prep_employees) %>% 
  mutate(
    across(
      .cols = starts_with(".cluster"),
      .fns = ~fct_na_value_to_level(., level = "(Missing)")
    )
  )
```

```{r include=FALSE}
d <- d %>% 
  mutate(
    .cluster_5 = fct_recode(.cluster_5,
                            "Telecommuters" = "1",
                            "Office visitors" = "2",
                            "Hybrid workers" = "3",
                            "In-office workers" = "4",
                            "Disconnected workers" = "5"),
    .cluster_5 = fct_relabel(.cluster_5, ~str_wrap(.x,width = 10))
  )
```

# What is the GSA priority?

-   Telework has become very common

-   Much of GSA staff telework

-   How employees use telework is not clear

-   Defining and describing different work patterns may help to understand organizational health

-   By understanding how staff work (remote, hybrid, or onsite), different strategies may be developed to improve productivity and staff engagement

# How did the evaluation work?

-   Data:

    -   GSA check-in data, GSA staff roster

    -   March - September 2023

-   All GSA staff

    -   Categorized work patterns by remote or onsite

    -   Rules: Mostly remote or onsite, staff category (flexible, onsite, etc), flexible onsite

    -   Clusters: K-means clusters defined using frequency of work location status, collocation with team members, and team remote work

# Position categories

```{r}
prep_d_01 <- 
  d %>% 
  group_by(position_category_description) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n))

prep_v_01 <- 
  prep_d_01 %>% 
  slice(3) %>% 
  pull(n)

prep_d_02 <-
  d %>%
  mutate(position_category_description = fct_lump_min(position_category_description, min = prep_v_01),
         position_category_description = fct_infreq(position_category_description)) %>%
  group_by(position_category_description) %>%
  count() %>%
  ungroup() %>%
  arrange(desc(n)) %>% 
  rename(num = n)
```

```{r}
p_1 <- 
  prep_d_02 %>% 
  ggplot(
    aes(x = position_category_description, y = num, fill = position_category_description)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::comma) +
  ylab("Employees") +
  theme_light() +
  oescolorrrs::scale_fill_oes() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_1
```

# Rule-based groups

```{r}
prep_d_01 <- 
  d %>%
  summarise(across(
    .cols = starts_with("rule_based"),
    .fns = ~ sum(., rm.na = T)
  )) %>%
  select(-contains("_position_category_")) %>%
  select(-ends_with("observation")) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  select(-rowname) %>%
  mutate(
    name = str_remove(name, "rule_based_"),
    name = str_remove(name, "_working"),
    name = str_replace_all(name, "_", " "),
    name = str_to_sentence(name),
    name = fct_relevel(
      name,
      "Primarily teleworking",
      "Mostly teleworking",
      "Mostly reporting to job site",
      "Primarily reporting to job site",
      "All remote",
      "Telework onsite minimum"
    ),
    name = fct_relabel(name, ~str_wrap(.,width = 20))
  ) %>% 
  arrange(name)

prep_d_02 <- 
  prep_d_01 %>% 
  mutate(tot = d %>% count() %>% pull(),
         prop = value/tot,
         text = scales::percent(prop, accuracy = .1))
```

```{r}
p_1 <- 
  prep_d_01 %>% 
  ggplot(
    aes(x = name, y = value, fill = name)
  ) +
  geom_col() +
  geom_text(data = prep_d_02, aes(x = name, y = .5*value, label = text), color = "white") +
  scale_y_continuous(labels = scales::comma) +
  ylab("Employees") +
  theme_light() +
  oescolorrrs::scale_fill_oes() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_1
```

# Clusters

1.  Telecommuters

2.  Office visitors

3.  Hybrid workers

4.  Onsite workers

5.  Disconnected workers

# Clusters

```{r}
prep_d_01 <- 
  d %>%
  group_by(.cluster_5) %>% 
  count() %>% 
  ungroup()

prep_d_02 <- 
  prep_d_01 %>% 
  mutate(tot = d %>% count() %>% pull(),
         prop = n/tot,
         text = scales::percent(prop, accuracy = .1))
```

```{r}
p_1 <- 
  prep_d_01 %>% 
  ggplot(
    aes(x = .cluster_5, y = n, fill = .cluster_5)
  ) +
  geom_col() +
  geom_text(data = prep_d_02, aes(x = .cluster_5, y = .5*n, label = text), color = "white") +
  scale_y_continuous(labels = scales::comma) +
  ylab("Employees") +
  theme_light() +
  oescolorrrs::scale_fill_oes() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_1
```

# Clusters: How much telework?

```{r include=FALSE}
prep_01 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  summarise(num = mean(daily_status_teleworking)) %>% 
  ungroup()

prep_02 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  summarise(y = max(density(daily_status_teleworking)$y),
            x = density(daily_status_teleworking)$x[density(daily_status_teleworking)$y == max(density(daily_status_teleworking)$y)]
  )
```

```{r}
p_1 <- 
  prep_01 %>% 
  ggplot(
    aes(x = .cluster_5, y = num, fill = .cluster_5)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::comma) +
  ylab("Mean days of telework") +
  theme_light() +
  oescolorrrs::scale_fill_oes() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_2 <- 
  d %>% 
  ggplot(
    aes(x = daily_status_teleworking, color = .cluster_5, fill = .cluster_5)
  ) +
  geom_density(alpha = .5) +
  # geom_vline(data = prep_01, aes(xintercept = num, color = .cluster_5)) +
  geom_text(data=prep_02, aes(x=x, y=y, label=.cluster_5), size = 3, hjust = "inward") +
  xlab("Days of telework") +
  ylab("Density") +
  theme_light() +
  oescolorrrs::scale_fill_oes() +
  oescolorrrs::scale_color_oes() +
  theme(legend.position = "None")

p_1|p_2
```

# Clusters: How much telework? (proportion of working)

```{r include=FALSE}
prep_01 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  summarise(num = mean(daily_status_teleworking_working_prop)) %>% 
  ungroup()

prep_02 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  summarise(y = max(density(daily_status_teleworking_working_prop)$y),
            x = density(daily_status_teleworking_working_prop)$x[density(daily_status_teleworking_working_prop)$y == max(density(daily_status_teleworking_working_prop)$y)]
  )
```

```{r}
p_1 <- 
  prep_01 %>% 
  ggplot(
    aes(x = .cluster_5, y = num, fill = .cluster_5)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::comma) +
  ylab("Mean days of telework") +
  theme_light() +
  oescolorrrs::scale_fill_oes() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_2 <- 
  d %>% 
  ggplot(
    aes(x = daily_status_teleworking_working_prop, color = .cluster_5, fill = .cluster_5)
  ) +
  geom_density(alpha = .5) +
  # geom_vline(data = prep_01, aes(xintercept = num, color = .cluster_5)) +
  geom_text(data=prep_02, aes(x=x, y=y, label=.cluster_5), size = 3, hjust = "inward") +
  xlab("Days of telework") +
  ylab("Density") +
  theme_light() +
  oescolorrrs::scale_fill_oes() +
  oescolorrrs::scale_color_oes() +
  theme(legend.position = "None")

p_1|p_2
```

# Clusters: How much telework? (proportion of working) (rain)

```{r}
d %>% 
  ggplot(
    aes(x = 1, y = daily_status_teleworking_working_prop, fill = .cluster_5)
  ) +
 geom_rain(alpha = .5,
           rain.side = "r") +
  coord_flip()
```

# Clusters: How much telework? (proportion of working) (rain)

```{r}
prep_01 <-
  d %>%
  group_by(.cluster_5) %>%
  summarise(
    num = mean(daily_status_teleworking_working_prop),
    l = quantile(daily_status_teleworking_working_prop, probs = .25),
    u = quantile(daily_status_teleworking_working_prop, probs = .75)
  ) %>%
  ungroup()

# d %>%
  ggplot(
  ) +
  geom_jitter(data = d,
    aes(
      x = .cluster_5,
      y = daily_status_teleworking_working_prop,
      fill = .cluster_5,
      color = .cluster_5
    ),
    position = position_jitter(0.2), color = "darkgray") +
  geom_pointrange(
    aes(
      x = .cluster_5,
      fill = .cluster_5,
      color = .cluster_5,
      y = num,
      ymin = l,
      ymax = u
    ),
    data = prep_01
  )

# Combine with violin plots
# d %>% 
#   ggplot(df, aes(dose, len)) +
#   geom_violin(color = "darkgray", trim = FALSE) + 
#   geom_pointrange(aes(ymin = len-sd, ymax = len+sd), data = df.summary)
```

# Clusters: How much onsite work?

```{r include=FALSE}
prep_01 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  summarise(num = mean(daily_status_reporting_to_job_site)) %>% 
  ungroup()

prep_02 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  summarise(xPos = max(daily_status_reporting_to_job_site),
            yPos = max((density(daily_status_reporting_to_job_site))$y))

prep_02 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  summarise(y = max(density(daily_status_reporting_to_job_site)$y),
            x = density(daily_status_reporting_to_job_site)$x[density(daily_status_reporting_to_job_site)$y == max(density(daily_status_reporting_to_job_site)$y)]
  )
```

```{r}
p_1 <- 
  prep_01 %>% 
  ggplot(
    aes(x = .cluster_5, y = num, fill = .cluster_5)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::comma) +
  ylab("Mean days of onsite work") +
  theme_light() +
  oescolorrrs::scale_fill_oes() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_2 <- 
  d %>% 
  ggplot(
    aes(x = daily_status_reporting_to_job_site, color = .cluster_5, fill = .cluster_5)
  ) +
  geom_density(alpha = .5) +
  # geom_vline(data = prep_01, aes(xintercept = num, color = .cluster_5)) +
  geom_text(data=prep_02, aes(x=x, y=y, label=.cluster_5), size = 3, hjust = "inward") +
  xlab("Days of onsite work") +
  ylab("Density") +
  theme_light() +
  oescolorrrs::scale_fill_oes() +
  oescolorrrs::scale_color_oes() +
  theme(legend.position = "None")

p_1|p_2
```

# Clusters: Telework and onsite

```{r}
prep_01 <-
  d %>%
  mutate(
    .cluster_5 = fct_relabel(.cluster_5,~str_wrap(.,width = 50))
  ) %>% 
  group_by(.cluster_5) %>%
  summarise(across
            (
              .cols = c(
                daily_status_teleworking,
                daily_status_reporting_to_job_site,
                working_days
              ),
              .fns = mean
            )) %>%
  ungroup()
```

```{r}
prep_gg_1 <- d %>%
  mutate(
    .cluster_5 = fct_relabel(.cluster_5,~str_wrap(.,width = 50))
  ) %>% 
  ggplot(
    aes(
      x = daily_status_reporting_to_job_site,
      y = daily_status_teleworking,
      color = .cluster_5,
      fill = .cluster_5,
      linetype = .cluster_5
    )
  )
```

```{r}
p_1 <- 
  prep_gg_1 +
  geom_jitter(# aes(size = working_days),
    alpha = .5) +
  stat_ellipse(
    geom = "polygon",
    alpha = .15,
    size = 2,
    level = .99
  ) +
  geom_text_repel(
    data = prep_01,
    aes(x = daily_status_reporting_to_job_site, y = daily_status_teleworking, 
        label = .cluster_5,
        fill = .cluster_5),
    # fill = "white",
    # alpha = .15,
    size = 3,
    force             = 0.5,
    # nudge_x           = 100,
    xlim = c(125),
    ylim = c(50,150),
    direction         = "y",
    hjust             = 0,
    segment.size      = 0.2,
    segment.curvature = -0.1,
    color = "black"
  ) +
  xlab("Days reporting to the site") +
  ylab("Days teleworking") +
  theme_light() +
  oescolorrrs::scale_fill_oes() +
  oescolorrrs::scale_color_oes() +
  theme(legend.position = "None")
```

```{r}
p_1
```

```{r}
p_2 <- 
  prep_gg_1 +
  geom_jitter(# aes(size = working_days),
    alpha = .5,
    size = .5) +
  stat_ellipse(
    geom = "polygon",
    alpha = .15,
    size = .5,
    level = .99
  ) +
  geom_text_repel(
    data = prep_01,
    aes(x = daily_status_reporting_to_job_site, y = daily_status_teleworking, 
        label = .cluster_5,
        fill = .cluster_5),
    # fill = "white",
    # alpha = .15,
    size = 3,
    force             = 0.5,
    # nudge_x           = 100,
    xlim = c(100, 150),
    ylim = c(75,150),
    direction         = "y",
    hjust             = 0,
    segment.size      = 0.2,
    segment.curvature = -0.1,
    color = "black"
  ) +
  xlab("Days reporting to the site") +
  ylab("Days teleworking") +
  theme_light() +
  oescolorrrs::scale_fill_oes() +
  oescolorrrs::scale_color_oes() +
  theme(legend.position = "None",
        axis.title = element_text(size = 8))
```

```{r}
p_2
```

```{r}
file_name <- "abstract_1.png"
fp <- file.path(params$google_drive_home,"04. Analysis", "output", "png",file_name)
```

```{r}
ggsave(filename = fp, plot = p_2, width = 3, height = 3, units = "in")
```

# Disconnected workers: Working days

```{r include=FALSE}
prep_01 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  summarise(num = mean(working_days)) %>% 
  ungroup()

prep_02 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  summarise(xPos = max(working_days),
            yPos = max((density(working_days))$y))

prep_02 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  summarise(y = max(density(working_days)$y),
            x = density(working_days)$x[density(working_days)$y == max(density(working_days)$y)]
  )
```

```{r}
p_1 <- 
  prep_01 %>% 
  ggplot(
    aes(x = .cluster_5, y = num, fill = .cluster_5)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::comma) +
  ylab("Mean working days") +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_2 <- 
  d %>% 
  ggplot(
    aes(x = working_days, color = .cluster_5, fill = .cluster_5)
  ) +
  geom_density(alpha = .5) +
  # geom_vline(data = prep_01, aes(xintercept = num, color = .cluster_5)) +
  geom_text(data=prep_02, aes(x=x, y=y, label=.cluster_5), size = 3, hjust = "inward") +
  xlab("Working days") +
  ylab("Density") +
  theme_light() +
  theme(legend.position = "None")

p_1|p_2
```

# Disconnected workers: Exiting employment

```{r include=FALSE}
prep_01 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  summarise(num = sum(exit_employment)) %>% 
  ungroup()

prep_02 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  count() %>% 
  ungroup()

prep_03 <- 
  prep_01 %>% 
  full_join(prep_02) %>% 
  mutate(
    num = num/n
  )
```

```{r}
p_1 <-
  prep_03 %>%
  ggplot(aes(x = .cluster_5, y = num, fill = .cluster_5)) +
  geom_col() +
  ylab("Percent exiting employment") +
  scale_y_continuous(labels = scales::percent) +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8)
  )

p_1
```

# Clusters and position categories

```{r}
prep_d_01 <- 
  d %>% 
  group_by(position_category_description) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n))

prep_v_01 <- 
  prep_d_01 %>% 
  slice(3) %>% 
  pull(n)

prep_d_02 <-
  d %>%
  mutate(position_category_description = fct_lump_min(position_category_description, min = prep_v_01),
         position_category_description = fct_infreq(position_category_description)) %>%
  group_by(position_category_description,.cluster_5) %>%
  count() %>%
  ungroup() %>%
  arrange(desc(n)) %>% 
  rename(num = n) %>% 
  mutate(
    position_category_description = fct_relevel(position_category_description, 
                                               "Offsite",
                                               "Onsite Flexible",
                                               "Onsite Required",
                                               "Other")
  )

prep_d_03 <- 
  prep_d_02 %>% 
  filter(position_category_description != "Other")

prep_d_04 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  rename(tot = n)

prep_d_05 <- 
  prep_d_03 %>% 
  full_join(prep_d_04) %>% 
  mutate(
    n = num/tot,
    text = scales::comma(num)
  ) 
```

```{r}
p_1 <- 
  prep_d_05 %>% 
  ggplot(
    aes(x = .cluster_5, y = n, fill = .cluster_5)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  ylab("Percent of cluster in position category") +
  facet_wrap(facets = vars(position_category_description), nrow = 1) +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_1
```

# Clusters and position categories (alluvial)

```{r}
is_alluvia_form(prep_d_05)
```

```{r}
p_1 <- 
  prep_d_05 %>% 
  mutate(.cluster_5 = fct_relabel(.cluster_5,~str_wrap(.,width = 50))) %>% 
  ggplot(aes(y = num, axis1 = .cluster_5, axis2 = position_category_description, fill = .cluster_5)) +
  geom_alluvium(width = 1/12) +
  geom_stratum(width = 1/12,
               # alpha = 1,
               fill = "black",
               color = "white"
               ) +
  geom_label(stat = "stratum", aes(label = after_stat(stratum)),
             fill = "white",
             # , alpha = .05,
             hjust = "inward"
             ) +
  ylab("Number of employees") +
  scale_y_continuous(labels = scales::comma) +
  oescolorrrs::scale_fill_oes() +
  oescolorrrs::scale_color_oes() +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_blank())
```

```{r}
p_1
```

```{r}
p_2 <- 
  prep_d_05 %>% 
  mutate(.cluster_5 = fct_relabel(.cluster_5,~str_wrap(.,width = 50))) %>% 
  ggplot(aes(y = num, axis1 = .cluster_5, axis2 = position_category_description, fill = .cluster_5)) +
  geom_alluvium(width = 1/12) +
  geom_stratum(width = 1/12,
               # alpha = 1,
               fill = "black",
               color = "white"
               ) +
  geom_label(stat = "stratum", aes(label = after_stat(stratum)),
             fill = "white",
             # , alpha = .05,
              size = 2,
             hjust = "inward"
             ) +
  ylab("Number of employees") +
  scale_y_continuous(labels = scales::comma) +
  oescolorrrs::scale_fill_oes() +
  oescolorrrs::scale_color_oes() +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.y = element_text(size = 8),
    axis.title.x = element_blank(),
    axis.text.x = element_blank())
```

```{r}
p_2
```

```{r}
file_name <- "abstract_2.png"
fp <- file.path(params$google_drive_home,"04. Analysis", "output", "png",file_name)
```

```{r}
ggsave(filename = fp, plot = p_2, width = 3, height = 3, units = "in")
```

# Telecommuters, office visitors, and regions

```{r}
prep_d_01 <- 
  d %>% 
  group_by(gsa_region,.cluster_5) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  rename(num = n)

prep_d_02 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  rename(tot = n)

prep_d_03 <- 
  prep_d_01 %>% 
  full_join(prep_d_02) %>% 
  mutate(
    n = num/tot,
    text = scales::comma(num)
  ) %>% 
  mutate(
    gsa_region = fct_reorder(gsa_region,as.numeric(gsa_region))
  )

prep_d_04 <- 
  prep_d_03 %>% 
  filter(.cluster_5 == "Telecommuters" |
           .cluster_5 == "Office\nvisitors"
           )
```

```{r}
p_1 <- 
  prep_d_04 %>% 
  ggplot(
    aes(x = gsa_region, y = n, fill = gsa_region)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  ylab("Percent of cluster in region") +
  facet_wrap(facets = vars(.cluster_5), nrow = 1) +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_1
```

# Hybrid, onsite, and disconnected workers and regions

```{r}
prep_d_01 <- 
  d %>% 
  group_by(gsa_region,.cluster_5) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  rename(num = n)

prep_d_02 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  rename(tot = n)

prep_d_03 <- 
  prep_d_01 %>% 
  full_join(prep_d_02) %>% 
  mutate(
    n = num/tot,
    text = scales::comma(num)
  ) %>% 
  mutate(
    gsa_region = fct_reorder(gsa_region,as.numeric(gsa_region))
  )

prep_d_04 <- 
  prep_d_03 %>% 
  filter(.cluster_5 == "Hybrid\nworkers" |
           .cluster_5 == "Onsite\nworkers" |
           .cluster_5 == "Disconnected\nworkers")
```

```{r}
p_1 <- 
  prep_d_04 %>% 
  ggplot(
    aes(x = gsa_region, y = n, fill = gsa_region)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  ylab("Percent of cluster in region") +
  facet_wrap(facets = vars(.cluster_5), nrow = 1) +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_1
```

# States: MD, PA, and VA

```{r}
prep_d_01 <- 
  d %>% 
  group_by(location_state,.cluster_5) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  rename(num = n)

prep_d_02 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  rename(tot = n)

prep_d_03 <- 
  prep_d_01 %>% 
  full_join(prep_d_02) %>% 
  mutate(
    n = num/tot,
    text = scales::comma(num)
  )# %>% 
  # mutate(
  #   gsa_region = fct_reorder(gsa_region,as.numeric(gsa_region))
  # )

prep_d_04 <- 
  prep_d_03 %>% 
  filter(location_state %in% c("MD","PA","VA")
           )
```

```{r}
p_1 <- 
  prep_d_04 %>% 
  ggplot(
    aes(x = .cluster_5, y = n, fill = .cluster_5)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  ylab("Percent of cluster in state") +
  facet_wrap(facets = vars(location_state), nrow = 1) +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_1
```

# States: CA and IL

```{r}
prep_d_01 <- 
  d %>% 
  group_by(location_state,.cluster_5) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  rename(num = n)

prep_d_02 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  rename(tot = n)

prep_d_03 <- 
  prep_d_01 %>% 
  full_join(prep_d_02) %>% 
  mutate(
    n = num/tot,
    text = scales::comma(num)
  )# %>% 
  # mutate(
  #   gsa_region = fct_reorder(gsa_region,as.numeric(gsa_region))
  # )

prep_d_04 <- 
  prep_d_03 %>% 
  filter(location_state %in% c("CA", "IL")
           )
```

```{r}
p_1 <- 
  prep_d_04 %>% 
  ggplot(
    aes(x = .cluster_5, y = n, fill = .cluster_5)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  ylab("Percent of cluster in state") +
  facet_wrap(facets = vars(location_state), nrow = 1) +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_1
```

# District of Columbia

```{r}
prep_d_01 <- 
  d %>% 
  group_by(location_state,.cluster_5) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  rename(num = n)

prep_d_02 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  rename(tot = n)

prep_d_03 <- 
  prep_d_01 %>% 
  full_join(prep_d_02) %>% 
  mutate(
    n = num/tot,
    text = scales::comma(num)
  )# %>% 
  # mutate(
  #   gsa_region = fct_reorder(gsa_region,as.numeric(gsa_region))
  # )

prep_d_04 <- 
  prep_d_03 %>% 
  filter(location_state %in% c("DC")
           )
```

```{r}
p_1 <- 
  prep_d_04 %>% 
  ggplot(
    aes(x = .cluster_5, y = n, fill = .cluster_5)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  ylab("Percent of cluster in state") +
  facet_wrap(facets = vars(location_state), nrow = 1) +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_1
```

# Clusters and staff offices

```{r}
prep_d_01 <- 
  d %>% 
  group_by(gsa_sso) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n))

prep_v_01 <- 
  prep_d_01 %>% 
  slice(2) %>% 
  pull(n)

prep_d_02 <-
  d %>%
  mutate(gsa_sso = fct_lump_min(gsa_sso, min = prep_v_01),
         gsa_sso = fct_infreq(gsa_sso)) %>%
  group_by(gsa_sso,.cluster_5) %>%
  count() %>%
  ungroup() %>%
  arrange(desc(n)) %>% 
  rename(num = n) %>% 
  mutate(gsa_sso = fct_relevel(gsa_sso,
                               "PBS",
                               "FAS",
                               "OCFO",
                               "Other"))

prep_d_03 <- 
  prep_d_02 %>% 
  filter(gsa_sso != "Other")

prep_d_04 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  rename(tot = n)

prep_d_05 <- 
  prep_d_03 %>% 
  full_join(prep_d_04) %>% 
  mutate(
    n = num/tot,
    text = scales::comma(num)
  ) 
```

```{r}
p_1 <- 
  prep_d_05 %>% 
  ggplot(
    aes(x = .cluster_5, y = n, fill = .cluster_5)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  ylab("Percent of cluster in staff office") +
  facet_wrap(facets = vars(gsa_sso), nrow = 1) +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 7))

p_1
```

# Clusters and series

```{r}
prep_d_01 <- 
  d %>% 
  group_by(occupational_series) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n))

prep_v_01 <- 
  prep_d_01 %>% 
  slice(3) %>% 
  pull(n)

prep_d_02 <-
  d %>%
  mutate(
    # occupational_series = fct_lump_min(occupational_series, min = prep_v_01),
    occupational_series = fct_infreq(occupational_series)
  ) %>%
  group_by(occupational_series, .cluster_5) %>%
  count() %>%
  ungroup() %>%
  arrange(desc(n)) %>%
  rename(num = n) %>%
  mutate(
    # occupational_series = fct_relevel(occupational_series,
    #                                   "1102",
    #                                   "1101",
    #                                   "0343",
    #                                   "Other"),
    occupational_series = recode(occupational_series,
                                 "1102" = "Contracting",
                                 "1101" = "General Business and Industry",
                                 "0343" = "Management and Program Analysis",
                                 "0501" = "Financial Administration and Program",
                                 "0560" = "Budget Analysis",
                                 "0301" = "Miscellaneous Administration and Program",
                                 "1170" = "Realty",
                                 "1176" = "Building Management",
                                 "2150" = "Transportation Operations",
                                 "2210" = "Information Technology Management"
                                 ),
    occupational_series = fct_relevel(
      occupational_series,
      "Budget Analysis",
      "Building Management",
      "Contracting",
      "Financial Administration and Program",
      "General Business and Industry",
      "Information Technology Management",
      "Management and Program Analysis",
      "Miscellaneous Administration and Program",
      "Realty",
      "Transportation Operations",
    )
  )

# prep_d_02
```

```{r}
prep_d_03 <-
  prep_d_02 %>%
  filter(
    occupational_series %in%
      c(
        # "Budget Analysis",
        # "Building Management",
        "Contracting",
        # "Financial Administration and Program",
        # "General Business and Industry",
        # "Information Technology Management",
        "Management and Program Analysis",
        "Miscellaneous Administration and Program"
        # "Realty",
        # "Transportation Operations"
      )
  )

prep_d_04 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  rename(tot = n)

prep_d_05 <- 
  prep_d_03 %>% 
  full_join(prep_d_04) %>% 
  mutate(
    n = num/tot,
    text = scales::comma(num)
  ) %>% 
  mutate(
    occupational_series = fct_relabel(occupational_series,~str_wrap(.,width = 15))
  )
```

```{r}
p_1 <- 
  prep_d_05 %>% 
  ggplot(
    aes(x = .cluster_5, y = n, fill = .cluster_5)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  ylab("Percent of cluster in series") +
  facet_wrap(facets = vars(occupational_series), nrow = 1) +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 7))

p_1
```

# Clusters and series

```{r}
prep_d_01 <- 
  d %>% 
  group_by(occupational_series) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n))

prep_v_01 <- 
  prep_d_01 %>% 
  slice(3) %>% 
  pull(n)

prep_d_02 <-
  d %>%
  mutate(
    # occupational_series = fct_lump_min(occupational_series, min = prep_v_01),
    occupational_series = fct_infreq(occupational_series)
  ) %>%
  group_by(occupational_series, .cluster_5) %>%
  count() %>%
  ungroup() %>%
  arrange(desc(n)) %>%
  rename(num = n) %>%
  mutate(
    # occupational_series = fct_relevel(occupational_series,
    #                                   "1102",
    #                                   "1101",
    #                                   "0343",
    #                                   "Other"),
    occupational_series = recode(occupational_series,
                                 "1102" = "Contracting",
                                 "1101" = "General Business and Industry",
                                 "0343" = "Management and Program Analysis",
                                 "0501" = "Financial Administration and Program",
                                 "0560" = "Budget Analysis",
                                 "0301" = "Miscellaneous Administration and Program",
                                 "1170" = "Realty",
                                 "1176" = "Building Management",
                                 "2150" = "Transportation Operations",
                                 "2210" = "Information Technology Management"
                                 ),
    occupational_series = fct_relevel(
      occupational_series,
      "Budget Analysis",
      "Building Management",
      "Contracting",
      "Financial Administration and Program",
      "General Business and Industry",
      "Information Technology Management",
      "Management and Program Analysis",
      "Miscellaneous Administration and Program",
      "Realty",
      "Transportation Operations",
    )
  )

# prep_d_02
```

```{r}
prep_d_03 <-
  prep_d_02 %>%
  filter(
    occupational_series %in%
      c(
        "Budget Analysis",
        # "Building Management",
        # "Contracting",
        "Financial Administration and Program",
        # "General Business and Industry",
        "Information Technology Management"
        # "Management and Program Analysis",
        # "Miscellaneous Administration and Program"
        # "Realty",
        # "Transportation Operations"
      )
  )

prep_d_04 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  rename(tot = n)

prep_d_05 <- 
  prep_d_03 %>% 
  full_join(prep_d_04) %>% 
  mutate(
    n = num/tot,
    text = scales::comma(num)
  ) %>% 
  mutate(
    occupational_series = fct_relabel(occupational_series,~str_wrap(.,width = 15))
  )
```

```{r}
p_1 <- 
  prep_d_05 %>% 
  ggplot(
    aes(x = .cluster_5, y = n, fill = .cluster_5)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  ylab("Percent of cluster in series") +
  facet_wrap(facets = vars(occupational_series), nrow = 1) +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 7))

p_1
```

# Clusters and series

```{r}
prep_d_01 <- 
  d %>% 
  group_by(occupational_series) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n))

prep_v_01 <- 
  prep_d_01 %>% 
  slice(3) %>% 
  pull(n)

prep_d_02 <-
  d %>%
  mutate(
    # occupational_series = fct_lump_min(occupational_series, min = prep_v_01),
    occupational_series = fct_infreq(occupational_series)
  ) %>%
  group_by(occupational_series, .cluster_5) %>%
  count() %>%
  ungroup() %>%
  arrange(desc(n)) %>%
  rename(num = n) %>%
  mutate(
    # occupational_series = fct_relevel(occupational_series,
    #                                   "1102",
    #                                   "1101",
    #                                   "0343",
    #                                   "Other"),
    occupational_series = recode(occupational_series,
                                 "1102" = "Contracting",
                                 "1101" = "General Business and Industry",
                                 "0343" = "Management and Program Analysis",
                                 "0501" = "Financial Administration and Program",
                                 "0560" = "Budget Analysis",
                                 "0301" = "Miscellaneous Administration and Program",
                                 "1170" = "Realty",
                                 "1176" = "Building Management",
                                 "2150" = "Transportation Operations",
                                 "2210" = "Information Technology Management"
                                 ),
    occupational_series = fct_relevel(
      occupational_series,
      "Budget Analysis",
      "Building Management",
      "Contracting",
      "Financial Administration and Program",
      "General Business and Industry",
      "Information Technology Management",
      "Management and Program Analysis",
      "Miscellaneous Administration and Program",
      "Realty",
      "Transportation Operations",
    )
  )

# prep_d_02
```

```{r}
prep_d_03 <-
  prep_d_02 %>%
  filter(
    occupational_series %in%
      c(
        # "Budget Analysis",
        "Building Management",
        # "Contracting",
        # "Financial Administration and Program",
        "General Business and Industry"
        # "Information Technology Management",
        # "Management and Program Analysis",
        # "Miscellaneous Administration and Program"
        # "Realty",
        # "Transportation Operations"
      )
  )

prep_d_04 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  rename(tot = n)

prep_d_05 <- 
  prep_d_03 %>% 
  full_join(prep_d_04) %>% 
  mutate(
    n = num/tot,
    text = scales::comma(num)
  )  %>% 
  mutate(
    occupational_series = fct_relabel(occupational_series,~str_wrap(.,width = 15))
  )
```

```{r}
p_1 <- 
  prep_d_05 %>% 
  ggplot(
    aes(x = .cluster_5, y = n, fill = .cluster_5)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  ylab("Percent of cluster in series") +
  facet_wrap(facets = vars(occupational_series), nrow = 1) +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 7))

p_1
```

# Clusters and series

```{r}
prep_d_01 <- 
  d %>% 
  group_by(occupational_series) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n))

prep_v_01 <- 
  prep_d_01 %>% 
  slice(3) %>% 
  pull(n)

prep_d_02 <-
  d %>%
  mutate(
    # occupational_series = fct_lump_min(occupational_series, min = prep_v_01),
    occupational_series = fct_infreq(occupational_series)
  ) %>%
  group_by(occupational_series, .cluster_5) %>%
  count() %>%
  ungroup() %>%
  arrange(desc(n)) %>%
  rename(num = n) %>%
  mutate(
    # occupational_series = fct_relevel(occupational_series,
    #                                   "1102",
    #                                   "1101",
    #                                   "0343",
    #                                   "Other"),
    occupational_series = recode(occupational_series,
                                 "1102" = "Contracting",
                                 "1101" = "General Business and Industry",
                                 "0343" = "Management and Program Analysis",
                                 "0501" = "Financial Administration and Program",
                                 "0560" = "Budget Analysis",
                                 "0301" = "Miscellaneous Administration and Program",
                                 "1170" = "Realty",
                                 "1176" = "Building Management",
                                 "2150" = "Transportation Operations",
                                 "2210" = "Information Technology Management"
                                 ),
    occupational_series = fct_relevel(
      occupational_series,
      "Budget Analysis",
      "Building Management",
      "Contracting",
      "Financial Administration and Program",
      "General Business and Industry",
      "Information Technology Management",
      "Management and Program Analysis",
      "Miscellaneous Administration and Program",
      "Realty",
      "Transportation Operations",
    )
  )

# prep_d_02
```

```{r}
prep_d_03 <-
  prep_d_02 %>%
  filter(
    occupational_series %in%
      c(
        # "Budget Analysis",
        # "Building Management",
        # "Contracting",
        # "Financial Administration and Program",
        # "General Business and Industry",
        # "Information Technology Management",
        # "Management and Program Analysis",
        # "Miscellaneous Administration and Program"
        "Realty",
        "Transportation Operations"
      )
  )

prep_d_04 <- 
  d %>% 
  group_by(.cluster_5) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  rename(tot = n)

prep_d_05 <- 
  prep_d_03 %>% 
  full_join(prep_d_04) %>% 
  mutate(
    n = num/tot,
    text = scales::comma(num)
  )  %>% 
  mutate(
    occupational_series = fct_relabel(occupational_series,~str_wrap(.,width = 15))
  )
```

```{r}
p_1 <- 
  prep_d_05 %>% 
  ggplot(
    aes(x = .cluster_5, y = n, fill = .cluster_5)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  ylab("Percent of cluster in series") +
  facet_wrap(facets = vars(occupational_series), nrow = 1) +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 7))

p_1
```

```{r}
states()
```

# TIL

-   Telework is very common - most GSA employees telework

-   Telecommuters and office visitors are very common for those that live in states neighboring DC

-   Employees living in DC are more likely to come to the office

-   PBS has considerable more onsite work than other offices

# Appendix

# Position Categories: How much telework?

```{r include=FALSE}
prep_d_01 <- 
  d %>% 
  group_by(position_category_description) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n))

prep_v_01 <- 
  prep_d_01 %>% 
  slice(3) %>% 
  pull(n)

prep_01 <- 
  d %>% 
   mutate(position_category_description = fct_lump_min(position_category_description, min = prep_v_01),
         position_category_description = fct_infreq(position_category_description)) %>%
  group_by(position_category_description) %>% 
  summarise(num = mean(daily_status_teleworking)) %>% 
  ungroup()

prep_02 <- 
  d %>% 
   mutate(position_category_description = fct_lump_min(position_category_description, min = prep_v_01),
         position_category_description = fct_infreq(position_category_description)) %>%
  group_by(position_category_description) %>% 
  summarise(y = max(density(daily_status_teleworking)$y),
            x = density(daily_status_teleworking)$x[density(daily_status_teleworking)$y == max(density(daily_status_teleworking)$y)]
  )
```

```{r}
p_1 <- 
  prep_01 %>% 
  ggplot(
    aes(x = position_category_description, y = num, fill = position_category_description)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::comma) +
  ylab("Mean days of telework") +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_2 <- 
  d %>% 
  mutate(position_category_description = fct_lump_min(position_category_description, min = prep_v_01),
         position_category_description = fct_infreq(position_category_description)) %>%
  ggplot(
    aes(x = daily_status_teleworking, color = position_category_description, fill = position_category_description)
  ) +
  geom_density(alpha = .5) +
  # geom_vline(data = prep_01, aes(xintercept = num, color = position_category_description)) +
  geom_text(data=prep_02, aes(x=x, y=y, label=position_category_description), size = 3, hjust = "inward") +
  xlab("Days of telework") +
  ylab("Density") +
  theme_light() +
  theme(legend.position = "None")

p_1|p_2
```

# Position Categories: How much onsite work?

```{r include=FALSE}
prep_d_01 <- 
  d %>% 
  group_by(position_category_description) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n))

prep_v_01 <- 
  prep_d_01 %>% 
  slice(3) %>% 
  pull(n)

prep_01 <- 
  d %>% 
   mutate(position_category_description = fct_lump_min(position_category_description, min = prep_v_01),
         position_category_description = fct_infreq(position_category_description)) %>%
  group_by(position_category_description) %>% 
  summarise(num = mean(daily_status_reporting_to_job_site)) %>% 
  ungroup()

prep_02 <- 
  d %>% 
   mutate(position_category_description = fct_lump_min(position_category_description, min = prep_v_01),
         position_category_description = fct_infreq(position_category_description)) %>%
  group_by(position_category_description) %>% 
  summarise(y = max(density(daily_status_reporting_to_job_site)$y),
            x = density(daily_status_reporting_to_job_site)$x[density(daily_status_reporting_to_job_site)$y == max(density(daily_status_reporting_to_job_site)$y)]
  )
```

```{r}
p_1 <- 
  prep_01 %>% 
  ggplot(
    aes(x = position_category_description, y = num, fill = position_category_description)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::comma) +
  ylab("Mean days of onsite work") +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_2 <- 
  d %>% 
  mutate(position_category_description = fct_lump_min(position_category_description, min = prep_v_01),
         position_category_description = fct_infreq(position_category_description)) %>%
  ggplot(
    aes(x = daily_status_reporting_to_job_site, color = position_category_description, fill = position_category_description)
  ) +
  geom_density(alpha = .5) +
  # geom_vline(data = prep_01, aes(xintercept = num, color = position_category_description)) +
  geom_text(data=prep_02, aes(x=x, y=y, label=position_category_description), size = 3, hjust = "inward") +
  xlab("Days of onsite work") +
  ylab("Density") +
  theme_light() +
  theme(legend.position = "None")

p_1|p_2
```

# Rule-based groups: How much telework?

```{r}
prep_d_01 <-
  d %>%
  select(person_id, 
         rule_based_primarily_teleworking_working,
         rule_based_mostly_teleworking_working,
         rule_based_mostly_reporting_to_job_site_working,
         rule_based_primarily_reporting_to_job_site_working,
         rule_based_all_remote,
         rule_based_telework_onsite_minimum,
         daily_status_teleworking) %>% 
  pivot_longer(starts_with("rule_based_")) %>% 
  filter(value == 1) %>% 
  mutate(
    name = str_remove(name, "rule_based_"),
    name = str_remove(name, "_working"),
    name = str_replace_all(name, "_", " "),
    name = str_to_sentence(name),
    name = fct_relevel(
      name,
      "Primarily teleworking",
      "Mostly teleworking",
      "Mostly reporting to job site",
      "Primarily reporting to job site",
      "All remote",
      "Telework onsite minimum"
    ),
    name = fct_relabel(name, ~str_wrap(.,width = 20))
  ) %>% 
  arrange(name)

prep_01 <- 
  prep_d_01 %>% 
  group_by(name) %>% 
  summarise(num = mean(daily_status_teleworking)) %>% 
  ungroup()

prep_02 <- 
  prep_d_01 %>% 
  group_by(name) %>% 
  summarise(y = max(density(daily_status_teleworking)$y),
            x = density(daily_status_teleworking)$x[density(daily_status_teleworking)$y == max(density(daily_status_teleworking)$y)]
  )
```

```{r}
p_1 <- 
  prep_01 %>% 
  ggplot(
    aes(x = name, y = num, fill = name)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::comma) +
  ylab("Mean days of telework") +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_2 <- 
  prep_d_01 %>% 
  ggplot(
    aes(x = daily_status_teleworking, color = name, fill = name)
  ) +
  geom_density(alpha = .5) +
  # geom_vline(data = prep_01, aes(xintercept = num, color = name)) +
  geom_text(data=prep_02, aes(x=x, y=y, label=name), size = 3, hjust = "inward") +
  xlab("Days of telework") +
  ylab("Density") +
  theme_light() +
  theme(legend.position = "None")

p_1|p_2
```

# Rule-based groups: How much telework?

```{r}
prep_d_01 <-
  d %>%
  select(person_id, 
         rule_based_primarily_teleworking_working,
         rule_based_mostly_teleworking_working,
         rule_based_mostly_reporting_to_job_site_working,
         rule_based_primarily_reporting_to_job_site_working,
         rule_based_all_remote,
         rule_based_telework_onsite_minimum,
         daily_status_reporting_to_job_site) %>% 
  pivot_longer(starts_with("rule_based_")) %>% 
  filter(value == 1) %>% 
  mutate(
    name = str_remove(name, "rule_based_"),
    name = str_remove(name, "_working"),
    name = str_replace_all(name, "_", " "),
    name = str_to_sentence(name),
    name = fct_relevel(
      name,
      "Primarily teleworking",
      "Mostly teleworking",
      "Mostly reporting to job site",
      "Primarily reporting to job site",
      "All remote",
      "Telework onsite minimum"
    ),
    name = fct_relabel(name, ~str_wrap(.,width = 20))
  ) %>% 
  arrange(name)

prep_01 <- 
  prep_d_01 %>% 
  group_by(name) %>% 
  summarise(num = mean(daily_status_reporting_to_job_site)) %>% 
  ungroup()

prep_02 <- 
  prep_d_01 %>% 
  group_by(name) %>% 
  summarise(y = max(density(daily_status_reporting_to_job_site)$y),
            x = density(daily_status_reporting_to_job_site)$x[density(daily_status_reporting_to_job_site)$y == max(density(daily_status_reporting_to_job_site)$y)]
  )
```

```{r}
p_1 <- 
  prep_01 %>% 
  ggplot(
    aes(x = name, y = num, fill = name)
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::comma) +
  ylab("Mean days of onsite work") +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_2 <- 
  prep_d_01 %>% 
  ggplot(
    aes(x = daily_status_reporting_to_job_site, color = name, fill = name)
  ) +
  geom_density(alpha = .5) +
  # geom_vline(data = prep_01, aes(xintercept = num, color = name)) +
  geom_text(data=prep_02, aes(x=x, y=y, label=name), size = 3, hjust = "inward") +
  xlab("Days of onsite work") +
  ylab("Density") +
  theme_light() +
  theme(legend.position = "None")

p_1|p_2
```

# Onsite Flexible: Rule-based groups

```{r}
prep_d_01 <- 
  d %>%
  filter(position_category_description == "Onsite Flexible") %>% 
  summarise(across(
    .cols = starts_with("rule_based"),
    .fns = ~ sum(., rm.na = T)
  )) %>%
  select(-contains("_position_category_")) %>%
  select(-ends_with("observation")) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  select(-rowname) %>%
  mutate(
    name = str_remove(name, "rule_based_"),
    name = str_remove(name, "_working"),
    name = str_replace_all(name, "_", " "),
    name = str_to_sentence(name),
    name = fct_relevel(
      name,
      "Primarily teleworking",
      "Mostly teleworking",
      "Mostly reporting to job site",
      "Primarily reporting to job site",
      "All remote",
      "Telework onsite minimum"
    ),
    name = fct_relabel(name, ~str_wrap(.,width = 20))
  ) %>% 
  arrange(name)

prep_d_02 <- 
  prep_d_01 %>% 
  mutate(tot = d %>% count() %>% pull(),
         prop = value/tot,
         text = scales::percent(prop, accuracy = .1))
```

```{r}
p_1 <- 
  prep_d_01 %>% 
  ggplot(
    aes(x = name, y = value, fill = name)
  ) +
  geom_col() +
  geom_text(data = prep_d_02, aes(x = name, y = .5*value, label = text), color = "white") +
  scale_y_continuous(labels = scales::comma) +
  ylab("Employees") +
  theme_light() +
  theme(
    legend.position = "None",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 8))

p_1
```

# CONTINUE FROM HERE

```{r}
beepr::beep(sound = 5)
```

```{r}

```